### 2022-01-27 _posts/welcome/welcome.Rmd
Add Distill skeleton

---
title: "Welcome to Steve Linberg"
description: |
  Welcome to our new blog, Steve Linberg. We hope you enjoy 
  reading what we have to say!
author:
  - name: Nora Jones 
    url: https://example.com/norajones
    affiliation: Spacely Sprockets
    affiliation_url: https://example.com/spacelysprokets
date: 01-27-2022
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Learn more about creating blogs with Distill at:
# https://rstudio.github.io/distill/blog.html

```

### 2022-01-28 _posts/welcome/welcome.Rmd
Edit default welcome post

@@
 ---
-title: "Welcome to Steve Linberg"
+title: "New blog under construction"
 description: |
-  Welcome to our new blog, Steve Linberg. We hope you enjoy 
-  reading what we have to say!
+  We're re-launching slinberg.net using Distill. Stand by!
 author:
-  - name: Nora Jones 
-    url: https://example.com/norajones
-    affiliation: Spacely Sprockets
-    affiliation_url: https://example.com/spacelysprokets
+  - name: Steve Linberg
+    url: https://slinberg.net
+    # affiliation: Spacely Sprockets
+    # affiliation_url: https://example.com/spacelysprokets
 date: 01-27-2022
@@
 ```
+
+This is a re-launch of slinberg.net, using the [Distill](https://rstudio.github.io/distill) platform to align with the UMASS DACSS program I am presently enrolled in.

### 2022-01-28 _posts/welcome/welcome.Rmd
update welcome and rebuild

@@
 
-This is a re-launch of slinberg.net, using the [Distill](https://rstudio.github.io/distill) platform to align with the UMASS DACSS program I am presently enrolled in.
+This is a re-launch of slinberg.net, using the [distill](https://rstudio.github.io/distill) platform to align with the [UMASS DACSS program](https://www.umass.edu/sbs/data-analytics-and-computational-social-science-program/ms) I am presently enrolled in.
+
+Pardon the dust while we redecorate. First up will be a soup-to-nuts, zero-to-hero walkthrough of how this site was built and deployed.

### 2022-01-29 _posts/welcome/welcome.Rmd
build
Add preview image to welcome post

@@
     self_contained: false
+preview: construction.png
 ---
@@
 
-Pardon the dust while we redecorate. First up will be a soup-to-nuts, zero-to-hero walkthrough of how this site was built and deployed.
+Pardon the dust while we redecorate. First up will be a soup-to-nuts, zero-to-hero walkthrough of how this site was built and deployed on a custom domain with GitHub.

### 2022-01-29 _meta/distill-and-github-pages-setup/distill-and-github-pages-setup.Rmd
build and dirs

---
title: "Distill and Github Pages Setup"
description: |
  A short description of the post.
author:
  - name: Steve Linberg
date: 2022-01-29
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

### 2022-01-29 _meta/distill-and-github-pages-setup/distill-and-github-pages-setup.Rmd
Add meta index page and categories

@@
 description: |
-  A short description of the post.
+  A walkthrough of setting up a new Distill site on GitHub Pages
 author:
@@
     self_contained: false
+categories:
+  - distill
+  - github pages
+  - git
+  - rstudio
 ---

### 2022-01-29 steve.Rmd
Add postcard page

---
title: "Steve Linberg"
image: "2021-sfl-photo-fadedbg-256.png"
links:
  - label: LinkedIn
    url: "https://www.linkedin.com/in/slinberg"
  - label: Twitter
    url: "https://twitter.com/stevelinberg"
  - label: GitHub
    url: "https://github.com/stevelinberg"
  - label: SO
    url: "https://stackoverflow.com/users/13603796/stevelinberg"
  - label: Email
    url: "mailto:steve@slinberg.net"
output:
  postcards::trestles
---

## Bio

I am a Masters student in the [DACSS program](https://www.umass.edu/sbs/data-analytics-and-computational-social-science-program) at the University of Massachusetts, Amherst; anticipated graduation Summer 2022. The day job is Senior Systems and Analytics Engineer at the [Gandara Center](https://www.gandaracenter.org/) in West Springfield, MA.

This site contains my academic [blog](index.html) with ruminations on computational social science, my own learning processes and discoveries, and related technical topics and snippets.

## Education

**Hampshire College** | Amherst, MA

BA, Social Science, 1990-3

**University of Massachusetts, Amherst** | Amherst, MA

M.S, Data Analytics and Computational Social Science | Anticipated graduation summer 2022

## Experience

**Gandara Center** | Senior Systems and Analytics Engineer / CIO / IT Director | 2005 - Present

**Silicon Goblin Technologies** | Owner / Programmer | 1999 - 2015

**National Center on Adult Literacy**, University of Pennsylvania Graduate School of Education | Systems Programmer | 1997-1999

**Holyoke Public Schools** | Literacy and GED Instructor | 1994-1997

**Unicom, Inc.** | Systems Programmer and Technical Writer | 1983-1997

## Interests

- Data analytics and statistics
- Programming languages
- The state of the world
- Drawing, painting, guitar, motorcycles, baking, gaming

### 2022-01-29 _meta/distill-and-github-pages-setup/distill-and-github-pages-setup.Rmd
Add first distill article

--- 2022-01-29 _meta/distill-and-github-pages-setup/distill-and-github-pages-setup.Rmd
@@
   distill::distill_article:
+    toc: true
     self_contained: false
+categories:
+  - distill
+  - github pages
+  - git
+  - rstudio
 ---
@@
 ```
+
+This document contains a walkthrough of setting up a new [Distill](https://rstudio.github.io/distill/) blog using [RStudio](https://www.rstudio.com/products/rstudio/) and [GitHub pages.](https://pages.github.com/) It's the first in a series of articles for social scientists showing how to create, publish and update a website using these packages.
+This is a relatively easy way to set up a lightweight website that's well-suited for blogging and writing articles, especially about and using R.
+Once it's built, you maintain it using RStudio just like your other projects.
+
+Although it's intended to be complete, and not overly technical where avoidable, it does assume a modest level of comfort using basic tools like RStudio and GitHub.
+If you're on Mac or Linux, you probably have [git](https://git-scm.com/) built in already, but you might need to install it if you're on Windows.
+That's outside the scope of this article, but you can find information about it on the main Git page.
+
+Assuming you have git and RStudio installed, and a GitHub account created, you're ready to proceed!
+
+## Create new GitHub repository
+The first step is to create a new repository for the site. Log into GitHub and create a new repo:
+
+![](Pasted image 20220127115109.png)
+
+Call it something sensible like "distill-blog"; make it public (unless you don't want to) and add a `.gitignore` file using the `R` template (just to start things off).
+Also, if you haven't already, change your [github repository settings](https://github.com/settings/repositories) to use "main" as your default branch name instead of "master".
+
+We'll come back in a bit to set up the Pages part, but for now we just want to get the project built and populated.
+We now have an empty Git repository sitting on GitHub, waiting to be used for something.
+
+## Create a new RStudio project using the repository
+Now we'll use our new, empty GitHub repository for our Distill blog. From RStudio, choose "File -> New Project" and choose "Version Control":
+
+![](Pasted image 20220127124549.png)
+
+Choose "Git":
+
+![](Pasted image 20220127124647.png)
+
+Copy in the repository URL from your project (found as shown below):
+
+![](Pasted image 20220127125016.png)
+
+Paste it into the `Repository URL` field in the wizard, choose where you want to store it on your computer, and I recommend checking `Open in new session` for a clean slate.
+Click "Create Project" to create it.
+
+![](Pasted image 20220127125055.png)
+
+The empty project will open in a new RStudio window.
+RStudio has added 3 new files (actually three directories) to the skeleton created by Github, which initially only contains the skeleton `.gitignore` file:
+
+![](Pasted image 20220127130928.png)
+
+RStudio adds the `.git` directory, the `.Rproj.user` directory, and the `distill-blog.Rproj` RStudio project file (it will be named with the "Project directory name" field from the final screen of the New Project Wizard).
+
+![](Pasted image 20220127131749.png)
+
+The `.gitignore` file created by Github tells git to ignore the `.Rproj.user` directory (which contains local settings used by RStudio and which you don't need to (and shouldn't) share or publish), and the `.git` directory is part of every repository and contains the repository settings.
+
+We want to keep github and RStudio in sync, generally speaking.
+More precisely, we want to keep *the repository* in sync between GitHub and RStudio.
+The local git repository doesn't yet know about the project file `distill-blog.Rproj`, since RStudio added that after cloning the repository from GitHub, so a good first task is to add it to the local repository, and push it back up to GitHub to get it in sync, and make sure that that core pipeline is working, since we'll be using it constantly.
+
+Although there is a "Git" pane in the Environments panel, which shows that `distill-blog.Rproj` is ready for staging:
+
+![](Pasted image 20220127133207.png)
+
+...I am an old-school curmudgeon and prefer to use the terminal for git commands, since not everything I normally do with git has a representation in this visual UI.
+In RStudio I have even remapped the shortcut control-backtick to the terminal for quick switching:
+
+![](Pasted image 20220127133440.png)
+
+Within the terminal, typing `git status` will show that git doesn't yet know about `distill-blog.Rpro`j:
+
+![](Pasted image 20220127133648.png)
+
+This isn't going to be a git tutorial, so we're just going to add the file to the repo and push it back to github.
+If you get stuck in this process, pause here until you're able to troubleshoot and continue.
+
+The git commands to update the repository with are:
+
+```sh
+git add .
+git commit -m "Add RStudio project skeleton"
+git push
+```
+
+(Note: the first command, `git add .`, says to add all files in the current directory to the repository.
+We could have explicitly typed `git add distill-blog.Rproj`, but (a) `git add .` is much faster, and (b) much easier in the steps below when we may not even know all of the files that need adding, and explicitly enumerating them could be troublesome and error-prone.
+Sometimes you want to be careful and specific about which files you're staging, but that's beyond the scope of this discussion, and we're keeping it simple here.)
+
+```sh
+[slinbergdistill-blog] (main)$ git add . 
+[slinbergdistill-blog] (main +)$ git commit -m "Add RStudio project skeleton"
+[main 6c67d2d] Add RStudio project skeleton
+ 1 file changed, 13 insertions(+)
+ create mode 100644 distill-blog.Rproj
+[slinbergdistill-blog] (main)$ git push Enter passphrase for key '/Users/slinberg/.ssh/sfl_id_rsa': 
+Enumerating objects: 4, done.
+Counting objects: 100% (4/4), done.
+Delta compression using up to 8 threads
+Compressing objects: 100% (3/3), done.
+Writing objects: 100% (3/3), 444 bytes | 444.00 KiB/s, done.
+Total 3 (delta 0), reused 0 (delta 0), pack-reused 0
+To github.com:stevelinberg/distill-blog.git
+   54aae63..6c67d2d  main -> main 
+```
+
+After `git push` uploads the changes back to the GitHub repository, now if we flip back to GitHub and refresh the page, we'll see that the project file has been added:
+
+![](Pasted image 20220127135410.png)
+
+At this point, we have an empty RStudio project committed to a GitHub repository, but we don't yet actually have any content.
+Now it's time to create the Distill blog skeleton and add it to the repository.
+
+## Add the Distill skeleton to the repository
+We will now manually (sort of) create the Distill skeleton.
+Although we could have created a new project of a Distill blog from the RStudio Create New Project wizard, it wouldn't have had the GitHub repository, and we would have needed to add that and do the github setup separately.
+Six of one, half-dozen of the other, but instead we created the GitHub repository first instead, and we'll add in Distill ourselves now.
+The end result should be the same.
+
+(Note: these commands are entered in the RStudio Console, which looks a lot like the Terminal, but speaks a different language.
+They're both areas where you type commands, but the difference is that the terminal is a shell that passes commands to the operating system, for doing things like creating directories, copying or deleting files, or issuing git commands, and the Console is an R-language construct that processes R commands.
+If you type an R command in the Terminal, or a shell command in the Console, they will both be confused and spit out an error message, but it won't do any damage beyond that.
+A glance at the tab titles will quickly orient you.)
+
+In the console, install the `distill` library if you haven't yet (reinstalling it won't do any harm):
+
+```r
+> install.packages("distill")
+trying URL 'https://cran.rstudio.com/bin/macosx/contrib/4.1/distill_1.3.tgz'
+Content type 'application/x-gzip' length 454710 bytes (444 KB)
+==================================================
+downloaded 444 KB
+
+The downloaded binary packages are in
+	/var/folders/jw/rmjpk2xj7z7_8txfqrjz0jhr0000gn/T//RtmpjDcuEY/downloaded_packages
+```
+
+Then load the library, which you should also do any time you want to work on your site in the future after opening the project:
+
+```r
+> library(distill)
+```
+
+With the `distill` library loaded, we're ready to create the site.
+One important point is that we're going to create the site in the root of our project, rather than inside a subdirectory.
+There's no need to bury it one level down, since it won't be a subcomponent of another project.
+The first argument in the command below, `"."`, means to work in the current directory, which is set by default as the directory containing the `distill-blog.Rproj` file.
+The argument is the blog title (which can be changed later if needed, but if you can avoid needing to do that it saves a few braincells), and the final argument causes it to be built with support for GitHub Pages, which is where our blog will ultimately end up.
+
+Here's the command I ran for this blog:
+
+```r
+> create_blog(".", "Steve Linberg", gh_pages = TRUE)
+```
+
+This takes a few seconds to run and creates a lot of new files.
+I like to go ahead and commit the project at this point as well, giving us a rollback point where we can wipe out changes we made and start over with a clean install if we get in trouble down the road.
+Switch back to the terminal, add all files in the directory, commit and push:
+
+```sh
+git add .
+git commit -m "Install Distill skeleton"
+git push
+```
+
+If we pop back over to GitHub and reload the repository view, we'll see a lot of new files added:
+
+![](Pasted image 20220127160301.png)
+
+## Build the site
+Before we do any further customizing, we can build the site and get it deploying on GitHub Pages.
+At this point, there's nothing to see because the site hasn't been built yet; doing the first build will construct the first blog post and the basic site skeleton.
+
+When you create the Distill site in this way, RStudio doesn't yet know that there is one, and so the Build tab may be missing in the Environment pane:
+
+![](Pasted image 20220127160912.png)
+
+If so, quit RStudio and re-open the project file, and it will appear:
+
+![](Pasted image 20220127161112.png)
+
+Click the Build tab and then "Build Website":
+
+![](Pasted image 20220127161335.png)
+
+It will grind for a few moments and do the first build, and present you with a glorious preview of your new blog in the Viewer panel:
+
+![](Pasted image 20220127161457.png)
+
+Let's commit this stage too, since a few new files were just created. Switch to the Terminal tab (ctrl + backtick) and issue these git commands:
+
+```sh
+git add .
+git commit -m "First build of site skeleton"
+git push
+```
+
+Breathtaking in its awesomeness, we are now ready to share it with the world!
+
+OK, you might prefer to do some work on it before you unveil it, and that's fine.
+When you're ready, continue below.
+(Read the RStudio page [Creating a Blog](https://rstudio.github.io/distill/blog.html?panelset=before-i-sleep) to learn the details of how Distill sites work, and how to add and customize content; like anything worth learning, it takes a bit of time, but the payoff will be your mighty blog powers.) In this case, we're going to go ahead and pretend we're done and test the deployment to make sure it all works.
+
+## Configuring GitHub Pages
+Next, we need to go back to GitHub and get GitHub Pages working, so we can publish our site.
+
+From your GutHub project page, as shown below:
+
+1. Click "Settings"
+2. On the left, click "Pages"
+3. Under "Source", set the branch to "main"
+4. Set the folder to "/docs"
+5. Click "Save".
+
+(Note: ignore the "Choose a theme" button; it doesn't apply to Distill projects.)
+
+![](Pasted image 20220127162652.png)
+
+GitHub will show a success message (hopefully):
+
+![](Pasted image 20220127163030.png)
+
+Note the "Custom domain" section at the bottom; we'll return to that if/when we want to move the site to a custom domain.
+As it is, GitHub generates a URL for the site based on your GitHub username and the repo name.
+In this case, the address is [https://stevelinberg.github.io/distill-blog/](https://stevelinberg.github.io/distill-blog/).
+Click the link, and you should see your site:
+
+![](Pasted image 20220127163306.png)
+
+This means it's live!
+
+You could just go from here, if you're content with the URL GitHub chose for you. Your basic workflow will be:
+- Work on content in RStudio
+- Commit changes with git as you go
+- Build the site when you're ready (or as often as you like)
+- Push the changes to GitHub, which publishes them to your site.
+
+If you want to brand a little more strongly and put the site on your own domain, continue with "Using a Custom Domain", the next article in this series, coming soon.

### 2022-01-29 _posts/2022-01-29-distill-setup-article-added/distill-setup-article-added.Rmd
fix paths and build

---
title: "Distill setup article added"
description: |
  The first Distill how-to article is up.
author:
  - name: Steve Linberg
    url: https://slinberg.net
date: 2022-01-29
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

[The first article about using Distill to build a website like this one](../../meta/distill-and-github-pages-setup/index.html) is now up in the [Meta](../../meta_index.html) section. More articles are coming!

### 2022-01-29 _posts/2022-01-29-distill-setup-article-added/distill-setup-article-added.Rmd
Text tweaks, build

@@
 description: |
-  The first Distill how-to article is up.
+  The first Distill how-to article is up in the Meta section.
 author:
@@
 More articles are coming!
+
+Note: the [source for this site](https://github.com/stevelinberg/distill-blog) can be found on github.

### 2022-01-29 _academics/ps753/ps753.Rmd
Add academics/PS753 page

---
title: "PS753"
description: |
  Front page for PS753 coursework, Spring 2022.
author:
  - name: Steve Linberg
date: 2022-01-29
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# PS753: Political and Social Networks, Spring 2022

From this index page, we will magically link to the glossary page.

### 2022-01-30 _academics/ps753-glossary/ps753-glossary.Rmd
PS753 glossary stuff

---
title: "PS753 Glossary"
description: |
  Glossary page for PS753 coursework
author:
  - name: Steve Linberg
    url: {}
date: 2022-01-29
output:
 distill::distill_article:
   self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

<p style="background: #eee">
[PS 753](../PS753/index.html)
&gt; Glossary
</p>

# Week 1

**From lecture 1a: [Nodes, Edges and Network Samples](https://www.loom.com/share/efc5252bb75b463481b93e55be6e4d67)**

node / vertex
: a junction in a network where two or more lines (edges) intersect; a dot connecting lines.<br>

edge / link
: lines connecting nodes, which indicate some sort of connection or relationship.

node or ego
: $i$ - the node being discussed or focused on

alter
: $j$ - the node that $i$ connects to

network population
: $n$ - the size of the population or count of nodes

With set notation, we define $i$ as a set of $n$ elements:
$$i \in 1, 2, 3 \dots n$$

and $j$ similarly as a set of $n$ elements, except that $j$ cannot equal $i (a node cannot connect to itself):
$$j \in 1, 2, 3 \dots n, i \ne j$$

interaction threshold
: a measure to determine whether two entities have a sufficient connection to be considered to have a link between them

snowball sample
: an entity group formed by starting with one node and "rolling outwards" to its connections until all nodes or actors (in a limited set) are located

egocentric name generator
: a mode of building an entity group (an egocentric network) defined by connection to a single central node; like a snowball sample that doesn't expand past the first set of connections

**From lecture 1b: [Ties and Adjacency Matrices](https://www.loom.com/share/a64c3a975e7147b9bb06bf921325bd96)**

adjacency matrix
: an $n \times n$ matrix depicting connections between $n$ nodes as 0 or 1, where 1 is a connection (or vertex/edge) and 0 is the absence of one:

||A|B|C|D|
|---|---|---|---|---|
|A|-|0|1|0|
|B|0|-|1|0|
|C|1|1|-|0|
|D|0|0|0|-|

directed tie
: a relationship where $W_{i,j} \ne W_{j,i}$, as might be used to represent a transfer of resources from one node to another; graphed with arrows

symmetric tie
: a relationship where $W_{i,j} = W_{j,i}$, with no direction; graphed with lines

binary tie
: a tie where $W_{i,j}$ is 0 or 1, as in the example above, indicating the absence or presence of a tie (also *dichotomous*)

valued tie
: a tie where $W_{i,j}$ is a value $v$, indicating a weight or magnitude of the connection; may be graphed with line attributes such as weight, color, etc

Ties are symmetric or directed, and binary or valued.

**From lecture 1c: [Edgelists](https://www.loom.com/share/9232d0ffe78d4ea4affc543bf4e353af)**

edgelist
: a table indicating edges in a network, with at least "from" and "to" columns, and possibly additional columns for attributes or values

|From|To|Value|
|---|---|---
|A|B|5|
|A|E|2|
|B|A|-1|
|B|C|4|
|B|D|2|
|C|A|-4|

Edgelists are more efficient in sparse networks, as they only list actual connections rather than being a matrix of all possible connections

### 2022-01-30 _academics/ps753/ps753.Rmd
PS753 glossary stuff

@@
 description: |
-  Front page for PS753 coursework, Spring 2022.
-author:
-  - name: Steve Linberg
-date: 2022-01-29
+  Political and Social Networks, Spring 2022.
+# author:
+#   - name: Steve Linberg
+# date: 2022-01-29
 output:
@@
     self_contained: false
+citation: false
 ---
@@
 
-# PS753: Political and Social Networks, Spring 2022
+<p style="background: #eee">
+PS 753
+</p>
 
-From this index page, we will magically link to the glossary page.
+Class work from this class will be listed here as we go. This part has to be done by hand in Distill, unfortunately.
+
+- My ongoing [glossary](../PS753-glossary/index.html) page; feel free to comment, or make additions or corrections via comments or github.

### 2022-01-30 _academics/ps753/ps753.Rmd
lowercase ps743 you idiot
text

@@
 
-Class work from this class will be listed here as we go. This part has to be done by hand in Distill, unfortunately.
+Work and assignments from this class will be added here as we go. (This part has to be done by hand in Distill, unfortunately.)
 
-- My ongoing [glossary](../PS753-glossary/index.html) page; feel free to comment, or make additions or corrections via comments or github.
+- My ongoing [glossary](../ps753-glossary/index.html) page; feel free to comment, or make additions or corrections via comments or github.

### 2022-01-30 _posts/2022-01-29-ps753-glossary-underway/ps753-glossary-underway.Rmd
Add blog post

---
title: "PS753 Glossary underway"
description: |
  Starting a section I've been planning for a while
author:
  - name: Steve Linberg
    url: https://slinberg.net
date: 2022-01-29
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

One of the things I want to have in the `Academics` section of the site is glossaries of terms from some of my classes, with the hope that they will be useful to others (as well as myself), and to also make it possible for people to add/correct/amend them. The latter is now possible via GitHub, where the complete source for this site can be found (hit the GitHub icon in the menu in the top right).

The [glossary](../../academics/ps753-glossary/) for [PS753, Political and Social Networks](../../academics/ps753), is underway with terms from week 1.

### 2022-01-30 _academics/ps753-glossary/ps753-glossary.Rmd
lowercase ps743 you idiot
Add pseudo top-level nav for /academics
typo

@@
 <p style="background: #eee">
-[PS 753](../PS753/index.html)
+Academics &gt; [PS 753](../ps753/index.html)
 &gt; Glossary
@@
 
-and $j$ similarly as a set of $n$ elements, except that $j$ cannot equal $i (a node cannot connect to itself):
+and $j$ similarly as a set of $n$ elements, except that $j$ cannot equal $i$ (a node cannot connect to itself):
 $$j \in 1, 2, 3 \dots n, i \ne j$$

### 2022-01-30 _academics/ps753-glossary/ps753-glossary.Rmd
Test of bibtex citation

@@
    self_contained: false
+bibliography: PS753.bib
 ---
@@
 
+(Note: some terms overlap somewhat or have context-dependent synonyms.)
+
 node / vertex
-: a junction in a network where two or more lines (edges) intersect; a dot connecting lines.<br>
+: a junction in a network where two or more lines (edges) intersect; a dot connecting lines.
 
-edge / link
+tie / edge / link
 : lines connecting nodes, which indicate some sort of connection or relationship.
 
-node or ego
+node / ego / actor
 : $i$ - the node being discussed or focused on
@@
 snowball sample
-: an entity group formed by starting with one node and "rolling outwards" to its connections until all nodes or actors (in a limited set) are located
+: an entity group formed by starting with "a focal actor or set of actors"^[@hanneman] and "rolling outwards" to its connections until all nodes or actors (in a limited set) are located
 
@@
 Edgelists are more efficient in sparse networks, as they only list actual connections rather than being a matrix of all possible connections
+
+---
+
+https://retorque.re/zotero-better-bibtex/

### 2022-01-30 _academics/ps753-glossary/ps753-glossary.Rmd
text updates

@@
 description: |
-  Glossary page for PS753 coursework
+  Glossary for PS753 coursework
 author:
@@
    self_contained: false
+   toc: true
 bibliography: PS753.bib
@@
 
-# Week 1
+> **A social network** is a set of actors (or points, or nodes, or agents) that may have relationships (or edges, or ties) with one another.
+Networks can have few or many actors, and one or more kinds of relations between pairs of actors.
+>
+>-- <cite>[@hanneman, chapter 2]</cite>
 
-**From lecture 1a: [Nodes, Edges and Network Samples](https://www.loom.com/share/efc5252bb75b463481b93e55be6e4d67)**
+## Week 1
+
+### Class lecture 1a: [Nodes, Edges and Network Samples](https://www.loom.com/share/efc5252bb75b463481b93e55be6e4d67)
 
@@
 
-tie / edge / link
+tie / edge / link / relation
 : lines connecting nodes, which indicate some sort of connection or relationship.
@@
 node / ego / actor
-: $i$ - the node being discussed or focused on
+: $i$ - the node being discussed or focused on; @hanneman also note "Actors are described by their relations, not by their attributes."
 
@@
 snowball sample
-: an entity group formed by starting with "a focal actor or set of actors"^[@hanneman] and "rolling outwards" to its connections until all nodes or actors (in a limited set) are located
+: an entity group formed by starting with "a focal actor or set of actors"^[@hanneman] and "rolling outwards" to its connections until all nodes or actors (in a limited set) are located, or a decision to stop is made
 
@@
 
-**From lecture 1b: [Ties and Adjacency Matrices](https://www.loom.com/share/a64c3a975e7147b9bb06bf921325bd96)**
+### Class lecture 1b: [Ties and Adjacency Matrices](https://www.loom.com/share/a64c3a975e7147b9bb06bf921325bd96)
 
@@
 
-**From lecture 1c: [Edgelists](https://www.loom.com/share/9232d0ffe78d4ea4affc543bf4e353af)**
+### Class lecture 1c: [Edgelists](https://www.loom.com/share/9232d0ffe78d4ea4affc543bf4e353af)
 
@@
 
----
+### @hanneman, chapter 1 (Social network data):
 
-https://retorque.re/zotero-better-bibtex/
+binary measures of relations
+: undirected relations, 0 or 1 for the absence or presence of a connection
+
+multiple-category nominal measures of relations
+: directed relations with categories (like "friend, lover, business relationship, kin, or no relationship")
+
+grouped ordinal measures of relations
+: measures that reflect a level of intensity or degree; often turned into binary measures by means of a threshold or cut-off
+
+full-rank ordinal measures of relations
+: an ordering from 1 to $n$ of an actor's relations (uncommon in social networks)
+
+interval measures of relations
+: continuous measures that express the strength of connections by comparison with others, to be able to say "this tie is twice as strong as that tie"; the "most advanced" level of measurement

### 2022-01-30 _academics/ps753-glossary/ps753-glossary.Rmd
Add Lazer terms

@@
 node / ego / actor
-: $i$ - the node being discussed or focused on; @hanneman also note "Actors are described by their relations, not by their attributes."
+: $i$ - the node being discussed or focused on; also note "Actors are described by their relations, not by their attributes."^[@hanneman]
 
@@
 : continuous measures that express the strength of connections by comparison with others, to be able to say "this tie is twice as strong as that tie"; the "most advanced" level of measurement
+
+### @lazer2011
+
+homophily
+: the idea that individuals who are similar to one another are more likely to form ties
+
+whole network data
+: relational information about a whole set (or subset) of data, with all of the actors' relations to each other considered
+
+egocentric data
+: relational information about a set of nodes connected to one particular node and not to each other
+
+diameter
+: the maximum degree of separation between any two nodes in the network
+
+one-mode data
+: ties among one set (or category) of agent, such as nations in the context of trade
+
+two-mode data
+: ties between different sets (or categories) of agents, such as ties between nations and international organizations

### 2022-01-31 _academics/ps753-glossary/ps753-glossary.Rmd
Add Borgatti terms

@@
 : ties between different sets (or categories) of agents, such as ties between nations and international organizations
+
+### @borgatti2009
+
+sociometry
+: "a technique for eliciting and graphically representing individuals' subjective feelings toward one another"
+
+strength of weak ties (SWT) theory
+: the theory that one is likelier to hear new information from people they aren't closely connected to in a network (c.f. homiphily)
+
+centrality
+: a family of positional properties of nodes in a network
+
+Freeman's betweenness
+: a type of centrality where a node is frequently along the shortest path between pairs of nodes, giving control over flow or power
+
+opportunity-based antecedents
+: "the likelihood that two nodes will come into contact" - when considering the formation of ties
+
+benefit-based antecedents
+: "some kind of utility maximization or discomfort minimization that leads to tie formation"
+
+node homogeneity
+: a category of node outcomes referring to the similarity of nodes
+
+performance
+: a category of node outcomes referring to some good (such as strong performance)

### 2022-01-31 _academics/ms797-glossary/ms797-glossary.Rmd
Add ms797 index and glossary

---
title: "MS797 Glossary"
description: |
  Glossary for MS797 coursework
author:
  - name: Steve Linberg
    url: {}
date: 2022-01-30
output:
 distill::distill_article:
   self_contained: false
   toc: true
# bibliography: PS753.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

<p style="background: #eee">
Academics &gt; [MS 797](../ms797/index.html)
&gt; Glossary
</p>

Page numbers in parenthesis after terms, from [ISLR](https://www.statlearning.com/) 2nd edition.

## Chapter 2

input variable (15)
: also *predictor*, *independent variable*, *feature*; usually written $X_1, X_2$, etc. The parameter or parameters we are testing to see if they are related to or affect the output.

output variable (15)
: also *response*, *dependent variable*; usually written $Y$. The outcome being measured.

error term (16)
: $\epsilon$ in the equation  
$$Y = f(X) + \epsilon$$
a random quantity of inaccuracy, *independent of X* and with *mean 0*.

systematic (16)
: $f$ in the equation  
$$Y = f(X) + \epsilon$$
the function that describes the (systematic) information $X$ provides about $Y$. This plus the error term equals $Y$.

reducible error (18)
: The amount of the error $\epsilon$ that could be eliminated by improving our estimator $\hat{f}$; the difference between $\hat{f}$ and $f$. This book and course is mostly about ways to minimize the reducible error.

irreducible error (18)
: The amount of $\epsilon$ that could not be reduced even if $f$ was a perfect estimator of $Y$. Always greater than 0. Could be due to hidden variables in $\epsilon$, or random fluctuations in Y, like a measure of "[a] patient's general feeling of well-being on that day".

expected value (19)
: *average value* of an expected measure.

training data (21)
: data used to develop the model for estimating $f$.

parametric methods (21)
: A model based on one or more input parameters, that yields a value for Y, as in:
$$f(X) = \beta_0 + \beta_1X_1 + \beta_2X_2 + \dots + \beta_pX_p$$
$$Y \approx \beta_0 + \beta_1X_1 + \beta_2X_2 + \dots + \beta_pX_p$$
$$\text{income} \approx \beta_0 + \beta_1 \times \text{education} + \beta_2 \times \text{seniority}$$
This creates a predictive, *inflexible* model which usually does not match the true $f$, but which has advantages of simplicity and interpretability. It can be used to predict values for $Y$ based on its parameters, or inputs. Linear and logistic regression are parametric.

non-parametric methods (23)
: methods that do not attempt to estimate $f$. More flexible and have the potential to very closely match observations, but with the risk of *overfitting* the data and increasing the *variance* of subsequent observations. They require much more data than parametric models, and may be difficult to interpret, K-Nearest Neighbor and Support Vector Machines are non-parametric.

prediction (26)
: seeking to guess the value of an response variable $y_i$ given a set of observations and a predictor $f$.

inference (26)
: a model that seems to better understand the relationship between the response and the predictors.

supervised learning (26)
: a category of model that allows us to guess a $y_i$ response to a set of predictor measurements $x_i, i = 1, \dots, n$.

unsupervised learning (26)
: a category of model in which there are observations/measurements $x_i, i = 1, \dots, n$, but no associated response $y_i$. Linear regression cannot be used because there is no response variable to predict.

cluster analysis (27)
: in unsupervised learning, a statistical method for determining whether a set of observations can be divided into "relatively distinct groups," looking for similarities within the groups. (Topic modeling may be an example of this.)

quantitative variables (28)
: numeric values; age, height, weight, quantity. Usually the response variable type for regression problems.

qualitative variables (28)
: also *categorical*: values from a discrete set. Eye color, name, yes/no. Usually the response variable type for classification problems.

regression problems (28)
: problems with quantitative response variables. Given predictors foo, bar, and baz, how big is the frob?

classification problems (28)
: problems with qualitative response variables. Given predictors foo, bar, and baz, is the outcome likely to be a frob, a frib or a freeb?

mean squared error (MSE) (29)
: the average squared error for a set of observations:
$$MSE = \frac{1}{n}\sum_{i=1}^{n}(y_i - \hat{f}(x_i))^2$$
MSE is small if the predicted responses are close to the true responses, and larger as it becomes less accurate; computed from training data, and Gareth *et al.* suggest it should be called *training MSE*.

variance (34)
: _"the amount by which $\hat{f}$ would change if we estimated it using a different training data set"_

bias (35)
: _"the error that is introduced by approximating a real-life problem, which may be extremely complicated, by a much simpler model"_, as in the error from the (presumed) linearity of a regression against non-linear data whose complexity it does not capture. More flexible models increase variance and decrease bias.

bias-variance trade-off (36)
: The tension in seeking the best model for the data between missing the true $f$ with an overly simple (biased) model, vs. an overfitted model with too much variance from mapping too closely to test data.

error rate (37)
: In classification, the proportion of classifications that are mistakes.
$$\frac{1}{n}\sum_{i=1}^{n}I(y_i \neq \hat{y}_i)$$
$I$ is 1 if $y_i \neq \hat{y}_i$ - if the guess for any given $y$ is wrong. The error rate is the percentage of incorrect classifications. Also the *training erorr rate*.

indicator variable (37)
: $I$ in the error rate definition above; a logical variable indicating the presence or absence of a characteristic or trait (such as an accurate classification).

test error rate (37)
: like the *training error rate* but applied to the test data. Uses $\text{Ave}$ instead of sum notation:
$$\text{Ave}(I(y_0 \neq \hat{y}_0))$$
$\hat{y}_0$ is the predicted class label from the classifier.

conditional probability (37)
: The chance that $Y = j$ given an observed $x_0$, as in the *Bayes classifier*:
$$\text{Pr}(Y = j|X = x_0)$$
In a two-class, yes/no classifier, we decide based on whether $\text{Pr}(Y = j|X = x_0)$ is $> 0.5$, or not. Note that $Y$ is the class, as in "ham"/"spam", not a $y$-axis coordinate.

Bayes decision boundary (38)
: a visual depiction of the line of 50% probability dividing (exactly two?) classes in a two-dimensional space

Bayes error rate (38)
: the expected (average) probability of classification error over all values of X in a data set.
$$1 - E(\underset{j}{maxPr}(Y = j|X))$$
The $\underset{j}{maxPr}$ whichever of the $j$ classes has the highest probability for any given value of $X$. Again, $Y$ is not a y-axis coordinate of a two-dimensional space, it's the *class* of the classification: "yes"/"no", "ham"/"spam", "infected"/"not infected". Also:
_"The Bayes error rate is analogous to the irreducible error, discussed earlier."_

K-nearest-neighbors (KNN) (39)
: a classifier that assigns a class Y to an observation based on the population proportions of its nearest neighbors; a circular "neighborhood" on a two-dimensional plot. It looks at actual data points that have been classified, and asks what any given non-classified point would be classified as based on its nearest neighbors.

## Chapter 3

Synergy effect / interaction effect (60)
: when two or more predictors affect each other as well as the outcome; when 50k each in TV or radio ads give different results than 100k in either one

Simple linear regression
: the simplest model, predicting $Y$ from a single predictor $X$.
$$Y \approx \beta_0 + \beta_1X$$
$\approx$ = "is approximately modeled as"

least squares (61)
: the most common measure of closeness of a regression line to its data points, the sum of squares of the distances between the points and the closest point on the line (directly above or below)

residual (61)
: the difference between $y_i$ and $\hat{y}_i$, also $e_i$; the difference between the $i$th response variable and the $i$th response variable predicted by the model

residual sum of squares (RSS) (62)
: the sum of the squared residuals for each point on the regression line
$$\text{RSS} = e_1^2 + e_2^2 + \dots + e_n^2$$
Formulas for $\hat{\beta_0}$ and $\hat{\beta_1}$ are on p. 62

intercept ($\beta_0$) (63)
: the expected value of $Y$ when $X = 0$

slope ($\beta_1$) (63)
: the average increase in $Y$ associated with a one-unit increase in $X$

error term ($\epsilon$) (63)
: whatever we missed with the model, due to the true model not being linear (it almost never is), measurement error, or other variables that cause variation in $Y$

population regression line (63)
: "the best linear approximation to the true relationship between $X$ and $Y$"
$$Y = \beta_0 + \beta_1X + \epsilon$$
least squares line (63)
: the regression line made of the least-squares estimates for $\beta_0$ and $\beta_1$

bias (65)
: in an estimator, something that systematically misses the true parameter; for an unbiased estimator, $\hat{\mu} = \mu$ when averaged over (huge) numbers of observations

standard error (SE) (65)
: the average amount that an estimate $\hat{\mu}$ (sample mean) differs from the actual value of $\mu$ (population mean)
$$\text{Var}(\hat{\mu}) = \text{SE}(\hat{\mu})^2 = \frac{\sigma^2}{n} (\text{also} = \frac{\sigma}{\sqrt{n}})$$
$\sigma$ is "the standard deviation of each of the realizations $y_i$ of $Y$. Since $\sigma^2$ is divided by $n$, the standard error shrinks as observations increase. It represents the amount we would expect means of additional samples to "jump around" simply due to random chance and the limitations of the model's accuracy.^[http://faculty.ucr.edu/~hanneman/nettext/C1_Social_Network_Data.html]

residual standard error (RSE) (66)
: the estimate of $\sigma$
$$\text{RSE} = \sqrt{RSS / (n-2)}$$

confidence interval (66)
: a range of values within which we have a measured probability (often 95%) of containing the true value of the parameter; a 95% confidence interval in linear regression takes the form
$$\hat{\beta_1} \pm 2 \cdot \text{SE}(\hat{\beta_1})$$

t-statistic (67)
: the number of standard deviations that $\hat{\beta_1}$ is away from $0$.
$$t = \frac{{\hat{\beta_1}} - 0}{\text{SE}(\hat{\beta_1})}$$
For there to be a relationship between $X$ and $Y$, $\hat{\beta_1}$ has to be nonzero (i.e. have a slope). The standard error (SE) of $\hat{\beta_1}$ (in the denominator above) measures its accuracy; if it is small, then $t$ will be larger, and if it is large, then $t$ will be smaller. $t$ is around 2 for a p-value of 0.05 (actually about 1.96, as 2 standard deviations is 95.45% of a normal distribution), and around 2.75 for a p-value of 0.01.

p-value (67)
: the probability of observing a value greater than $|t|$ by chance.

total sum of squares (TSS) (70)
: the total variance in the response $Y$
$$\text{TSS} = \sum(y_i - \bar{y})^2$$
compare with RSS, the amount of variability left unexplained after the regression. TSS - RSS is the amount of variability (or error) explained by the regression.

$R^2$ statistic (70)
: the proportion of variance in $Y$ explained by $X$, a range from 0 to 1
$$R^2 = \frac{\text{TSS - RSS}}{\text{TSS}} = 1 - \frac{\text{RSS}}{\text{TSS}}$$
$R^2$ values close to 1 indicate a regression that explains a lot of the variability in the response, and a stronger model. A value close to 0 indicates that the regression doesn't explain much of the variability.

correlation (70)
: a measure of the linearity of the relationship between $X$ and $Y$; values close to 0 indicate weak-to-no relationship, values near 1 or -1 indicate strong positive or negative correlation
(horrible formula on p.70)

### 2022-01-31 _academics/ms797/ms797.Rmd
Add ms797 index and glossary

---
title: "MS797"
description: |
  Machine Learning for Social Sciences, Spring 2022.
# author:
#   - name: Steve Linberg
# date: 2022-01-30
output:
  distill::distill_article:
    self_contained: false
citation: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

<p style="background: #eee">
Academics &gt; ML797
</p>

Work and assignments from this class will be added here as we go. (This part has to be done by hand in Distill, unfortunately.)

- My ongoing [glossary](../ml797-glossary/index.html) page; feel free to comment, or make additions or corrections via comments or github.

### 2022-01-31 _posts/2022-01-30-ms797-glossary-underway/ms797-glossary-underway.Rmd
New post

---
title: "MS797 Glossary underway"
description: |
  The other big glossary for this semester is underway.
author:
  - name: Steve Linberg
    url: https://slinberg.net
date: 2022-01-30
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

The term's other major [glossary](../../academics/ms797-glossary/), for [MS797, Machine Learning for Social Science](../../academics/ms797), is up. At this point it's about halfway through chapter 3... lots to go. This was initially in a standalone repo, but I think it makes more sense to migrate it here.

Again, comments, clarifications or corrections are all welcome.

### 2022-01-31 _academics/ms797/ms797.Rmd
correct path

@@
 <p style="background: #eee">
-Academics &gt; ML797
+Academics &gt; MS797
 </p>
@@
 
-- My ongoing [glossary](../ml797-glossary/index.html) page; feel free to comment, or make additions or corrections via comments or github.
+- My ongoing [glossary](../ms797-glossary/index.html) page; feel free to comment, or make additions or corrections via comments or github.

### 2022-02-01 _academics/ps753-glossary/ps753-glossary.Rmd
Add week 2 terms

@@
 : a category of node outcomes referring to some good (such as strong performance)
+
+## Week 2
+
+### Class lecture 2a: [Network Structure - Walks, Paths, and Distance](https://www.loom.com/share/a8a7949244f04ed5995d572340c20aa0)
+
+Connections between nodes:
+
+adjacent
+: a direct connection between nodes; not necessarily bilateral in directed networks. A leading to B ("A adj B") does not mean that B leads to A ("B adj A)
+
+reachable
+: whether a node is reachable from another node, regardless of distance
+
+distance
+: the number of ties (steps, edges) that must be traversed to reach a target node
+
+walk
+: sequence (not *path*, see below) that connects two nodes, consisting of the nodes and edges
+
+trail
+: a walk that can only go through each edge / tie once, but can hit the same node more than once
+
+path
+: a trail that only hits each node and edge once; the start and end node may be the same
+
+geodesic distance
+: the shortest path between two nodes; by definition, not a trail or walk because repeated segments wouldn't be the shortest; with binary data, the number of steps; with weighted data, might be a sum or some other calculation of "effort"
+
+### Class lecture 2b: [Graph Substructures and Components](https://www.loom.com/share/515068ae6dc641be984c0ef974fb8e52)
+
+Network substructures
+
+dyad, triad, clique
+: two, three, or four-or-more connected nodes
+
+complete graph
+: a network where every node is directly connected to every other node
+
+connected graph
+: a network where every node is indirectly connected to every other node
+
+unconnected graph
+: a network where at least one node is unreachable
+
+component
+: the set of all points that constitutes a connected subgraph within a network
+
+main component
+: the largest component within a network
+
+minor component
+: a smaller one, possibly one of many
+
+pendant
+: a node with only one link or edge to a network, "dangling"
+
+isolate
+: an unconnected node
+
+### Class lecture 2c: [Dyad and Triad Census](https://www.loom.com/share/96aaf1d2982d427c83564c338d2e0b83)
+
+mutual dyad
+: a dyad where both nodes connect to each other, as in an undirected network
+
+asymmetric dyad
+: a dyad (in a directed network) where one node connects to another, but non-reciprocally (one way only)
+
+null
+: a dyad of two unconnected nodes
+
+(empty / one edge / two path / triangle) triad
+: a triad with zero, one, two or three edges between three nodes (all four possible permutations)
+
+balance theory
+: the theory that two nodes connected to a common node will also develop connections to each other
+
+global transitivity index
+: the proportion of triads in a network that are complete (with 3 connections between them)
+
+There is a vocabulary for triads in directed networks describing the 16 possible permutations of ties (or the absence thereof) among 3 nodes, counting the number of mutual, asymmetric and null dyads, with direction indicators, like `003` or `120D` - see the slide at 3:45
+
+vacuously transitive triad
+: a triad where (5 of 16 possibilities)
+
+intransitive triad
+: a triad where (7 of 16 possibilities)
+
+transitive triad
+: a triad where (4 of 16 possibilities)
+
+xxx see https://corescholar.libraries.wright.edu/cgi/viewcontent.cgi?article=2427&context=knoesis
+
+### Class lecture 2d: [Transitivity and Clustering Coefficient](https://www.loom.com/share/ff29331e900741dd8634d921e4abafdc)
+
+local transitivity / local clustering coefficient
+: the likelihood that the neighbors of a node are also connected to each other; the number of connections that *do* exist over the number of connections that *could* exist

### 2022-02-01 _academics/ps753-glossary/ps753-glossary.Rmd
build
update citation

@@
 
-xxx see https://corescholar.libraries.wright.edu/cgi/viewcontent.cgi?article=2427&context=knoesis
+(See @alhazmiUnderstandingSocialEffects2015)
 
@@
 : the likelihood that the neighbors of a node are also connected to each other; the number of connections that *do* exist over the number of connections that *could* exist
+
+![homer](homer.png)
+In the example above, there are 8 nodes that Homer (center) could connect to; among those 8 nodes, there are 7 + 6 + 5 + 4 + 3 + 2 + 1 = 28 possible undirected connections (not connecting to Homer), and 9 of those 28 do exist, for a local clustering coefficient of 9/28 = 0.32.

### 2022-02-01 _academics/ps753-glossary/ps753-glossary.Rmd
update geodesic distance definition

@@
 geodesic distance
-: the shortest path between two nodes; by definition, not a trail or walk because repeated segments wouldn't be the shortest; with binary data, the number of steps; with weighted data, might be a sum or some other calculation of "effort"
+: the shortest path between two nodes; by definition, not a trail or walk because repeated segments wouldn't be the shortest; with binary data, the number of edges between the nodes; with weighted data, might be a sum or some other calculation of "effort"
 

### 2022-02-01 _academics/ps753-glossary/ps753-glossary.Rmd
Text tweaks

@@
 vacuously transitive triad
-: a triad where (5 of 16 possibilities)
+: a triad where (...) (5 of 16 possibilities)
 
 intransitive triad
-: a triad where (7 of 16 possibilities)
+: a triad where (...) (7 of 16 possibilities)
 
 transitive triad
-: a triad where (4 of 16 possibilities)
+: a triad where (...) (4 of 16 possibilities)
 

### 2022-02-01 _meta/distill-and-github-pages-setup/distill-and-github-pages-setup.Rmd
Text tweaks, build
Text tweaks

@@
 If you're on Mac or Linux, you probably have [git](https://git-scm.com/) built in already, but you might need to install it if you're on Windows.
-That's outside the scope of this article, but you can find information about it on the main Git page.
+That's outside the scope of this article, but you can find information about it on the [Git for Windows download page](https://git-scm.com/download/win).
+
+The very high-level overview of the process here is:
+
+1. Create a new, empty GitHub repository that will contain your site
+2. Create a new, empty RStudio project using the repository you just created
+3. Install the Distill framework into this project
+4. Configure and build the bare-bones, empty site and push it up to GitHub
+5. Configure GitHub Pages to publish the site
+
+Once you've done that, you can begin writing content and publishing blogs, and spend time fine-tuning and tweaking your site.
+(You can also do that before setting up GitHub Pages if you want to wait to unveil it until it has more content and polish - as you prefer.)
 
@@
  create mode 100644 distill-blog.Rproj
-[slinbergdistill-blog] (main)$ git push Enter passphrase for key '/Users/slinberg/.ssh/sfl_id_rsa': 
+[slinbergdistill-blog] (main)$ git push 
+Enter passphrase for key '/Users/slinberg/.ssh/sfl_id_rsa': 
 Enumerating objects: 4, done.
@@
 
-Then load the library, which you should also do any time you want to work on your site in the future after opening the project:
+Then load the library, which you should also do as your first step (after opening the project) any time you work on it in the future:
 
@@
 You could just go from here, if you're content with the URL GitHub chose for you. Your basic workflow will be:
+
 - Work on content in RStudio

### 2022-02-01 _academics/ps753-glossary/ps753-glossary.Rmd
text

--- 2022-02-01 _academics/ps753-glossary/ps753-glossary.Rmd
@@
 vacuously transitive triad
-: a triad where (5 of 16 possibilities)
+: a triad where (to be continued...) (5 of 16 possibilities)
 
 intransitive triad
-: a triad where (7 of 16 possibilities)
+: a triad where (to be continued...) (7 of 16 possibilities)
 
 transitive triad
-: a triad where (4 of 16 possibilities)
+: a triad where (to be continued...) (4 of 16 possibilities)
 

### 2022-02-01 _academics/ps753-glossary/ps753-glossary.Rmd
add bit o' latex

@@
 ![homer](homer.png)
-In the example above, there are 8 nodes that Homer (center) could connect to; among those 8 nodes, there are 7 + 6 + 5 + 4 + 3 + 2 + 1 = 28 possible undirected connections (not connecting to Homer), and 9 of those 28 do exist, for a local clustering coefficient of 9/28 = 0.32.
+In the example above, there are 8 nodes that Homer (center) could connect to; among those 8 nodes, there are $7 + 6 + 5 + 4 + 3 + 2 + 1 = 28$ possible undirected connections (not connecting to Homer), and 9 of those 28 do exist, for a local clustering coefficient of $9/28 \approx 0.32$.

### 2022-02-02 _posts/2022-02-02-pythagorean-theorem/pythagorean-theorem.Rmd
new post

---
title: "Pythagorean Theorem"
description: |
  How did I not know this
author:
  - name: Steve Linberg
    url: https://slinberg.net
date: 2022-02-02
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

How did I make it to `$adulthood * 2` without knowing that our old friend, the Pythagorean Theorem ($a^2 + b^2 = c^2$), works in any number of dimensions?

The Euclidean distance $d$ between the origin ($0, 0, \dots 0$) is just

$$d = \sqrt{x_1^2 + x_2^2 + \dots + x_n^2}$$

Add up all of the squares and take the square root. Dope-smack for me.

### 2022-02-02 _posts/2022-02-02-pythagorean-theorem/pythagorean-theorem.Rmd
add category

@@
     self_contained: false
+categories:
+  - math
 ---

### 2022-02-05 _academics/ms797-glossary/ms797-glossary.Rmd
typo
Many additions

@@
    toc: true
-# bibliography: PS753.bib
 ---
@@
 <p style="background: #eee">
-Academics &gt; [MS 797](../ms797/index.html)
+Academics &gt; [MS797](../ms797/index.html)
 &gt; Glossary
@@
 Page numbers in parenthesis after terms, from [ISLR](https://www.statlearning.com/) 2nd edition.
+Non-page numbers indicate other sources; "biostats" references material from Biostatistics 690Z (Health Data Science: Statistical Modeling), fall 2021.
 
@@
 
+model sum of squares (MSS) (biostats)
+: Also sometimes **ESS**, "explained sum of squares": the total variance in the response $Y$ that can be accounted for by the model
+$$\text{MSS} = \sum(\hat{y_i} - \bar{y})^2$$
+
+residual sum of squares (RSS) (biostats)
+: the total variance in the response $Y$ that cannot be accounted for by the model
+$$\text{RSS} = \sum(y_i - \hat{y_i})^2$$
+also
+$$\text{RSS} = e_i^2 + e_2^2 + \dots + e_n^2$$
+or
+$$\text{RSS} = (y_1 - \hat{\beta_0} - {\hat{\beta_1}x_1})^2 + (y_2 - \hat{\beta_0} - {\hat{\beta_1}x_2})^ + \dots + (y_n - \hat{\beta_0} - {\hat{\beta_1}x_n})^2$$
+
 total sum of squares (TSS) (70)
-: the total variance in the response $Y$
+: the total variance in the response $Y$; the total variability of the response about its mean
 $$\text{TSS} = \sum(y_i - \bar{y})^2$$
-compare with RSS, the amount of variability left unexplained after the regression. TSS - RSS is the amount of variability (or error) explained by the regression.
+compare with RSS, the amount of variability left unexplained after the regression. TSS - RSS is the amount of variability (or error) explained by the regression (MSS).
+
+**NOTE**: there is a nice visual [here on stackexchange](https://stats.stackexchange.com/a/512216/338749); if anybody knows how to tell Zotero to use a custom bibtex citation entry over the ones it generates, please let me know so I can integrate it better here :frown:
 
@@
 : a measure of the linearity of the relationship between $X$ and $Y$; values close to 0 indicate weak-to-no relationship, values near 1 or -1 indicate strong positive or negative correlation
-(horrible formula on p.70)
+(horrible formula on p.70 I don't have time to LaTeX at the moment)
+
+standard linear regression model (72)
+: The modal used for standard linear models, used to interpret the the effect on $Y$ of a one-unit increase in any predictor $\beta_j$ while holding all other predictors constant
+$$Y = \beta_0 + \beta_1X_1 + \beta_2X_2 + \dots + \beta_pX_p + \epsilon$$
+
+variable selection (78)
+: the task of refining a model to include only the variables associated with the response
+
+null model (79)
+: a model that conatins an intercept, but no predictors; used as a first stage in forward selection
+
+forward selection (79)
+: a variable selection method that starts with a null model, and then runs simple linear regressions on all predictors $p$ and adding the one that results in the lowest RSS; repeated until some threshold is reached
+
+backwards selection (79)
+: a variable selection method that starts with a model with all predictors, and removing the one with the lowest $p$-value until all remaining predictors are significant, whether by $p$-value or some other criterion
+
+mixed selection (79)
+: a hybrid approach starting with a null model, adding predictors one at a time that produce the best fit, and removing any that acquire a larger $p$-value in the process until all predictors are added or eliminated
+
+interaction (81)
+: when predictors affect each other, in addition to providing their own effect on the model
+
+confidence interval (82)
+: a range with a percentage component in which there is that percentage chance that the true value of an estimated parameter lies; a 95% confidence interval is a range in which we can be 95% certain $f(X)$ will be found
+
+prediction interval (82)
+: similar to confidence interval, but a prediction range within which we are $X$% certain that any singular future observation will fall, rather than a statistic like an overall mean; a 95% prediction interval is a range in which we are confident that 95% of future observations will fall.
+Prediction ranges are substantially wider than confidence intervals.
+
+qualitative predictor / factor (83)
+: a categorical predictor with a fixed number of factors, like "yes" / "no" or "red" / "yellow" / "green"
+
+dummy variable (83)
+: a numeric representation of a factor to use in a model, as in representing "yes" / "no" factor variables as 1 / 0 in a regression
+
+baseline (86)
+: the factor level where there is no dummy variable; a factor with 3 levels will use 2 dummy variables, with the factor's absence signifying the 3rd value (usually 0)
+
+additivity assumption (87)
+: the assumption that the association between a predictor $X$ and the response $Y$ does not depend on the value of other predictors; used by the standard linear regression model
+
+linearity assumption (87)
+: the assumption, also used by the standard linear regression model, that unit changes in $X_j$ result in the same change to Y regardless of its value
+
+interaction term (88)
+: the product of two predictors in a multiple regression model, quantifying their effect on each other
+
+main effect (89)
+: isolated effects; the effect of a single predictor on the outcome
+
+hierarchical principle (89)
+: the principle that main effects should be left in a model even if they are statistically insignificant, if they are also part of an interaction that is significant
+
+polynomial regression (91)
+: an extension of linear regression to accommodate non-linear relationships
+
+residual plot (93)
+: a plot of the residuals or errors ($e_i = y_i - \hat{y}_i$), used to check for non-linearity (a potential problem that would likely indicate something was missed in the model)
+
+time series (94)
+: data consisting of observation made at discrete points in time
+
+tracking (95)
+: when (residuals / variables?) tend to have similar values
+
+heteroscedasticity (96)
+: non-constant variances in errors; "unequal scatter"
+
+homoscedasticity (extra)
+: constant variances in errors; follows the assumption of equal variance required by most methods
+
+weighted least squares (97)
+: an extension to ordinary least squares used in circumstances of heteroscedasticity, to weight data points proportionally with the inverse variances
+
+outlier (97)
+: an observation whose value is very far from its predicted value
+
+studentized residual (98)
+: a residual divided by its estimated standard error; observations with student errors higher than 3 (indicating 3 standard deviations) are likely outliers
+
+high leverage (98)
+: observations with an unusual $x_i$ value, far from other / expected $x$ values
+
+leverage statistic (99)
+: a quantification of a point's leverage
+$$h_i = \frac{1}{n} + \frac{(x_i - \bar{x})^2}{\sum_{i'=1}^{n}(x_{i`} - \bar{x})^2}$$
+
+collinearity (99)
+: when two or more predictor variables are closely related to each other
+
+power (101)
+: the probability of a test correctly detecting a nonzero coefficient (and correctly rejecting $H_0 : \beta_j = 0$)
+
+multicollinearity (102)
+: when collinearity exists between three or more predictors even when no pair of predictors is collinear (or correlated)
+
+variance inflation factor (102)
+: "the ratio of the variance of $\hat{\beta_j}$ when fitting the full model divided by the variance of $\hat{\beta_j}$ on its own"; smallest possible value of 1 indicates the absence of collinearity, 5-10 indicates a "problematic amount".
+Hairy forumla not reproduced.

### 2022-02-05 _academics/ms797-glossary/ms797-glossary.Rmd
correct term

@@
 studentized residual (98)
-: a residual divided by its estimated standard error; observations with student errors higher than 3 (indicating 3 standard deviations) are likely outliers
+: a residual divided by its estimated standard error; observations with student residuals higher than 3 (indicating 3 standard deviations) are likely outliers
 

### 2022-02-07 _academics/ms797-useful-r-tidbits/ms797-useful-r-tidbits.Rmd
build

---
title: "Useful R tidbits"
description: |
  Tasty R snacks that do useful things
author:
  - name: Steve Linberg
    url: {}
date: 2022-02-06
output:
  distill::distill_article:
    self_contained: false
    toc: true
preview: ISLR-cover.png
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = FALSE)
```

<p style="background: #eee">
Academics &gt; [MS797](../ms797/index.html)
&gt; Useful R tidbits
</p>

Most of this comes directly from the [resources from the 2nd edition](https://www.statlearning.com/resources-second-edition) of [ISLR](https://www.statlearning.com/), either direct quotes or things that build on them. The `Lab` section of each chapter contains useful walkthroughs and code illustrating the chapter's key concepts, and there's always something to learn about R and how different people use it.

<aside>
[![ISLRv2](ISLR-cover.png){width=150px}](https://www.statlearning.com)
</aside>

Everything here is, unless otherwise marked, either the work of, or derivative of the work of, the original authors of the book.

## Chapter 3: Linear Regression

```{r chunk1, echo=FALSE}
library(MASS)
library(ISLR2)
```
### names(): see the attributes of a model

`names()` is a nice, simple way to see the attributes of a linear model:

```{r chunk4}
attach(Boston)
lm.fit <- lm(medv ~ lstat)
names(lm.fit)
```

### coef(): see the model's coefficients

`coef()` is a convenient way to get the `coefficients` attribute:

```{r chunk6}
coef(lm.fit)
```

Note: this is equivalent to `lm.fit$coefficients`, but slightly easier on the eyes.

### confint(): see confidence intervals for a model

```{r chunk7}
confint(lm.fit)
```

Note: levels can be added with the `level` parameter (the default is 0.95):

```{r}
confint(lm.fit, level = 0.90)
```

### predict(): see confidence OR prediction intervals for a model

```{r chunk8}
predict(lm.fit, data.frame(lstat = (c(5, 10, 15))),
    interval = "confidence")
predict(lm.fit, data.frame(lstat = (c(5, 10, 15))),
    interval = "prediction")
```

Note: prediction intervals are defined on p. 82; in brief, it is the interval in which we are $X$% certain that any future observation will fall.

Note: the `data.frame()` call in the above parameter lists produce a nifty little data frame that is used by `predict` to identify the column to do calculations on, and the desired levels:

```{r}
data.frame(lstat = (c(5, 10, 15)))
```

Note: the book uses the base R graphics:

```{r chunk9}
plot(lstat, medv)
abline(lm.fit)
```

The tidyverse/ggplot version of this is:

```{r chunk9a}
library(tidyverse)
ggplot(Boston, mapping = aes(x = lstat, y = medv)) +
  geom_point() +
  geom_abline(intercept = coef(lm.fit)[1],
              slope = coef(lm.fit)[2])
```

...a bit more work, but more modern.

### arranging plots in grids

The book uses base R, again, with `par()` and `mfrow()` to arrange plots:

```{r chunk11}
par(mfrow = c(2, 2))
plot(lm.fit)
```

ggplot uses `facet()` to arrange plots from the same data set in grids.

Plots from different data sets need additional packages to be combined; one option is the `cowplot` library and the `plot_grid()` function.

XXX
<hr>

Alternatively, we can compute the residuals from a linear regression
fit using the `residuals()` function. The function
`rstudent()` will return the studentized residuals, and we
can use this function to plot the residuals against the fitted values.

```{r chunk12}
plot(predict(lm.fit), residuals(lm.fit))
plot(predict(lm.fit), rstudent(lm.fit))
```

On the basis of the residual plots, there is some evidence of non-linearity.
Leverage statistics can be computed for any number of predictors using the `hatvalues()` function.

```{r chunk13}
plot(hatvalues(lm.fit))
which.max(hatvalues(lm.fit))
```

The `which.max()` function identifies the index of the largest element of a vector. In this case, it tells us which observation has the largest leverage statistic.

## Multiple Linear Regression

In order to fit a multiple linear regression model using least squares, we again use the `lm()` function. The syntax {\R{lm(y $\sim$ x1 + x2 + x3)}} is used to fit a model with three predictors, `x1`, `x2`, and `x3`.
The `summary()` function now outputs the regression coefficients for all the predictors.

```{r chunk14}
lm.fit <- lm(medv ~ lstat + age, data = Boston)
summary(lm.fit)
```

The `Boston` data set contains 12 variables, and so it would be cumbersome to have to type all of these in order to perform a regression using all of the predictors.
Instead, we can use the following short-hand:

```{r chunk15}
lm.fit <- lm(medv ~ ., data = Boston)
summary(lm.fit)
```

We can access the individual components of a summary object by name
(type `?summary.lm` to see what is available). Hence
`summary(lm.fit)$r.sq` gives us the $R^2$, and
`summary(lm.fit)$sigma` gives us the RSE. The `vif()`
function, part of the `car` package, can be used to compute variance
inflation factors.   Most VIF's are
low to moderate for this data. The `car` package is not part of the base `R` installation so it must be downloaded the first time you use it via the `install.packages()` function in `R`.

```{r chunk16}
library(car)
vif(lm.fit)
```

What if we would like to perform a regression using all of the variables but one?  For example, in the above regression output,  `age` has a high $p$-value. So we may wish to run a regression excluding this predictor.
 The following syntax results in a regression using all predictors except `age`.

```{r chunk17}
lm.fit1 <- lm(medv ~ . - age, data = Boston)
summary(lm.fit1)
```

Alternatively, the `update()` function can be used.

```{r chunk18}
lm.fit1 <- update(lm.fit, ~ . - age)
```

## Interaction Terms

It is easy to include interaction terms in a linear model using the `lm()` function. The syntax `lstat:black` tells `R` to include an interaction term between `lstat` and `black`.
The syntax `lstat * age` simultaneously includes `lstat`, `age`, and the interaction term `lstat`$\times$`age` as predictors; it is a shorthand for `lstat + age + lstat:age`.
  %We can also pass in transformed versions of the predictors.

```{r chunk19}
summary(lm(medv ~ lstat * age, data = Boston))
```

## Non-linear Transformations of the Predictors

The `lm()` function can also accommodate non-linear transformations of the predictors. For instance, given a predictor $X$, we can create a predictor $X^2$ using
 `I(X^2)`. The function `I()` is needed since the `^` has a special meaning in a formula object; wrapping as we do allows the standard usage in `R`, which is to raise `X` to the power `2`. We now
perform a regression of `medv` onto `lstat` and `lstat^2`.

```{r chunk20}
lm.fit2 <- lm(medv ~ lstat + I(lstat^2))
summary(lm.fit2)
```

The near-zero $p$-value associated with the quadratic term suggests that it leads to an improved model.
We use the `anova()` function  to further quantify the extent to which the quadratic fit is superior to the linear fit.

```{r chunk21}
lm.fit <- lm(medv ~ lstat)
anova(lm.fit, lm.fit2)
```

Here Model 1 represents the linear submodel containing only one predictor, `lstat`, while Model 2 corresponds to the larger quadratic model that has two predictors, `lstat` and `lstat^2`.
The `anova()` function performs a hypothesis test
comparing the two models. The   null hypothesis is that the two models fit the data equally well,  and the alternative hypothesis is that the full model is superior. Here the $F$-statistic is $135$
 and the associated $p$-value is virtually zero. This provides very clear evidence that the model containing the predictors `lstat` and `lstat^2` is far superior to the model that only contains the predictor `lstat`.
 This is not surprising, since earlier we saw evidence for non-linearity in the relationship between `medv` and `lstat`. If we type

```{r chunk22}
par(mfrow = c(2, 2))
plot(lm.fit2)
```

 then we see that when the `lstat^2` term is included in the model, there is little discernible pattern in the residuals.

In order to create a cubic fit, we can include a predictor of the form `I(X^3)`. However, this approach can start to get cumbersome for higher-order polynomials. A better approach involves using the `poly()` function to create the polynomial within `lm()`. For example, the following command produces a
fifth-order polynomial fit:

```{r chunk23}
lm.fit5 <- lm(medv ~ poly(lstat, 5))
summary(lm.fit5)
```

This suggests that including additional  polynomial terms, up to fifth order, leads to an improvement in the model fit! However, further investigation of the data reveals that no polynomial terms beyond fifth order have significant $p$-values
in a regression fit.

 By default, the `poly()` function orthogonalizes the predictors:
 this means that the features output by this function are not simply a
 sequence of powers of the argument. However, a linear model applied to the output of the `poly()` function will have the same fitted values as a linear model applied to the raw polynomials (although the coefficient estimates, standard errors, and p-values will differ). In order to obtain the raw polynomials from the `poly()` function,  the argument `raw = TRUE` must be used.

Of course, we are in no way restricted to using polynomial transformations of the predictors. Here we try a log transformation.

```{r chunk24}
summary(lm(medv ~ log(rm), data = Boston))
```

## Qualitative Predictors

We will now examine the `Carseats` data, which is part of the
`ISLR2` library. We will  attempt to predict `Sales`
(child car seat sales) in $400$ locations based on a number of
predictors.

```{r chunk25}
head(Carseats)
```

The `Carseats` data includes qualitative predictors such as `shelveloc`, an indicator of the quality of the shelving location---that is, the  space within a store in which the car seat is displayed---at each location. The predictor `shelveloc` takes on three possible values:  *Bad*, *Medium*, and *Good*. Given a qualitative variable such as `shelveloc`, `R` generates dummy variables automatically. Below we fit a multiple regression model that includes some interaction terms.

```{r chunk26}
lm.fit <- lm(Sales ~ . + Income:Advertising + Price:Age, 
    data = Carseats)
summary(lm.fit)
```

The `contrasts()` function returns the coding that `R` uses for the dummy variables.

```{r chunk27}
attach(Carseats)
contrasts(ShelveLoc)
```

Use `?contrasts` to learn about other contrasts, and how to set them.

`R` has created a `ShelveLocGood` dummy variable that takes on a value of 1 if the shelving location is good, and 0 otherwise. It has also created a `ShelveLocMedium` dummy variable that equals 1 if the shelving location is medium, and 0 otherwise.
A bad shelving location corresponds to a zero for each of the two dummy variables.
The fact that the coefficient for `ShelveLocGood` in the regression output is positive indicates that a good shelving location is associated with high sales (relative to a bad location). And `ShelveLocMedium` has a smaller positive coefficient, indicating that a medium shelving location is associated with higher sales than a bad shelving location but lower sales than a good shelving location.

### 2022-02-08 _posts/2022-02-05-saturday-mornings-with-tufte/saturday-mornings-with-tufte.Rmd
new post

---
title: "Saturday mornings with Tufte"
description: |
  The other half of data science skillbuilding
author:
  - name: Steve Linberg
    url: https://slinberg.net
date: 2022-02-05
output:
  distill::distill_article:
    self_contained: false
preview: tuftebooks.jpg
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

It takes a lot of sustained effort to build, maintain, and grow robust quantitative skills. Data science is hard, and it's not enough to just have a wealth of formulas and statistical techniques committed to memory: data investigation requires instincts, which can only come as the hard-won results of a great deal of study, practice and inquiry.

As the world re-aligns itself around the tsunami of data that is already hitting us, though, we all know that quant-fu is only half the battle. Being able to manipulate and interpret data fluently is absolutely essential, but no less so is the ability to present and communicate it effectively. Data presentation skills are meaningless (and dangerous) if built atop a weak quantitative foundation, but all the quantitative skills in the world won't save us if we can't make the results of our work understandable.

![Saturday reading](tuftebooks.jpg)

I'm taking a few hours every Saturday morning to read [Edward Tufte](https://www.edwardtufte.com)'s "[The Visual Display of Quantitative Information](https://www.edwardtufte.com/tufte/books_vdqi)" and "[Beautiful Evidence](https://www.edwardtufte.com/tufte/books_be)". I'm not far enough in yet to articulate his vision or integrate his concepts into my own work, but it has the feeling of a correct path that I've learned to recognize from early forays in other fields. A lot of it is about decluttering, stripping away what is superfluous to leave only what is essential, which is one aspect of great art.

![Sample page on time series](vdqi_spread.jpg)

One of the things I love about data science is that this principle is equally critical in both the quantitative and qualitative aspects of our field. Statistical and engineering skills help strip away unneeded or irrelevant data; information design skills help visually present only what is the most important of what remains.

(And, right out of the gate on page 1 of *VDQI*, as the cool kids call it, is an idea for a short quantitative project that I'll take up in the next post.)

![VDQI p1](vdqi_p1.jpg)

### 2022-02-08 _academics/ps753-glossary/ps753-glossary.Rmd
build

@@
 <p style="background: #eee">
-Academics &gt; [PS 753](../ps753/index.html)
+[<i class="fas fa-home"></i>](/) &gt; Academics &gt; [PS 753](../ps753/index.html)
 &gt; Glossary
@@
 
-## Week 1
+## Week 1 - Introduction
 
@@
 
-two-mode data
+two-mode data / bipartite data
 : ties between different sets (or categories) of agents, such as ties between nations and international organizations
@@
 
-## Week 2
+## Week 2 - Network Structure
 
@@
 In the example above, there are 8 nodes that Homer (center) could connect to; among those 8 nodes, there are $7 + 6 + 5 + 4 + 3 + 2 + 1 = 28$ possible undirected connections (not connecting to Homer), and 9 of those 28 do exist, for a local clustering coefficient of $9/28 \approx 0.32$.
+
+## Week 3 - Network Degree
+
+### Class lecture 3a: [Degree](https://www.loom.com/share/49273add09ce4b26924b7f5ab2983810)
+
+degree
+: the number of links that a node has; the number of nodes it's connected to
+
+degree distribution
+: a distribution showing the number of nodes of a network that have each degree level
+
+indegree
+: the number of links that a node *receives* in a directed network
+
+outdegree
+: the number of links that a node *sends* in a directed network
+
+### Class lecture 3b: [Centrality vs. Centralization](https://www.loom.com/share/d2ed1cbb9bc84d27adfbb3febedfeac4)
+
+centrality
+: a measure of the prominence of one node relative to others; can be variously defined
+
+(degree) centrality
+: proportional the the number of other nodes to which a node is linked
+
+centralization
+: a property of a graph, referring to its overall cohesion; comparing most central point to all other points; ratio of the actual sum of differences to the maximum possible sum of differences
+
+### Class lecture 3c: [Network Density](https://www.loom.com/share/6e43e1fca3d44e1881f1a1d4895c26d3)
+
+network density
+: number of ties as a proportion of the maximum possible number of ties; varies from 0 to 1, calculation will vary by whether network is undirected or directed (twice as many potential connections)

### 2022-02-08 _posts/2022-02-05-saturday-mornings-with-tufte/saturday-mornings-with-tufte.Rmd
Add categories

@@
 preview: tuftebooks.jpg
+categories:
+  - tufte
+  - design
 ---

### 2022-02-08 _posts/2022-02-05-saturday-mornings-with-tufte/saturday-mornings-with-tufte.Rmd
text tweak

@@
 
-One of the things I love about data science is that this principle is equally critical in both the quantitative and qualitative aspects of our field.
-Statistical and engineering skills help strip away unneeded or irrelevant data; information design skills help visually present only what is the most important of what remains.
+One of the things I love about data science is that this principle is equally critical in both the quantitative and qualitative aspects of the field.
+Statistical and engineering skills help strip away unneeded or irrelevant data; information design skills help visually present only what is the most important of what remains, to deliver the message as clearly as possible.
 

### 2022-02-08 _posts/2022-02-05-saturday-mornings-with-tufte/saturday-mornings-with-tufte.Rmd
build

--- 2022-02-08 _posts/2022-02-05-saturday-mornings-with-tufte/saturday-mornings-with-tufte.Rmd
@@
 
-One of the things I love about data science is that this principle is equally critical in both the quantitative and qualitative aspects of our field.
-Statistical and engineering skills help strip away unneeded or irrelevant data; information design skills help visually present only what is the most important of what remains.
+One of the things that I love about data science is that this principle is equally critical in both the quantitative and qualitative aspects of the field.
+Statistical and engineering skills help strip away unneeded or irrelevant data; information design skills help visually focus on the most important aspects of what remains, to deliver the message as clearly as possible.
 

### 2022-02-08 _academics/ms797-useful-r-tidbits/ms797-useful-r-tidbits.Rmd
build
Link to ps753 tidbits page, title change

@@
 ---
-title: "Useful R tidbits"
+title: "Useful R tidbits: Machine Learning"
 description: |
-  Tasty R snacks that do useful things
+  Tasty R snacks that do useful things in Machine Learning
 author:
@@
 <p style="background: #eee">
-Academics &gt; [MS797](../ms797/index.html)
+[<i class="fas fa-home"></i>](/) &gt; Academics &gt; [MS797](../ms797/index.html)
 &gt; Useful R tidbits
 </p>
+
+This document contains R snippets pertaining to work in the [Machine Learning class](../ms797/index.html). There is also one for the [Social Networks class](../px753/index.html).
 

### 2022-02-08 _academics/ps753-useful-r-tidbits/ps753-useful-r-tidbits.Rmd
Stub creation

---
title: "Useful R tidbits: Social Neteorks"
description: |
  Tasty R snacks that do useful things in Social Networks
author:
  - name: Steve Linberg
    url: {}
date: 2022-02-08
output:
  distill::distill_article:
    self_contained: false
    toc: true
---

```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = FALSE)
```

<p style="background: #eee">
[<i class="fas fa-home"></i>](/) &gt; Academics &gt; [PS753](../ps753/index.html)
&gt; Useful R tidbits
</p>

This document contains R snippets pertaining to work in the [Social Networks class](../ps753/index.html). There is also one for the [Machine Learning class](../ms797/index.html).

### 2022-02-08 _academics/ps753-glossary/ps753-glossary.Rmd
build
Add useful bits link

@@
 
-degree
+(vertex) degree
 : the number of links that a node has; the number of nodes it's connected to
@@
 
+(closeness) centrality
+: (forthcoming)
+
+(betweenness) centrality
+: (forthcoming)
+
+(eigenvector) centrality
+: (forthcoming)
+
 centralization
-: a property of a graph, referring to its overall cohesion; comparing most central point to all other points; ratio of the actual sum of differences to the maximum possible sum of differences
+: a property of a graph or network, referring to its overall cohesion; comparing most central point to all other points; ratio of the actual sum of differences to the maximum possible sum of differences
 

### 2022-02-08 _academics/ps753-useful-r-tidbits/ps753-useful-r-tidbits.Rmd
Add useful bits link

@@
 ---
-title: "Useful R tidbits: Social Neteorks"
+title: "Useful R tidbits: Social Networks"
 description: |
@@
 
-This document contains R snippets pertaining to work in the [Social Networks class](../ps753/index.html). There is also one for the [Machine Learning class](../ms797/index.html).
+This document contains R snippets pertaining to work in the [Social Networks class](../ps753/index.html). There is also one for the [Machine Learning class](../ms797-useful-r-tidbits/index.html).

### 2022-02-09 _academics/ps753-useful-r-tidbits/ps753-useful-r-tidbits.Rmd
new content

@@
 This document contains R snippets pertaining to work in the [Social Networks class](../ps753/index.html). There is also one for the [Machine Learning class](../ms797-useful-r-tidbits/index.html).
+
+## Week 1 tutorial cheatsheet
+
+||igraph|statnet|
+|---|---|---|
+|Count of vertices|`vcount()`||
+|Count of edges|`ecount()`||
+|Both||`print()`|
+|Bipartite or single mode?|`is_bipartite()`|`print()`|
+|Edges directed or undirected?|`is_directed()`|`print()`|
+|Weighted? (or binary)|`is_weighted()`|`print()`|
+
+vertex attribute
+: vector of additional information about nodes in a network
+
+edge attribute
+: vector of additional information about the edges in a network
+
+Q: how to get missing edge count in igraph?
+
+||igraph|statnet|
+|---|---|---|
+|display vertex attributes|`vertex_attr_names()`|`list.vertex.attributes()`|
+|display edge attributes|`edge_attr_names()`|`list.edge.attributes()`|
+
+In igraph, attributes are accessed via `$`, using the `V` and `E` functions, 
+as in:
+
+```r
+V(karate.ig)$name
+E(karate.ig)$weight
+```
+
+In statnet, they are accessed via the `%v%` and `%e%` mechanisms, as in:
+
+```r
+karate.stat %v% "vertex.names"
+karate.stat %e% "weight"
+```
+
+## Week 2 cheatsheet
+
+A dyad census will count the reciprocal (`mut`), asymmetric (`asym`) and absent (`null`) dyads, based on directed graphs. In igraph:
+
+```r
+igraph::dyad.census(trade2007.ig)
+$mut
+[1] 11444
+
+$asym
+[1] 3148
+
+$null
+[1] 2244
+```
+
+In statnet:
+```r
+sna::dyad.census()
+ [1]   6225  19035  40611   6442   7044  10097  55355  44966   9200   1876
+[11] 146537  25578  17167  30894 225908 374449
+```
+
+Triad census is similar:
+
+```r
+igraph::triad_census(trade2007.ig)
+```
+
+```r
+sna::triad.census(gotbook.stat, mode="graph")  # undirected
+
+sna::triad.census(trade2007.stat)              # directed
+      003   012   102 021D 021U 021C  111D  111U 030T 030C   201  120D
+[1,] 6225 13655 26489 4225 4657 4821 34812 23635 4374  537 97088 15073
+     120U  120C    210    300
+[1,] 7947 14249 136169 283855
+
+```
+
+Note that the `statnet` version gives us information about the types of triads as column names in the matrix it returns.
+The `igraph` version also breaks them into the 16 categories, but returns them in a fixed order not detailed in its return value (described in its help documentation).
+
+The total number of possible triads in a 298 vertex network is (298 x 297 x 296) / (3 x 2 x 1) - the 3 countdown comes from "triad". Quads would be (298 x 297 x 296 x 295) / (4 x 3 x 2 x 1).
+
+## Transitivity
+
+Transitivity is the percentage of potential connected triads - how many are complete.
+Basic way in igraph is `transitivity()`.
+The statnet version is `gtrans()`, but it only works in directed networks.
+`print()` will say whether the network is directed.
+(Note: in the tutorial, we see that the climate network IS directed, but it returns a different result than igraph: 0.627 vs 0.724.
+Not clear why.
+"it is calculating a transitivity score based on an understanding of network structure rooted in hierarchy")
+
+Local transitivity is the `local clustering coefficient` - how many nodes of an ego are connected to each other. Have to unpack this, but the magic is:
+
+```r
+transitivity(gotbook.ig, type="local", 
+  vids=V(gotbook.ig)[
+  c("Petyr Baelish","Jon Snow", "Tyrion Lannister")]) 
+```
+
+The global clustering coefficient in igraph is
+
+```r
+transitivity(trade2007.ig, type="global")
+[1] 0.8837142
+```
+
+The local coefficient is:
+
+```r
+transitivity(trade2007.ig, type="average")
+[1] 0.8862707
+```
+
+Network transitivity in statnet is `gtrans()`:
+
+```r
+gtrans(trade2007.stat)
+[1] 0.9993143
+```
+
+||igraph|statnet|
+|---|---|---|
+|global clustering coefficient|`transitivity(trade2007.ig,  type="global")`|`gtrans(trade2007.stat)` (directed only)|
+|local clustering coefficient|`transitivity(trade2007.ig,  type="local")`|???|
+|average local clustering coefficient|`transitivity(trade2007.ig,  type="average")`|???|
+
+## Degree
+
+`igraph::degree()` and `statnet::degree()`, and once again they give different results; igraph includes loops, statnet doesn't.
+Force igraph to ignore them with `loops = FALSE`.
+"Note that setting `diag=TRUE` in `sna::degree` does not guarantee equivalence as statnet only single counts the loops in a directed network, while igraph double-counts the loops."
+
+igraph shows the node names, statnet doesn't.
+
+Getting the degree of a particular set of nodes in `igraph`:
+
+```r
+igraph::degree(trade2007.ig, v = V(trade2007.ig)[c("China", "Canada", "United Kingdom", "Denmark")])
+         China         Canada United Kingdom        Denmark 
+           364            364            364            362 
+```
+
+### *Indegree* and *outdegree* calculations
+
+||igraph|statnet|
+|---|---|---|
+|indegree|`igraph::degree(climate.ig,mode="in", loops = FALSE)`|`sna::degree(climate.stat, cmode="indegree")`|
+|outdegree|`igraph::degree(climate.ig,mode="out", loops = FALSE)`|`sna::degree(climate.stat, cmode="outdegree")`|
+
+Code from the tutorial to create `data.frame`s with degree statistics:
+
+```r
+#igraph:
+
+trade2007.nodes <- data.frame(name = V(trade2007.ig)$name,
+    totdegree = igraph::degree(trade2007.ig, loops = FALSE),
+    indegree = igraph::degree(trade2007.ig, mode = "in", loops = FALSE),
+    outdegree = igraph::degree(trade2007.ig, mode = "out", loops = FALSE))
+
+#statnet version:
+
+trade2007.nodes <- data.frame(name = trade2007.stat%v%"vertex.names",
+    totdegree = sna::degree(trade2007.stat),
+    indegree = sna::degree(trade2007.stat, cmode = "indegree"),
+    outdegree = sna::degree(trade2007.stat, cmode = "outdegree"))
+```
+
+## Components
+
+Shortest path length between 2 nodes: igraph `distances()` does this.
+
+```r
+distances(gotbook.ig,"Petyr Baelish","Robb Stark")
+# Calculate distance using unweighted edges
+distances(gotbook.ig,"Petyr Baelish","Robb Stark", weights=NA)
+# list shortest paths between 2 nodes
+all_shortest_paths(gotbook.ig,"Bronn","Varys", weights=NA)$res
+#find average shortest path for network
+average.path.length(gotbook.ig,directed=F)
+```
+
+Component structure and membership
+
+```r
+# What element are returned by components
+names(igraph::components(gotbook.ig))
+
+# Number of components
+igraph::components(gotbook.ig)$no
+
+# Size of each component
+igraph::components(gotbook.ig)$csize
+
+# retrieve the index of isolate nodes
+# (nodes with component count of 1 from "components" above)
+isolates(gotbook.stat)
+
+# subset vertex.names attribute to get names of isolates
+as.vector(gotbook.stat %v% 'vertex.names')[c(isolates(gotbook.stat))]
+```
+
+## Graph density
+
+Note: `network.density()` (statnet) ignores edge values "at present".
+
+```r
+#get network density: igraph
+graph.density(climate.ig)
+
+## [1] 0.4117647
+
+#get network density: statnet
+network.density(climate.stat)
+
+## [1] 0.399654
+```
+
+Adding `loops = TRUE` to `graph.density()` appears to fix the problem and gets the two packages to agree:
+
+```r
+#get network density without loops: igraph
+graph.density(climate.ig, loops=TRUE)
+
+## [1] 0.399654
+```
+
+*SO*, it's safest to always do either:
+- `graph.density(climate.ig, loops=TRUE)` (igraph), OR
+- `network.density(climate.stat)` (statnet)
+
+In statnet, we can get network density with loops (nodes connecting to themselves) omitted:
+
+```r
+#get network density without loops: statnet
+gden(climate.stat, diag=FALSE)
+
+## [1] 0.3921569
+```
+
+## Network degree centralization
+
+In `statnet`, call `centralization()` with the `degree` function and appropriate parameters for `degree` in the `cmode` argument:
+
+```r
+centralization(climate.stat, degree, cmode="indegree")
+centralization(climate.stat, degree, cmode="outdegree")
+centralization(climate.stat, degree, cmode="freeman") # default
+```
+
+Could also call it with other `sna` functions like `betweenness`, `closeness`; tutorial references `eigenvalue` but there is no such function in `sna`. `evcent` perhaps?
+
+The `igraph` version uses `centr_degree()` and returns an object with several components, of which `centralization` is one:
+
+```r
+centr_degree(climate.ig, loops = FALSE, mode = "in")$centralization
+centr_degree(climate.ig, loops = FALSE, mode = "out")$centralization
+```

### 2022-02-09 _academics/ps753-glossary/ps753-glossary.Rmd
text tweaks

@@
 two-mode data / bipartite data
-: ties between different sets (or categories) of agents, such as ties between nations and international organizations
+: ties between different sets (or categories) of agents, such as ties between nations and international organizations; a network split into two parts, each of whose nodes only connect to nodes in the opposite part, not to nodes within its own part
 
@@
 
+average clustering coefficient
+: the average of the local clustering coefficient of all nodes in the network
+
 ## Week 3 - Network Degree

### 2022-02-09 _academics/ms797-glossary/ms797-glossary.Rmd
build

@@
    toc: true
+categories:
+  - machine learning
+  - glossary
 ---
@@
 <p style="background: #eee">
-Academics &gt; [MS797](../ms797/index.html)
+[<i class="fas fa-home"></i>](/) &gt; Academics &gt; [MS797](../ms797/index.html)
 &gt; Glossary

### 2022-02-09 _academics/ms797-useful-r-tidbits/ms797-useful-r-tidbits.Rmd
Add useful bits link
build

@@
 preview: ISLR-cover.png
+categories:
+  - r
+  - machine learning
 ---
@@
 ```{r setup, include=FALSE}
-# knitr::opts_chunk$set(echo = FALSE)
+knitr::opts_chunk$set(echo = FALSE)
 ```
@@
 
-This document contains R snippets pertaining to work in the [Machine Learning class](../ms797/index.html). There is also one for the [Social Networks class](../px753/index.html).
+This document contains R snippets pertaining to work in the [Machine Learning class](../ms797/index.html). There is also one for the [Social Networks class](../ps753-useful-r-tidbits/index.html).
 

### 2022-02-09 _academics/ms797/ms797.Rmd
build

@@
 citation: false
+categories:
+  - machine learning
 ---
@@
 <p style="background: #eee">
-Academics &gt; MS797
+[<i class="fas fa-home"></i>](/) &gt; Academics &gt; MS797
 </p>
@@
 - My ongoing [glossary](../ms797-glossary/index.html) page; feel free to comment, or make additions or corrections via comments or github.
+- Some [useful R tidbits](../ms797-useful-r-tidbits/index.html) from, or inspired by, the text.

### 2022-02-09 _academics/ps753-glossary/ps753-glossary.Rmd
build

@@
 bibliography: PS753.bib
+categories:
+  - social networks
+  - glossary
 ---

### 2022-02-09 _academics/ps753-useful-r-tidbits/ps753-useful-r-tidbits.Rmd
build

@@
     toc: true
+categories:
+  - r
+  - social networks
 ---

### 2022-02-09 _academics/ps753/ps753.Rmd
Add pseudo top-level nav for /academics
build
Add useful bits link

@@
 citation: false
+categories:
+  - social networks
 ---
@@
 <p style="background: #eee">
-PS 753
+[<i class="fas fa-home"></i>](/) &gt; Academics &gt; PS 753
 </p>
@@
 - My ongoing [glossary](../ps753-glossary/index.html) page; feel free to comment, or make additions or corrections via comments or github.
+- Some [useful R tidbits](../ps753-useful-r-tidbits/index.html) from the course material and related sources.

### 2022-02-09 _academics/ms797-useful-r-tidbits/ms797-useful-r-tidbits.Rmd
couple of tweaks

@@
 
-XXX
-<hr>
+# EVERYTHING BELOW THIS LINE
+
+is direct paste from the book's sources which I have not yet written in my own words or added anything to. Stay tuned.
+
+<hr />
 
@@
 In order to fit a multiple linear regression model using least squares, we again use the `lm()` function.
-The syntax {\R{lm(y $\sim$ x1 + x2 + x3)}} is used to fit a model with three predictors, `x1`, `x2`, and `x3`.
+The syntax `lm(y ~ x1 + x2 + x3)` is used to fit a model with three predictors, `x1`, `x2`, and `x3`.
 The `summary()` function now outputs the regression coefficients for all the predictors.
@@
 The syntax `lstat * age` simultaneously includes `lstat`, `age`, and the interaction term `lstat`$\times$`age` as predictors; it is a shorthand for `lstat + age + lstat:age`.
-  %We can also pass in transformed versions of the predictors.
 

### 2022-02-10 _academics/ps753-hw3/ps753-hw3.Rmd
New post

---
title: "Homework 3: The Florentine Wedding dataset"
description: |
  A first-look network analysis
author:
  - name: Steve Linberg
    url: {}
date: 2022-02-09
output:
  distill::distill_article:
    self_contained: false
categories:
  - r
  - social networks
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
library(network)
library(igraph)
library(tidyverse)

data(flo)
ntwk.ig <- graph_from_adjacency_matrix(flo)

ntwk.vcount <- vcount(ntwk.ig)
ntwk.ecount <- ecount(ntwk.ig)
ntwk.is_bipartite <- is_bipartite(ntwk.ig)
ntwk.is_directed <- is_directed(ntwk.ig)
ntwk.is_weighted <- is_weighted(ntwk.ig)
ntwk.vertex_attr_names <- vertex_attr_names(ntwk.ig)
ntwk.edge_attr_names <- edge_attr_names(ntwk.ig)
ntwk.names <- V(ntwk.ig)$name
ntwk.dyad_census <- igraph::dyad.census(ntwk.ig)
ntwk.triad_census <- igraph::triad_census(ntwk.ig)
ntwk.global_clustering_coef <- transitivity(ntwk.ig, type = "global")
ntwk.local_avg_clustering_coef <- transitivity(ntwk.ig, type = "average")
ntwk.avg_path_length <- average.path.length(ntwk.ig, directed = T)
ntwk.component_names <- names(igraph::components(ntwk.ig))
ntwk.components_no <- igraph::components(ntwk.ig)$no
ntwk.components_csize <- igraph::components(ntwk.ig)$csize

# distances(ntwk.ig,"Acciaiuoli","Strozzi", weights=NA)

ntwk.graph_density <- graph.density(ntwk.ig)
ntwk.graph_density_noloops <- graph.density(ntwk.ig, loops = TRUE)

ntwk.avg_degree <- igraph::degree(ntwk.ig)

# Get the degree info and make the data frame

ntwk.nodes <- data.frame(
  name = V(ntwk.ig)$name, 
  degree = igraph::degree(ntwk.ig),
  indegrees = igraph::degree(ntwk.ig, mode = "in", loops = FALSE),
  outdegrees = igraph::degree(ntwk.ig, mode = "out", loops = FALSE)
  )
# ntwk.nodes[ntwk.nodes$name %in% c("Medici", "Strozzi"),]

# Get degree centralization

ntwk.indegree_centralization <- centr_degree(ntwk.ig, loops = FALSE, mode = "in")$centralization
ntwk.outdegree_centralization <- centr_degree(ntwk.ig, loops = FALSE, mode = "out")$centralization
ntwk.total_centralization <- centr_degree(ntwk.ig, loops = FALSE, mode = "total")$centralization
```

<p style="background: #eee">
[<i class="fas fa-home"></i>](/) &gt; Academics &gt; [PS753](../ps753/index.html)
&gt; Homework 3: The Florentine Wedding dataset
</p>

For this assignment, I am looking at the Florentine Wedding dataset by J. F. Padgett, representing intermarriages between prominent families in Florence between 1282 and 1500. This is an area of interest to me as I have some background in fine art, and the Medici family in particular was known for its patronage of some of history's greatest artists throughout the Renaissance, including Boticelli, Leonardo da Vinci, and Michelangelo. No less than Galileo Galilei himself tutored several generations of Medici children.

The great families of this era were frequently fierce rivals, with delicately balanced power structures measured in various spheres of influence; intermarriages were one way that alliances were forged between families in the ceaseless effort to gain political or financial advantage over their competitors.

The Medici family was the most prominent in Florence throughout the renaissance. Two other prominent families, the Strozzi and Guadagni, also exercised considerable influence (they were all banking families), and did not intermarry with the Medici, though they were not far removed. This essay looks at the marriage network between the various families as one measure of social power and centrality in Renaissance Florence.

The `flo` dataset in the R `network` library is an adjacency matrix of 16 families of this era, including the aforementioned. It contains the 16 nodes of the families, and 20 unweighted, directed edges between them. The Pucci family is an isolate, making the network disconnected; the remaining nodes form a single component.

The families in the dataset, with their total intermarriage count (`degree`), marriages into each family (`indegrees`) and marriages out to other families (`outdegrees`) are:

```{r}
ntwk.nodes[,c("degree", "indegrees", "outdegrees")]
```

A visualization shows each marriage as a line connecting two families:

```{r}
V(ntwk.ig)$size <- ntwk.nodes[,"degree"] * 5
V(ntwk.ig)$label.cex <- ntwk.nodes[,"degree"] / 8
V(ntwk.ig)$label.cex <- pmax(V(ntwk.ig)$label.cex, 0.75)

plot(ntwk.ig, layout = layout_in_circle(ntwk.ig), edge.arrow.mode = 0, vertex.label.color="black", vertex.shape = "circle", vertex.label.font = 2, main = "Florence interfamily marriages, 1282-1500")
```

Each node is sized according to its number of connections; it is easy to see from this diagram that the Medici family had the most marriages within the Florence family network, with six. The Strozzi and Guadagni families were next at four apiece.

A component layout offers a different view that makes distances and paths easier to read:

```{r}
set.seed(1235)
V(ntwk.ig)$size <- ntwk.nodes[,"degree"] * 3
plot(ntwk.ig, layout = layout_components(ntwk.ig), edge.arrow.mode = 0, vertex.label.color="black", vertex.shape = "circle", vertex.label.font = 2, main = "Florence interfamily marriages, 1282-1500")

```
From this layout, we can see that the Medici, Strozzi and Guadagni never directly intermarried, but were also only separated from each other by one node. With an overall average path length of approximately 2.5, this suggests some calculus and deliberation: not directly joining with their main rivals, but remaining only one degree removed.

The overall network density is $\approx$ 0.16; 20 connections by marriage out of a possible total of 120. Four of the 16 nodes are pendants. We might expect the network density to increase if we looked at more families than the 16 listed here; it is very unlikely that only 16% of families experienced any marriages at all! This would need to be weighed against the prominence of the additional families according to the criteria with which this initial set was selected.

It would be interesting to me to look deeper into a bipartite network of connections between the families and who they patronized, especially the artists and scientists of the era. This would require more data than is present in this set, but could be very revealing about other ways in which the families interacted, competing for social and political influence in the dynamic period of the Renaissance.

### 2022-02-10 _academics/ps753-hw3/ps753-hw3.Rmd
build

@@
 	message = FALSE,
-	warning = FALSE
+	warning = FALSE,
+	fig.height = 8
 )
@@
 V(ntwk.ig)$size <- ntwk.nodes[,"degree"] * 5
-V(ntwk.ig)$label.cex <- ntwk.nodes[,"degree"] / 8
-V(ntwk.ig)$label.cex <- pmax(V(ntwk.ig)$label.cex, 0.75)
+V(ntwk.ig)$label.cex <- ntwk.nodes[,"degree"] / 4
+V(ntwk.ig)$label.cex <- pmax(V(ntwk.ig)$label.cex, 1)
 
@@
 
-It would be interesting to me to look deeper into a bipartite network of connections between the families and who they patronized, especially the artists and scientists of the era.
+It would be interesting to me to look deeper into a bipartite network of connections between the families and whom they patronized, especially the artists and scientists of the era.
 This would require more data than is present in this set, but could be very revealing about other ways in which the families interacted, competing for social and political influence in the dynamic period of the Renaissance.

### 2022-02-10 _academics/ps753-hw3/ps753-hw3.Rmd
Add code fold options

@@
     self_contained: false
+    code_folding: true
 categories:
@@
 
+<p style="background: #eee">
+[<i class="fas fa-home"></i>](/) &gt; Academics &gt; [PS753](../ps753/index.html)
+&gt; Homework 3: The Florentine Wedding dataset
+</p>
+
 ```{r setup, include=FALSE}
@@
 	echo = FALSE,
+	fig.height = 8,
 	message = FALSE,
-	warning = FALSE,
-	fig.height = 8
+	warning = FALSE
 )
@@
 library(tidyverse)
+```
 
+```{r}
 data(flo)
@@
 ```
-
-<p style="background: #eee">
-[<i class="fas fa-home"></i>](/) &gt; Academics &gt; [PS753](../ps753/index.html)
-&gt; Homework 3: The Florentine Wedding dataset
-</p>
+</details>
 

### 2022-02-10 _academics/ps753-hw3/ps753-hw3.Rmd
tidy code

@@
 
-ntwk.indegree_centralization <- centr_degree(ntwk.ig, loops = FALSE, mode = "in")$centralization
-ntwk.outdegree_centralization <- centr_degree(ntwk.ig, loops = FALSE, mode = "out")$centralization
-ntwk.total_centralization <- centr_degree(ntwk.ig, loops = FALSE, mode = "total")$centralization
+ntwk.indegree_centralization <- 
+  centr_degree(ntwk.ig, loops = FALSE, mode = "in")$centralization
+ntwk.outdegree_centralization <- 
+  centr_degree(ntwk.ig, loops = FALSE, mode = "out")$centralization
+ntwk.total_centralization <- 
+  centr_degree(ntwk.ig, loops = FALSE, mode = "total")$centralization
 ```
@@
 
-plot(ntwk.ig, layout = layout_in_circle(ntwk.ig), edge.arrow.mode = 0, vertex.label.color="black", vertex.shape = "circle", vertex.label.font = 2, main = "Florence interfamily marriages, 1282-1500")
+plot(
+  ntwk.ig,
+  layout = layout_in_circle(ntwk.ig),
+  edge.arrow.mode = 0,
+  vertex.label.color = "black",
+  vertex.shape = "circle",
+  vertex.label.font = 2,
+  main = "Florence interfamily marriages, 1282-1500"
+)
 ```
@@
 V(ntwk.ig)$size <- ntwk.nodes[,"degree"] * 3
-plot(ntwk.ig, layout = layout_components(ntwk.ig), edge.arrow.mode = 0, vertex.label.color="black", vertex.shape = "circle", vertex.label.font = 2, main = "Florence interfamily marriages, 1282-1500")
+plot(
+  ntwk.ig,
+  layout = layout_components(ntwk.ig),
+  edge.arrow.mode = 0,
+  vertex.label.color = "black",
+  vertex.shape = "circle",
+  vertex.label.font = 2,
+  main = "Florence interfamily marriages, 1282-1500"
+)
 

### 2022-02-10 _academics/ms797-glossary/ms797-glossary.Rmd
Add correlation formula LaTeX

@@
 : a measure of the linearity of the relationship between $X$ and $Y$; values close to 0 indicate weak-to-no relationship, values near 1 or -1 indicate strong positive or negative correlation
-(horrible formula on p.70 I don't have time to LaTeX at the moment)
+$$\text{Cor(X, Y)} = {\frac {\sum _{i=1}^{n}(x_{i}-{\bar {x}})(y_{i}-{\bar {y}})}{{\sqrt {\sum _{i=1}^{n}(x_{i}-{\bar {x}})^{2}}}{\sqrt {\sum _{i=1}^{n}(y_{i}-{\bar {y}})^{2}}}}}$$
 

### 2022-02-15 _academics/ps753-useful-r-tidbits/ps753-useful-r-tidbits.Rmd
Add isolotes command for igraph

@@
 isolates(gotbook.stat)
+# No direct command in igraph, but can do this:
+deg_counts <- igraph::degree(gotbook.ig, loops = FALSE)
+length(deg_counts[deg_counts == 0])
 

### 2022-02-16 _academics/ps753-glossary/ps753-glossary.Rmd
knits

@@
 
-(closeness) centrality
-: (forthcoming)
-
-(betweenness) centrality
-: (forthcoming)
-
-(eigenvector) centrality
-: (forthcoming)
-
 centralization
@@
 : number of ties as a proportion of the maximum possible number of ties; varies from 0 to 1, calculation will vary by whether network is undirected or directed (twice as many potential connections)
+
+## Week 4 - Status & Eigenvectors
+
+### Class lecture 4a: [Status and Hierarchy](https://www.loom.com/share/91c521e65705482cb7249140be3b0274)
+
+closeness centrality
+: the sum of geodesic distances (shortest paths) to all other points in the graph. Divide by $n-1$, then invert. A measure of how close a node is to all of the other nodes in the network.
+
+### Class lecture 4b: [Status and Prestige](https://www.loom.com/share/e0484b4af1bb4d3b8242715c8cba30f8)
+
+prestige
+: signal of the quality of a node, of a node's visibility within the network
+
+eigenvector centrality
+: "takes into account the centrality of other nodes to which a node is connected.
+That is, being connected to other central nodes increases centrality." Takes into account the centrality of the nodes a node is connected to, considering the importance of the connected nodes and not just their quantity or path length.
+$$\lambda \text{c} = \text{Wc}$$
+
+Bonacich Power
+: "A closely related concept, but includes a weighting factor that emphasizes global vs. local structure (and negative connections)"
+$$\text{c}(\alpha, \beta) = \alpha(\text{I} - \beta\text{W})^{-1}\text{W1}$$
+
+### Class lecture 4c: [Hubs and Bridges](https://www.loom.com/share/e85fa1918f70438e9bb67232ade95594)
+
+bridge
+: a node with few ties to central actors
+
+hub
+: a node with many ties to peripheral actors
+
+reflected centrality
+: the degree of centrality that a node receives from a connected node that is due to that node's centrality.
+If A has influential friend B, then A's eigenvector centrality will be higher; reflected centrality is the portion of A's centrality that comes from B's centrality.
+
+derived centrality
+: the remainder of the eigenvector centrality that A receives from B in the example above, that is not due to B's centrality but just from B being a connected node
+
+Reflected and derived centrality can be represented in the following matrix by Mizruchi et al:
+
+||High reflected centrality|Low reflected centrality|
+|---|---|---|
+|High derived centrality|Cosmopolitans|Pure bridges|
+|Low derived centrality|Pure hubs|Peripherals|
+
+The four prototypes are:
+
+cosmopolitans
+: well-connected nodes that are connected to other well-connected nodes; the "cool kids" table
+
+pure hubs
+: well-connected nodes that are connected to nodes that are not well-connected; the cool kid who talks to uncool kids
+
+pure bridge
+: a node connected to few but high-centrality nodes; a friend of the cool kid with few other friends
+
+peripheral
+: a node with few and low-centrality connections; me and my D&D nerd friends in high school
+
+### Class lecture 4d: [Status and Core/Periphery Networks](https://www.loom.com/share/2b5fdf986b164bb785a4c8c0fca0d1e9)
+
+betweenness centrality
+: (forthcoming)

### 2022-02-16 _academics/ps753-useful-r-tidbits/ps753-useful-r-tidbits.Rmd
knits

--- 2022-02-09 _academics/ps753-useful-r-tidbits/ps753-useful-r-tidbits.Rmd
@@
 isolates(gotbook.stat)
+# There is no direct command in igraph, but we can do this:
+# Create a list of the degree of each node in the network
+deg_counts <- igraph::degree(gotbook.ig, loops = FALSE)
+# filter and count the nodes with 0 degrees (or any other quantity of interest)
+length(deg_counts[deg_counts == 0])
 
@@
 
-Could also call it with other `sna` functions like `betweenness`, `closeness`; tutorial references `eigenvalue` but there is no such function in `sna`. `evcent` perhaps?
+Could also call it with other `sna` functions like `betweenness`, `closeness`
 
@@
 ```
+
+### Eigenvector centralization
+
+`statnet` uses `evcent()` to calculate the eigenvalue centrality score for each node in the network:
+
+```r
+evcent(imf2014.stat, ignore.eval=TRUE))
+```
+
+Eigenvector centrality index for the network:
+
+```r
+centralization(imf2014.stat, evcent)
+```
+
+In `igraph`, a set of eigenvector-related information is created with `centr_eigen()`:
+
+```r
+# If the network is directed, specify "directed - T" - will not auto-detect
+eigen_info <- centr_eigen(imf2014.ig, directed = T)
+# Centrality score for node 3:
+eigen_info[3]$vector
+# Eigenveector centrality index for the network:
+eigen_info$centralization
+```
+
+**The scores calculated by `igraph` and `statnet` are different.
+We aren't sure why.** It appears that igraph [counts incoming ties to calculate eigenvector centrality](https://stackoverflow.com/questions/21035598/what-is-evcent-returning-for-directed-graphs), and statnet [recommends using Bonachic power instead for directed networks](http://mailman13.u.washington.edu/pipermail/statnet_help/2015/002170.html).
+
+### Bonacich Power Centrality
+
+`igraph`:
+
+```r
+power_centrality(imf2014.ig)
+```
+
+`statnet`:
+
+```r
+bonpow(imf2014.stat)
+```
+
+Again, there appear to be some inconsistency between `igraph` and `statnet` in the calculations, with statnet apparently not incorporating weights and failing on singluar matrices.
+
+### Derived and Reflected Centrality
+
+There are no library routines for these calculations. Convert the data to a matrix first:
+
+```r
+mat2014 <- as.matrix(as_adjacency_matrix(imf2014.ig, attr="weight"))
+```
+
+> To calculate the proportion of centrality that is received, we first square the adjacency matrix.
+The diagonal of the adjacency matrix is equal to the the square of node degree.
+We then divide this diagonal (sqared degree) by total sqaured indegree (calculated by rowSums) to get the proportion of received centrality.
+
+```r
+# square the adjacency matrix
+mat2014sq <- t(mat2014) %*% mat2014
+
+# Calculate the proportion of reflected centrality.
+imf2014.nodes$rc <- diag(mat2014sq) / rowSums(mat2014sq)
+
+# freplace missing values with 0
+imf2014.nodes$rc <- ifelse(is.nan(imf2014.nodes$rc), 0, imf2014.nodes$rc)
+
+# Calculate received eigenvalue centrality
+imf2014.nodes$eigen.rc <- imf2014.nodes$eigen * imf2014.nodes$rc
+```
+
+> If total centraltiy is 1, then derived centrality is simply 1 - the proportion of eigenvector centrality due to received centrality.
+
+```r
+# Calculate the proportion of derived centrality.
+imf2014.nodes$dc <- 1 - diag(mat2014sq) / rowSums(mat2014sq)
+
+# replace missing values with 0
+imf2014.nodes$dc <- ifelse(is.nan(imf2014.nodes$dc), 1, imf2014.nodes$dc)
+
+# Calculate received eigenvalue centrality
+imf2014.nodes$eigen.dc <- imf2014.nodes$eigen * imf2014.nodes$dc
+```

### 2022-02-17 _academics/ps753-hw4/ps753-hw4.Rmd
Initial commit

--- 2022-02-10 _academics/ps753-hw3/ps753-hw3.Rmd
@@
 ---
-title: "Homework 3: The Florentine Wedding dataset"
+title: "Homework 4: Centrality Analysis of the Florentine Wedding dataset"
 description: |
@@
 
+<p style="background: #eee">
+[<i class="fas fa-home"></i>](/) &gt; Academics &gt; [PS753](../ps753/index.html)
+&gt; Homework 4: Centrality Analysis of the Florentine Wedding dataset
+</p>
+
 ```{r setup, include=FALSE}
@@
 
-ntwk.indegree_centralization <- centr_degree(ntwk.ig, loops = FALSE, mode = "in")$centralization
-ntwk.outdegree_centralization <- centr_degree(ntwk.ig, loops = FALSE, mode = "out")$centralization
-ntwk.total_centralization <- centr_degree(ntwk.ig, loops = FALSE, mode = "total")$centralization
+ntwk.indegree_centralization <- 
+  centr_degree(ntwk.ig, loops = FALSE, mode = "in")$centralization
+ntwk.outdegree_centralization <- 
+  centr_degree(ntwk.ig, loops = FALSE, mode = "out")$centralization
+ntwk.total_centralization <- 
+  centr_degree(ntwk.ig, loops = FALSE, mode = "total")$centralization
 ```
 
-<p style="background: #eee">
-[<i class="fas fa-home"></i>](/) &gt; Academics &gt; [PS753](../ps753/index.html)
-&gt; Homework 3: The Florentine Wedding dataset
-</p>
-
-For this assignment, I am looking at the Florentine Wedding dataset by J.
-F.
-Padgett, representing intermarriages between prominent families in Florence between 1282 and 1500.
-This is an area of interest to me as I have some background in fine art, and the Medici family in particular was known for its patronage of some of history's greatest artists throughout the Renaissance, including Boticelli, Leonardo da Vinci, and Michelangelo.
-No less than Galileo Galilei himself tutored several generations of Medici children.
-
-The great families of this era were frequently fierce rivals, with delicately balanced power structures measured in various spheres of influence; intermarriages were one way that alliances were forged between families in the ceaseless effort to gain political or financial advantage over their competitors.
-
-The Medici family was the most prominent in Florence throughout the renaissance.
-Two other prominent families, the Strozzi and Guadagni, also exercised considerable influence (they were all banking families), and did not intermarry with the Medici, though they were not far removed.
-This essay looks at the marriage network between the various families as one measure of social power and centrality in Renaissance Florence.
-
-The `flo` dataset in the R `network` library is an adjacency matrix of 16 families of this era, including the aforementioned.
-It contains the 16 nodes of the families, and 20 unweighted, directed edges between them.
-The Pucci family is an isolate, making the network disconnected; the remaining nodes form a single component.
-
-The families in the dataset, with their total intermarriage count (`degree`), marriages into each family (`indegrees`) and marriages out to other families (`outdegrees`) are:
-
-```{r}
-ntwk.nodes[,c("degree", "indegrees", "outdegrees")]
-```
-
-A visualization shows each marriage as a line connecting two families:
-
-```{r}
-V(ntwk.ig)$size <- ntwk.nodes[,"degree"] * 5
-V(ntwk.ig)$label.cex <- ntwk.nodes[,"degree"] / 4
-V(ntwk.ig)$label.cex <- pmax(V(ntwk.ig)$label.cex, 1)
-
-plot(ntwk.ig, layout = layout_in_circle(ntwk.ig), edge.arrow.mode = 0, vertex.label.color="black", vertex.shape = "circle", vertex.label.font = 2, main = "Florence interfamily marriages, 1282-1500")
-```
-
-Each node is sized according to its number of connections; it is easy to see from this diagram that the Medici family had the most marriages within the Florence family network, with six.
-The Strozzi and Guadagni families were next at four apiece.
-
-A component layout offers a different view that makes distances and paths easier to read:
-
-```{r}
-set.seed(1235)
-V(ntwk.ig)$size <- ntwk.nodes[,"degree"] * 3
-plot(ntwk.ig, layout = layout_components(ntwk.ig), edge.arrow.mode = 0, vertex.label.color="black", vertex.shape = "circle", vertex.label.font = 2, main = "Florence interfamily marriages, 1282-1500")
-
-```
-From this layout, we can see that the Medici, Strozzi and Guadagni never directly intermarried, but were also only separated from each other by one node.
-With an overall average path length of approximately 2.5, this suggests some calculus and deliberation: not directly joining with their main rivals, but remaining only one degree removed.
-
-The overall network density is $\approx$ 0.16; 20 connections by marriage out of a possible total of 120.
-Four of the 16 nodes are pendants.
-We might expect the network density to increase if we looked at more families than the 16 listed here; it is very unlikely that only 16% of families experienced any marriages at all!
-This would need to be weighed against the prominence of the additional families according to the criteria with which this initial set was selected.
-
-It would be interesting to me to look deeper into a bipartite network of connections between the families and whom they patronized, especially the artists and scientists of the era.
-This would require more data than is present in this set, but could be very revealing about other ways in which the families interacted, competing for social and political influence in the dynamic period of the Renaissance.
+This week, we're going to look at various measures of *degree centrality* in the Florentine Wedding dataset.

### 2022-02-17 _academics/ps753-hw4/ps753-hw4.Rmd
knit

@@
 This week, we're going to look at various measures of *degree centrality* in the Florentine Wedding dataset.
+
+[Last week](../ps753-hw3/index.html), we looked at some basic network plots that showed the overall structure of the network, sizing the nodes by their individual degrees.
+One of the primary features of this data set is all of the ways in which it can be shown how dominant the Medici family was in medieval Florence.
+
+```{r}
+V(ntwk.ig)$size <- ntwk.nodes[,"degree"] * 5
+V(ntwk.ig)$label.cex <- ntwk.nodes[,"degree"] / 4
+V(ntwk.ig)$label.cex <- pmax(V(ntwk.ig)$label.cex, 1)
+
+set.seed(1235)
+V(ntwk.ig)$size <- ntwk.nodes[,"degree"] * 3
+plot(
+  ntwk.ig,
+  layout = layout_components(ntwk.ig),
+  edge.arrow.mode = 0,
+  vertex.label.color = "black",
+  vertex.shape = "circle",
+  vertex.label.font = 2,
+  main = "Florence interfamily marriages, 1282-1500"
+)
+```
+
+## Closeness Centrality
+
+*Closeness centrality* is a measure of how close a node is, via the shortest path, to every other node in the network.
+It is more revealing in hierarchies, where dominant nodes that are closer to the center of the graph will have, on average, shorter paths to the rest of the node than the nodes at the periphery.
+
+The closeness centrality measures of the nodes in the Florentine data may come as a bit of a surprise:
+
+```{r}
+closeness_centrality <- closeness(ntwk.ig, mode = "all")
+# Calculate node size by 1000 x closeness
+V(ntwk.ig)$size <- closeness_centrality * 1000
+# Ensure minimum node size
+V(ntwk.ig)$size <- pmax(V(ntwk.ig)$size, 1)
+# Base text size on node size
+V(ntwk.ig)$label.cex <- V(ntwk.ig)$size / 15
+V(ntwk.ig)$label.cex <- pmax(V(ntwk.ig)$label.cex, 1)
+set.seed(1235)
+closeness_labels <- sprintf("%s\n%s", ntwk.names,
+                            # Omit leading "0" can't be done in sprintf
+                            substr(as.character(round(
+                              closeness_centrality, 3
+                            )), 2, 10))
+plot(
+  ntwk.ig,
+  layout = layout_components(ntwk.ig),
+  vertex.label = closeness_labels,
+  edge.arrow.mode = 0,
+  vertex.label.color = "black",
+  vertex.shape = "circle",
+  vertex.label.font = 2,
+  main = "Florence nodes by closeness centrality"
+)
+```
+
+Although careful examination shows that the Medici family is still the dominant node, with the highest closeness centrality score, there is very little variation among all but the disconnected Pucci node.
+The chief reason for this is that the network itself is very small, and tightly connected.
+As we saw last week, the average path length for this network is `r round(ntwk.avg_path_length, 3)`; the Florentine families were all in close proximity to each other, as might be expected given that one of its primary constraints is geographical, but it is interesting to see how the power of the Medici family did *not* derive primarily from its closeness centrality, or its position in a hierarchy.
+
+## Betweenness Centrality
+
+Another useful measure is *betweenness centrality*, which is a count of the number of shortest paths between nodes that pass through another node.
+It can be seen as a measure of importance or control, in that it reflects how well a node is positioned along the optimal path between two other nodes.
+
+```{r}
+betweenness_centrality <- betweenness(ntwk.ig)
+# Calculate node size by 2 x betweenness
+V(ntwk.ig)$size <- betweenness_centrality / 2
+# Ensure minimum node size
+V(ntwk.ig)$size <- pmax(V(ntwk.ig)$size, 10)
+# Base text size on node size
+V(ntwk.ig)$label.cex <- V(ntwk.ig)$size / 15
+V(ntwk.ig)$label.cex <- pmax(V(ntwk.ig)$label.cex, 1)
+
+colors <- case_when(
+  betweenness_centrality == 0 ~ "#FFFF33",
+  betweenness_centrality < 20 ~ "#FF8033",
+  TRUE ~ "#FF0033"
+)
+set.seed(1235)
+plot(
+  ntwk.ig,
+  vertex.color = colors,
+  layout = layout_components(ntwk.ig),
+  edge.arrow.mode = 0,
+  vertex.label.color = "black",
+  vertex.shape = "circle",
+  vertex.label.font = 2,
+  main = "Florence nodes by betweenness centrality"
+)
+```
+
+This chart uses heatmap colors to indicate the strength of betweenness centrality.
+The yellow nodes - the network's pendants and isolate - are yellow, which indicates a betweenness measure of 0.
+The orange nodes are nodes with 20 or less optimal paths running through them, and the red nodes have more than 20.
+In fact, the scale had to be forced in this graph to make the smaller-weighted betweenness nodes visible at all compared to the overwhelming dominance of the Medici node, with a betweenness score more than twice that of its nearest rival:
+
+```{r}
+sort(round(betweenness_centrality, 2), decreasing = TRUE)
+```
+
+95 of the `r sum(betweenness_centrality)` optimal paths in the Florence network ran through the Medici.
+It is easy to imagine many different ways that such centrality could manifest as various forms of power.
+
+## Eigenvector and Bonacich Power
+
+Eigenvector centrality is a measure of not just how many alters a node is connected to, but of how well-connected those alters are.
+A node with a high eigenvector centrality score is connected to significant numbers of other highly central nodes.
+For this chart, we can return to standard-area scaling (except for slightly inflating the smallest to keep them visible), and we see that once again, the Medici are in a dominant position, well-connected to the other well-connected Florence families:
+
+```{r}
+eigen_info <- centr_eigen(ntwk.ig, directed = T)
+
+# Calculate node size by 1000 x closeness
+V(ntwk.ig)$size <- eigen_info$vector * 30
+# Ensure minimum node size
+V(ntwk.ig)$size <- pmax(V(ntwk.ig)$size, 5)
+# Base text size on node size
+V(ntwk.ig)$label.cex <- V(ntwk.ig)$size / 15
+V(ntwk.ig)$label.cex <- pmax(V(ntwk.ig)$label.cex, 1)
+
+# colors <- case_when(
+#   betweenness_centrality == 0 ~ "#FFFF33",
+#   betweenness_centrality < 20 ~ "#FF8033",
+#   TRUE ~ "#FF0033"
+# )
+set.seed(1235)
+eigen_labels <- sprintf("%s\n%.2f", ntwk.names, eigen_info$vector)
+
+plot(
+  ntwk.ig,
+  # vertex.color = colors,
+  layout = layout_components(ntwk.ig),
+  vertex.label = eigen_labels,
+  edge.arrow.mode = 0,
+  vertex.label.color = "black",
+  vertex.shape = "circle",
+  vertex.label.font = 2,
+  main = "Florence nodes by eigenvector centrality"
+)
+
+```
+It is interesting to look at the Strozzi family in this diagram, which has the second-highest eigenvector centrality score behind the Medici.
+Recall that in terms of simple degree, the Strozzi and Guadagni families were equal, and in terms of betweenness centrality, the Guadagini family had a significantly higher score than the Strozzi, showing that more optimal / shortest paths between nodes passed through the Guadagini.
+This could be interpreted as a measure of positional influence and power.
+However, here we see that the Strozzi's eigenvector centrality is higher than the Guadagini, and it is plain to see why; the Strozzi are connected to four other families with relatively high eigenvector centrality (the Bischeri, Peruzzi and Castellani), where the Guadagni family is somewhat penalized by its connection to the Lamberteschi pendant node that has no other connections.
+So, although the Guadagni family was more centrally positioned in terms of betweenness, the Strozzi family's fewer optimal paths were nonetheless connected to more central families.
+
+Finally, it is worth examining the Bonacich power connections, which is an interesting measure that in some way inverts the eigenvector concept, proposing that being well-connected to other well-connected nodes does not mean that a node has the ability to *influence* those nodes, as they are also well-connected, and so therefore there is advantage in being connected to weaker nodes.
+When viewed purely in terms of potential power dynamics, this has some interesting merit.
+
+```{r}
+bon_info <- power_centrality(ntwk.ig)
+
+# H/T: range01 <- function(x){(x-min(x))/(max(x)-min(x))}
+range01 <- function(x){(x-min(x))/(max(x)-min(x))}
+scaled_bon_info <- range01(bon_info)
+
+# Calculate node size by 1000 x closeness
+V(ntwk.ig)$size <- scaled_bon_info * 30
+# Ensure minimum node size
+V(ntwk.ig)$size <- pmax(V(ntwk.ig)$size, 5)
+# Base text size on node size
+V(ntwk.ig)$label.cex <- V(ntwk.ig)$size / 15
+V(ntwk.ig)$label.cex <- pmax(V(ntwk.ig)$label.cex, 1)
+
+# colors <- case_when(
+#   betweenness_centrality == 0 ~ "#FFFF33",
+#   betweenness_centrality < 20 ~ "#FF8033",
+#   TRUE ~ "#FF0033"
+# )
+set.seed(1235)
+eigen_labels <- sprintf("%s\n%.2f", ntwk.names, bon_info)
+
+plot(
+  ntwk.ig,
+  # vertex.color = colors,
+  layout = layout_components(ntwk.ig),
+  vertex.label = eigen_labels,
+  edge.arrow.mode = 0,
+  vertex.label.color = "black",
+  vertex.shape = "circle",
+  vertex.label.font = 2,
+  main = "Florence nodes by Bonacich centrality"
+)
+
+```
+
+This offers a very different view of the power dynamics in Florence.
+The Medici are greatly reduced, because their direct connections are all powerful families themselves, and so although they are well-connected by other measures, they may not be in a position of power over those other families.
+More influential families according to Bonacich would include the Ridolfi and Tornabuoni families, who suddenly become dominant from this perspective.
+The Bonacich calculation is complex and includes notions of centrality and power, and is not purely a measure of adjacency; see how small the Albizzi's Bonacich's score is, even though one of its three connections is to the pendant Ginori family, which is nonetheless in a slightly stronger position according to Bonacich.
+A fuller investigation of this dynamic would require some careful research.

### 2022-02-17 _academics/ps753-hw4/ps753-hw4.Rmd
Add toc

@@
     code_folding: true
+    toc: true
 categories:

### 2022-02-17 _academics/ps753-hw4/ps753-hw4.Rmd
Update yaml description

--- 2022-02-17 _academics/ps753-hw4/ps753-hw4.Rmd
@@
 description: |
-  A first-look network analysis
+  Examining centrality measures in the Florentine network
 author:

### 2022-02-19 _posts/2022-02-19-playfair-and-the-flatten-the-curve-graphic/playfair-and-the-flatten-the-curve-graphic.Rmd
new post

---
title: "Playfair and the \"Flatten the Curve\" Graphic"
description: |
  The importance of simplicity in statistical graphics
author:
  - name: Steve Linberg
    url: https://slinberg.net
date: 2022-02-19
output:
  distill::distill_article:
    self_contained: false
categories:
  - tufte
  - design
bibliography: bib.bib
preview: nyt_curve.png
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

One of the early tidbits from [Edward Tufte](https://www.edwardtufte.com)'s "[The Visual Display of Quantitative Information](https://www.edwardtufte.com/tufte/books_vdqi) is this quote from Scottish political economist William Playfair's 1786 book "The Commercial and Political Atlas":

> Information, that is imperfectly acquired, is generally as imperfectly retained; and a man who has carefully investigated a printed table, finds, when done, that he has only a very faint and partial idea of what he has read; and that like a figure imprinted on sand, is soon totally erased and defaced.

This speaks to the general principle, widely accepted now but less so at that time, that tables of data can be hard to interpret and understand, and that knowledge gained from them is fleeting. Deep truths revealed by diligent statistical analysis are of limited value if they can't make themselves readily understood, and if they don't stick in the minds of viewers.

One of the great recent triumphs of visual design was the justly-famous "Flatten the Curve" graphic that illustrated the need to keep the rates of COVID-19 transmission low enough that hospitals could keep up and not be overwhelmed. It actually originated from the CDC in a 2007 publication with a title only a career bureaucrat could love: "Interim Pre-Pandemic Planning Guidance : Community Strategy for Pandemic Influenza Mitigation in the United States : Early, Targeted, Layered Use of Nonpharmaceutical Interventions" [@centersfordiseasecontrolandpreventionu.s.InterimPrepandemicPlanning2007a]:

![The CDC's original "Flatten the Curve" graphic (2007)](r6601a1f1.gif)

Ponderous as the document's title may be, this graphic nonetheless showed this nontrivial concept very well. The solid-colored curve was the bad scenario of an outbreak of a pandemic with no intervention, and the transparent curve was a view of squishing that outbreak down over time so that it would be more manageable.

As brilliant as the design was - far, far more effective than paragraphs of explanatory text with supporting tables of data - it was still a bit text-heavy, with two text axes and five chunks of (small) inline text that had to be read to fully understand its meaning, and similarly-colored curves showing the two different outcomes, so it wasn't quite a glance-level chart that could be instantly understood. When COVID-19 hit in 2019, this graphic was popularized and went through many iterations of further simplification, to makes its message as clear and impactful as possible. The New York Times published this version in 2020 [@robertsFlatteningCoronavirusCurve2020]:

![The NYT's simplified graphic (2020)](nyt_curve.png)

This is even more effective: the text and colors are bold and easy to read, the axis text is simplified, there are only three elements of text in the graph (each only three words), and a third item has been added: a horizontal line indicating the healthcare system's capacity. This is about as simple and direct as it's possible to make this information. It communicated a complex problem in a way that could be instantly understood, with minimal reading, by almost anyone regardless of education levels. It may have saved millions of lives, by being shared around the world and absorbed into the collective human understanding of what needed to happen to make the pandemic as manageable as possible.

(Another very subtle and effective change was the rendering of both curves as slightly irregular in shape, as opposed to the mathematically perfect normal-distribution forms in the CDC's original graphic; this would help to reinforce the notion that infection / hospitalization / death rates from the pandemic would not follow shapes that could be precisely predicted, though overall trends could still be forecast.)

Not every complex social phenomenon can be rendered into such an effective message, but good data science should always strive for this level of simplicity and clarity when the purpose is to communicate something important as broadly and effectively as possible.

### 2022-02-19 _posts/2022-02-17-double-checking-basic-stats/double-checking-basic-stats.Rmd
build

---
title: "Double-Checking Basic Stats"
description: |
  Putting the magnifying glass on page 1 of Tufte's "Visual Display of Quatitative Information".
author:
  - name: Steve Linberg
    url: https://slinberg.net
date: 2022-02-17
output:
  distill::distill_article:
    self_contained: false
    toc: true
    code_folding: true
draft: true
preview: tufte-anscombe-plots.png
categories:
  - tufte
  - r
  - statistics
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(pander)
library(cowplot) # for side-by-side plots from different data
```

Before taking the plunge into [Edward Tufte](https://www.edwardtufte.com)'s "[The Visual Display of Quantitative Information](https://www.edwardtufte.com/tufte/books_vdqi)" , we're going to take a brief side trip into basic statistics to make sure we're on a solid conceptual footing.

Sometimes people make mistakes, even when careful and well-intentioned. Worse, sometimes people use statistics and data to deliberately mislead. Fact-checking is hard, and tedious, and frequently unrewarding. We want to believe what people say and write. We want to look at charts and plots and believe that the underlying data is sound. Most of the time, hopefully, it is. However, part of our role as data scientists is to be ready, willing and able to check the data on any claim, no matter how trustworthy it appears.

Right on page 1 of *VQDI*, Tufte gives us an opportunity to do just this. He shows [Anscombe's Quartet](https://en.wikipedia.org/wiki/Anscombe%27s_quartet), a set of four small sets of data that share nearly identical summary statistics:

![[The Visual Display of Quantitative Information](https://www.edwardtufte.com/tufte/books_vdqi), p1](tufte-anscombe-data.png)

...but that look quite different when plotted, to illustrate the point that visualizing data is a critical aspect of analysis:

![[The Visual Display of Quantitative Information](https://www.edwardtufte.com/tufte/books_vdqi), p2](tufte-anscombe-plots.png)

Anscombe's Quartet is a brilliant project that shows many shortcomings of common statistical methods, and rewards deep study. In our case here, though, we're going to do something far less ambitious: simply verify (using [R](https://www.r-project.org/)) that the summary statistics Tufte shows with the data are accurate, and that we understand what they mean.

My intent for this article is that it should be accessible to non-statisticians, non-mathematicians and non-programmers, though the R program code can be shown with the "Show code" disclosure triangles for those who are interested.

## The data

As luck would have it, the Anscombe dataset is built right into R, in the `datasets` library, so we don't have to retype anything:

```{r}
library(datasets)
datasets::anscombe %>% 
  select(x1, y1, x2, y2, x3, y3, x4, y4) %>%
  pander()
```

<aside>
Careful observers will note that the `x1`, `x2` and `x3` columns are identical, though the `y` columns vary, and that the `x4` column contains the same value for all but the 8th row.
</aside>

For our purpose here, we'll just look at the first data set; naturally, the verification process would be identical for any of the other data sets, or for any other (similar) data set.

```{r}
# extract the first dataset to a separate dataframe, rename columns x and y
set_1 <- datasets::anscombe %>%
  select(x1, y1) %>%
  rename(x = x1, y = y1)
set_1 %>% pander()
```

## The claims

The claim is that all four of the data sets share the following summary statistics:

|Statistic|Value|
|---|---|
|$N$|11|
|mean of $X$'s|9.0|
|mean of $Y$'s|7.5|
|equation of regression line|$Y = 3 + 0.5X$|
|standard error of estimate of slope|0.118|
|$t$|4.24|
|sum of squares $X - \bar{X}$|110.0|
|regression sum of squares|27.50|
|residual sum of squares of Y|13.75|
|correlation coefficient|.82|
|$r^2$|.67|

Let's examine each of the claims in turn.

### $N$, mean of $X$'s and $Y$'s

Simple enough: `nrow()` counts rows, and `mean()` calculates means.

```{r}
set_1.xbar <- mean(set_1$x)
set_1.ybar <- mean(set_1$y)
set_1.n <- nrow(set_1)
```

|Statistic|Value|
|---|---|
|$N$ (row count)|`r set_1.n`|
|mean of x ($\bar{x}$)|`r set_1.xbar`|
|mean of y ($\bar{y}$)|`r set_1.ybar`|

The mean of `y` is slightly higher than `7.5`, but that's a nit we can disregard; in nearly every real-world situation, there is no practical difference between `7.5` and `7.5009091`, and the purpose of the Anscombe data is to show the importance of visualization, not to achieve perfect mathematical precision. Not a lot more needs to be said here; these are the most basic of summary statistics, but nothing is ever so basic that checking it is a waste of time.

### equation of regression line: $Y = 3 + 0.5X$

OK, so what is a regression line? In general terms, it's a line of best fit, a straight line that get as close as possible to all of the points in the data set. If we look at the plot of the first set's points:

```{r}
set_1.baseplot <- ggplot(set_1, mapping = aes(x = x, y = y)) +
  geom_point(color = "darkgreen", size = 3) +
  scale_x_continuous(limits = c(0, 20)) +
  theme_minimal() +
  theme(axis.line = element_line(colour = "darkblue",
                    size = 1, linetype = "solid"))

set_1.baseplot
```

A regression line is a straight line through the graph that best represents, to the extent that a straight line can, the overall shape of the data. It consists of an *intercept* (the value of X where the line crosses the X axis, where Y is 0), and a *slope* (the "rise over the run", or how much each unit of increase in X, as we move left to right, increases or decreases the value / height of Y). It is expressed with the equation

$$y = \beta_0 + \beta_1 \times x$$
where $\beta_0$ ("beta zero") is the intercept, and $\beta_1$ ("beta one") is the slope, or the change in $y$ for every change of one unit in $x$.

Sounds easy enough, but if we did it by eye, trying to draw the line that best shows the middle of the data, how would we do it?

The simplest guess might be to use the basic statistic

We can then add an abline:

```{r}
set_1.baseplot_abline <- set_1.baseplot +
    geom_abline(slope = 0.5, intercept = 3, color = "red")
set_1.baseplot_abline
```

### standard error of estimate of slope = 0.118
### $t$ = 4.24
### sum of squares $X - \bar{X}$ = 110.0
### regression sum of squares = 27.50
### residual sum of squares of Y = 13.75
### correlation coefficient = .82
### $r^2$ = .67

### 2022-02-19 _posts/2022-02-19-playfair-and-the-flatten-the-curve-graphic/playfair-and-the-flatten-the-curve-graphic.Rmd
update thumb size limit, build

@@
     self_contained: false
+bibliography: bib.bib
+preview: nyt_curve.png
 categories:
@@
   - design
-bibliography: bib.bib
-preview: nyt_curve.png
 ---

### 2022-02-19 _posts/2022-02-19-playfair-and-the-flatten-the-curve-graphic/playfair-and-the-flatten-the-curve-graphic.Rmd
typo

@@
 As brilliant as the design was - far, far more effective than paragraphs of explanatory text with supporting tables of data - it was still a bit text-heavy, with two text axes and five chunks of (small) inline text that had to be read to fully understand its meaning, and similarly-colored curves showing the two different outcomes, so it wasn't quite a glance-level chart that could be instantly understood.
-When COVID-19 hit in 2019, this graphic was popularized and went through many iterations of further simplification, to makes its message as clear and impactful as possible.
+When COVID-19 hit in 2019, this graphic was popularized and went through many iterations of further simplification, to make its message as clear and impactful as possible.
 The New York Times published this version in 2020 [@robertsFlatteningCoronavirusCurve2020]:

### 2022-02-19 _posts/2022-02-19-playfair-and-the-flatten-the-curve-graphic/playfair-and-the-flatten-the-curve-graphic.Rmd
edits
build

@@
 
-One of the early tidbits from [Edward Tufte](https://www.edwardtufte.com)'s "[The Visual Display of Quantitative Information](https://www.edwardtufte.com/tufte/books_vdqi) is this quote from Scottish political economist William Playfair's 1786 book "The Commercial and Political Atlas":
+One of the early tidbits from [Edward Tufte](https://www.edwardtufte.com)'s "[The Visual Display of Quantitative Information](https://www.edwardtufte.com/tufte/books_vdqi)" is this quote from Scottish political economist William Playfair's 1786 book "The Commercial and Political Atlas":
 
@@
 This is about as simple and direct as it's possible to make this information.
-It communicated a complex problem in a way that could be instantly understood, with minimal reading, by almost anyone regardless of education levels.
+It communicated a complex problem in a way that could be instantly understood, by almost anyone, regardless of education levels.
 It may have saved millions of lives, by being shared around the world and absorbed into the collective human understanding of what needed to happen to make the pandemic as manageable as possible.

### 2022-02-19 _posts/2022-02-19-playfair-and-the-flatten-the-curve-graphic/playfair-and-the-flatten-the-curve-graphic.Rmd
build

@@
 
-One of the early tidbits from [Edward Tufte](https://www.edwardtufte.com)'s "[The Visual Display of Quantitative Information](https://www.edwardtufte.com/tufte/books_vdqi)" is this quote from Scottish political economist William Playfair's 1786 book "The Commercial and Political Atlas":
+One of the good early tidbits from [Edward Tufte](https://www.edwardtufte.com)'s "[The Visual Display of Quantitative Information](https://www.edwardtufte.com/tufte/books_vdqi)" is this quote from Scottish political economist William Playfair's 1786 book "The Commercial and Political Atlas":
 

### 2022-02-19 _academics/ms797-glossary/ms797-glossary.Rmd
New terms, knit

@@
 : "the ratio of the variance of $\hat{\beta_j}$ when fitting the full model divided by the variance of $\hat{\beta_j}$ on its own"; smallest possible value of 1 indicates the absence of collinearity, 5-10 indicates a "problematic amount".
-Hairy forumla not reproduced.
+$$\text{VIF}(\hat{\beta_j}) = \frac{1}{1 - {R_{X_j|X_-j}^2}}$$
+"$R_{X_j|X_-j}^2$ is the $R^2$ from a regression of $X_j$ onto all of the other predictors"
+
+K-nearest neighbors regression (105)
+: a mode of regression that seeks to classify observations by their proximity to classified neighbors
+
+curse of dimensionality (107)
+: when an observation has no nearby neighbors due to a high number of dimensions exponentially increasing the available space for other observations to be spread out in
+
+# Chapter 4
+
+qualitative / categorical (129)
+: interchangeable term for a variable with a non-quantitative value, such as color
+
+classification (129)
+: predicting a qualitative response for an observation
+
+classifier (129)
+: a classification technique, such as: logistic regression, linear discriminant analysis, quadratic discriminant analysis, naive Bayes, and K-nearest neighbors
+
+logistic function (134)
+: a specific function returning an S-shaped curve with values between 0 and 1, used in logistic regression
+$$p(X) = \frac{e^{\beta_0 + \beta_1X}}{1 + e^{\beta_0 + \beta_1X}}$$
+
+odds (134)
+: the likelihood of a particular outcome: the ratio of the number of results that produce the outcome versus the number that do not, between 0 and $\infty$ (very low or very high probabilities)
+$$odds = \frac{p(X)}{1 - p(X)} = e^{\beta_0 + \beta_1X}$$
+Note: odds is not the same as probability! Odds of 1/4 means a 20% probability, not 25%.
+
+To convert from odds to probability: divide odds by (1 + odds), as in:
+
+$$\frac{1}{4} \div \left[ 1 + \frac{1}{4} \right] = \frac{1}{4} \div \frac{5}{4} = 1/5 = 0.2$$
+
+To convert from probability to odds, divide probability by (1 - probability), as in:
+$$\frac{1}{5} \div \left[ 1 - \frac{1}{5} \right] = \frac{1}{5} \div \frac{4}{5} = \frac{1}{5} \times \frac{5}{4} = 1/4 = 0.25$$
+log odds / logit (135)
+: the log of the odds 
+$$\text{log} \left( \frac{p(X)}{1 - p(X)} \right)$$
+
+likelihood function (135)
+: a function that produces the closest possible match of a set of observations, for use in predicting the classifications of other points
+hairy equation not rendered
+
+confounding (139)
+: when predictors correlate; usually this is something to work to avoid
+
+multinomial logistic regression (140)
+: logistic regression into more than 2 categories
+
+softmax coding (141)
+: an alternative coding for multiple logistic regression that treats all classes symmetrically, rather than establishing one as a baseline
+
+(probability) density function (142)
+: a function returning the likelihood of a particular value occurring in a given sample, between 0 and 1; often used in pairs to establish likelihood of a value within a range rather than the likelihood of an infinitely-thin single "slice"
+
+prior (142)
+: the probability that a random observation comes from class $k$
+
+posterior (142)
+: the probability that an observation belongs to a class $k$ given its predictor value
+
+normal / Gaussian (143)
+: characterized by a bell-shaped curve, not uniformly distributed but tending towards a likeliest/central value
+
+overfitting (148)
+: mapping too closely to idiosyncrasies in training data, increasing variance in a model
+
+null classifier (148)
+: a classifier that always predicts a zero or null status
+
+confusion matrix (148)
+: a matrix showing how many predictions were made, and how accurate the classifications were (predicted x actual, hits on TL-BR diagonal and misses on BL-TR diagonal)
+
+sensitivity (149)
+: the percentage of a class (like defaulters) that is correctly identified
+
+specificity (149)
+: the percentage of a different class (like non-defaulters) that is correctly identified
+
+ROC curve (150)
+: Receiver Operating Characteristics; name of a curve showing the overall performance of a classifier resembling the top left corner of a rounded rectangle; area under the curve (AUC) indicates the percentage of correct classifications; the closer it is to square, the bigger the AUC, the better the classifier
+
+marginal distribution (155)
+: the distribution of an individual predictor
+
+joint distribution (155)
+: the association between different predictors
+
+kernel density estimator (156)
+: "essentially a smoothed version of a histogram"
+
+# Chapter 5
+
+model assessment (197)
+: the process of evaluating a model's performance
+
+model selection (197)
+: the process of selecting the proper level of flexibility for a model
+
+validation set approach (198)
+: randomly dividing a set of observations into a training set and a validation or hold-out set, to assess the test error rate
+
+leave one out cross-validation (LOOCV) (200)
+: using a single observation for a validation set and the rest as a training set
+
+k-fold cross-validation (CV) (203)
+: dividing a set of observation into $k$ groups of approximately equal size, using the first as a validation set and fitting the method on the remaining $k-1$ folds
+
+bootstrap (209)
+: a tool to quantify the uncertainty associated with a given estimator or statistical learning method
+
+sampling with replacement (211)
+: picking from a set without removing the picked elements

### 2022-02-21 _academics/2022-02-19-ms797-bigger-concepts/ms797-bigger-concepts.Rmd
Initial commit

---
title: "MS797 Bigger Concepts"
description: |
  Bigger concepts for MS797 coursework
author:
  - name: Steve Linberg
    url: {}
date: 2022-02-19
output:
 distill::distill_article:
   self_contained: false
   toc: true
   code_folding: Show code
categories:
  - machine learning
---

<p style="background: #eee">
[<i class="fas fa-home"></i>](/) &gt; Academics &gt; [MS797](../ms797/index.html)
&gt; `r rmarkdown::metadata$title`
</p>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ISLR2)
library(tidyverse)
library(GGally) # for ggpairs
library(pander) # for prettier output
library(cowplot) # for side-by-side plots from different data
```

This page is for concepts that don't fit into the [glossary](../ms797-glossary/index.html).

# Classification

The general umbrella categories for *classification*&mdash;where the outcome variable to be predicted is categorical, rather than continuous, like eye color rather than weight&mdash;are *discriminative" and "generative".

Logistic regression is discriminative. Linear discriminant analysis, quadratic discriminant analysis, and naive Bayes are generative, K-nearest neighbors appears to be neither.

## Logistic Regression

Logistic regression's output is a *probability* that a response variable belongs to a particular category.

The general form of a logistic model is:

$$p(X) = \text{PR}(Y = 1|X)$$
The right side of the equation reads "The probability that Y equals 1, given X" - or the chance that Y is a particular category for a specific value of X.

A linear regression model would use this form:

$$p(X) = \beta_0 + \beta_1X$$

And the magic formula for this is the *logistic function*, in which we raise $e$ to this power.

$$p(X) = \frac{e^{\beta_0 + \beta_1X}}{1 + e^{\beta_0 + \beta_1X}}$$

(Why $e^{\beta_0 + \beta_1X}$ instead of, say, $2^{\beta_0 + \beta_1X}$ or $\pi^{\beta_0 + \beta_1X}$ or $123456^{\beta_0 + \beta_1X}$?)

Dividing both sides by $1 - p(X)$ (somehow) yields:

$$\frac{p(X)}{1 - p(X)} = e^{\beta_0 + \beta_1X}$$

which is also the *odds* that Y is 1 given X. If the probability of something is 20%, or 1/5, then the odds are 1/4:

$$\frac{1}{5} \div \left[ 1 - \frac{1}{5} \right] = \frac{1}{5} \div \frac{4}{5} = \frac{1}{5} \times \frac{5}{4} = 1/4 = 0.25$$

### When to use logistic regression

When the response variable is categorical, and has two possible classes, like "Yes" or "No", "Up" or "Down", etc. However, see Linear Discriminant Analysis below for examples of when *not* to use it.

In some cases, a categorical response variable could be analyzed with a linear regression by treating the categorical variable as numeric, like treating "not infected" as 0 and "infected" as 1, and using the rounded value as the predictor; however, beyond 2 possible responses it's unlikely that a continuous scale could be established between the possible values, and it gets messy.

### How to do logistic regression

Use `glm()` with the argument `family = binomial` to indicate logistic regression.

Using the `Boston` data set from the `ISLR2` package, the following will model the probability of a tract having a `crim` rate above the median, based on `zn`, using logistic regression:

```{r}
data(Boston)
boston <- Boston %>%
  # Create the crim_above_median response variable
  mutate(crim_above_med = as.factor(
    ifelse(crim > median(crim), "Yes", "No")))

# Fit the model
zn.fit <- glm(crim_above_med ~ zn, data = boston, family = binomial)
summary(zn.fit)
```

The `foo.fit` summary will show the coefficients and p-values; the `$coef` aspect of the summary will show more detail.

`predict()` can then be fed the fit to make predictions. The argument `type = "response"` will tell it to output probabilities "of the form $P(Y = 1|X)$" from the data that was used in the fit, also described as the training data. It returns a vector of probabilities for every data element.

```{r}
# Create the probability vector from the fit
zn.probs <- predict(zn.fit, type = "response")
head(zn.probs) %>% pander()
```

The responses can be converted into categories and turned into a confusion matrix to show how the regression did with the training data.

```{r}
# Convert the probability vector into yes-or-no predictions
zn.pred <- ifelse(zn.probs > 0.5, "Yes", "No")
# Display a confusion matrix
table(zn.pred, boston$crim_above_med) %>% pander()
```

The overall success rate can be calculated with `mean()`, comparing each predicted value to the actual value:

```{r}
# See how many of the predictions match the actual value
mean(zn.pred == boston$crim_above_med)
```

The model can then be run on test data that is held out from the main set, with the expectation that it won't do quite as well as the training data. The efficacy can then be assessed in the same way, and if it's not high, we can return to the model and try to find other forms that make better predictions - or perhaps conclude that we can't make a good prediction at all.

## Linear Discriminant Analysis (LDA)

The simplest use case for LDA is when there are more than two possible response classes; logistic regression gives the probability that an observation is in a class A or B, but it can only handle two classes.

LDA also gives better results when "there is substantial separation between the two classes", or where the sample size is small and the distribution of the predictors is approximately normal in each of the classes.

<aside>
How do we determine this?
</aside>

<!--
### When to use it

### How to do it

### How to interpret it
-->

### 2022-02-25 steve.Rmd
update photo

@@
 title: "Steve Linberg"
-image: "2021-sfl-photo-fadedbg-256.png"
+image: "2022-sfl-photo-400x400.jpg"
 links:

### 2022-02-25 _academics/ms797-bigger-concepts/ms797-bigger-concepts.Rmd
build

--- 2022-02-21 _academics/2022-02-19-ms797-bigger-concepts/ms797-bigger-concepts.Rmd
@@
 
+**This document is in process.**
+
 # Classification
 
-The general umbrella categories for *classification*&mdash;where the outcome variable to be predicted is categorical, rather than continuous, like eye color rather than weight&mdash;are *discriminative" and "generative".
+The general umbrella categories for *classification*&mdash;where the outcome variable to be predicted is categorical, rather than continuous, like eye color rather than weight&mdash;are "discriminative" and "generative".
+Discriminative models take sets of data and separate them into classes, like finding the decision boundary between two or more classes, but can't predict the values of new data.
+Generative models (used in unsupervised learning) can generate new data, but are more susceptible to outliers (misclassified observations) than discriminative models, require more data to create, and are more computationally expensive.
+
+A generative model&mdash;somehow&mdash;tries to guess how the data was generated, so it can take a new bit of unseen data, assess the probability that each class would have produced it, and pick the class with the highest probability.
+
+A discriminative model can only look at existing data and create a decision boundary separating the existing data into classes.
+
+![Figure from https://stanford.edu/%7Eshervine/teaching/cs-229/cheatsheet-supervised-learning](disc-gen.png)
+
+Note that there is no decision boundary in the generative model, just levels of probability into which new data could be classified with varying levels of confidence.
 
@@
 
-The `foo.fit` summary will show the coefficients and p-values; the `$coef` aspect of the summary will show more detail.
+The `zn.fit` summary above will show the coefficients and p-values; the `$coef` aspect of the summary will show more detail.
 

### 2022-02-28 _academics/ms797-glossary/ms797-glossary.Rmd
knit and clean build

@@
 : _"the amount by which $\hat{f}$ would change if we estimated it using a different training data set"_
+More practically: the *average* of squared differences from the mean, often expressed as $\sigma^2$, where $\sigma$ (or the square root of the variance) is the standard deviation
 

### 2022-02-28 _academics/ps753-glossary/ps753-glossary.Rmd
Add terms

@@
 That is, being connected to other central nodes increases centrality." Takes into account the centrality of the nodes a node is connected to, considering the importance of the connected nodes and not just their quantity or path length.
+A node with high eigenvector centrality is connected to significant numbers of other highly central nodes.
 $$\lambda \text{c} = \text{Wc}$$
@@
 Bonacich Power
-: "A closely related concept, but includes a weighting factor that emphasizes global vs. local structure (and negative connections)"
+: "A closely related concept, but includes a weighting factor that emphasizes global vs.
+local structure (and negative connections)" In contrast to eigenvector centrality, penalizes nodes connected to other well-connected nodes, under the theory that being connected to strong or powerful nodes means a node has less influence and power than it would if it were connected to fewer or weaker nodes.
 $$\text{c}(\alpha, \beta) = \alpha(\text{I} - \beta\text{W})^{-1}\text{W1}$$
@@
 betweenness centrality
-: (forthcoming)
+: a count of the number of shortest paths between nodes that pass through another node

### 2022-03-05 _posts/2022-03-05-tufte-takes-no-prisoners/tufte-takes-no-prisoners.Rmd
new post

---
title: "Tufte Takes No Prisoners"
description: |
  Edward Tufte gets raw on chartjunk.
author:
  - name: Steve Linberg
    url: https://slinberg.net
date: 2022-03-05
output:
  distill::distill_article:
    self_contained: false
categories:
  - tufte
  - design
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

The gloves come off in in "Chartjunk", chapter 5 of [Edward Tufte](https://www.edwardtufte.com)'s "[The Visual Display of Quantitative Information](https://www.edwardtufte.com/tufte/books_vdqi)". Here I am, innocently beginning to think about a poster presentation I have coming up in a couple of months and to imagine design elements that will beautify the project and help make the information easier to understand, and then right in the first paragraph, Tufte drops this:

> Graphical decoration, which propers in technical publications as well as in commercial and media graphics, comes cheaper than the hard work required to produce intriguing numbers and secure evidence.

*...ouch.*

OK, let's not overreact: he's not saying "don't strive for beauty" here. The main point is that visual clutter in charts and graphs, which can take many forms, at best pollutes the message and obfuscates the data, and at worst might be used to (attempt to) conceal a lack of data. That's fair. He then goes on to show many examples of charts rendered almost unintelligible by over-use of fill patterns, made simple by modern computer-generated graphing systems, pointless use of three-dimensional effects that don't represent data, and heavy gridlines that visually compete with data points.

On the subject of [graph paper](#graph_paper), he doesn't hold back:

> Most ready-made graph paper comes with a darkly printed grid. The reverse (unprinted) side should be used, for then the lines show through faintly and do not clutter the data. If the paper is heavily gridded on both sides, throw it out.

Taking no prisoners, he shows this monstrosity a little further on:

```{r fig.cap="Figure from [The Visual Display of Quantitative Information](https://www.edwardtufte.com/tufte/books_vdqi), p. 118", out.width=400}
knitr::include_graphics("badgraphic.png")
```

...introduced by this utter decimation:

> A series of weird three-dimensional displays appearing in the magazine *American Education* in the 1970s delighted connoisseurs of the graphically preposterous. Here five colors report, almost by happenstance, only five pieces of data (since the division within each year adds to 100 percent). This may well be the worst graphic ever to find its way into print.

On first read, I thought to myself: "*I* want to be a connoisseur of the graphically preposterous". Then I thought it over a bit more and decided I wasn't so sure.

In any event, I'm around the 2/3 point now in VDQI, enjoying many savagely entertaining examples of what *not* to do in the visual presentation of data. There are a few good examples scattered in as well, though with much lower overall frequency so far. Studying the wreckage of bad design can absolutely be educational, but with all due deference to the important work of learning from the mistakes of others, it is always easier to critique than to create. I am not "peeking ahead" out of sequence to see if he does this more in subsequent chapters; *Saturday Mornings with Tufte&trade;* are deliberately slow and meditative, with a cup of coffee and no electronic devices or distractions. So far this has been an informative and educational read, and I'm looking forward to the rest of it and then to moving on to Beautiful Evidence. But I'm still planning to try to create something beautiful in my upcoming poster, and hopefully I can find a way to do that that would also dodge the well-deserved barbs aimed at the many failed efforts shown in this book.

***
(*<span id="graph_paper">Endnote</span>*: I am a big fan of graph paper, and although most of my note-taking these days is on the ipad, I still have a big physical stash of my favorite, [whitelines](https://www.whitelinespaper.com/) (non-referral link):

![From one of my [whitelines](https://www.whitelinespaper.com/) notebooks](whitelines.jpg)

It uses a light gray tone with white gridlines instead of black, *much* easier on the eyes.)

### 2022-03-05 _posts/2022-03-05-tufte-takes-no-prisoners/tufte-takes-no-prisoners.Rmd
linked BE

@@
 I am not "peeking ahead" out of sequence to see if he does this more in subsequent chapters; *Saturday Mornings with Tufte&trade;* are deliberately slow and meditative, with a cup of coffee and no electronic devices or distractions.
-So far this has been an informative and educational read, and I'm looking forward to the rest of it and then to moving on to Beautiful Evidence.
+So far this has been an informative and educational read, and I'm looking forward to the rest of it and then to moving on to "[Beautiful Evidence](https://www.edwardtufte.com/tufte/books_be)".
 But I'm still planning to try to create something beautiful in my upcoming poster, and hopefully I can find a way to do that that would also dodge the well-deserved barbs aimed at the many failed efforts shown in this book.

### 2022-03-05 _academics/ms797-bigger-concepts/ms797-bigger-concepts.Rmd
build

@@
 
+## Resampling methods
+
+Resampling methods are used to refine and test models against training data by repeatedly testing against different subsets within the data.
+
+*Cross-validation* can be used to estimate the test error rate of a method (*model assessment*), and *bootstrap* is mostly to measure the accuracy of a parameter or of a method.
+
+### Cross-validation: validation set approach
+
+This approach randomly splits the data into two sets: a *training* set and a *validation* or *test* set.
+Size ratios seem to be on the order of between 50-50 to 3:1 or 4:1 training-to-test data.
+The model is fit against the training data, and then `predict()` is used to predict the results for the validation data; the results of `predict` can then be compared against the actual classifications in the validation data to compute the error rate.
+
+This process can be run repeatedly with different splits.
+
+The general process is:
+
+1. Randomly split the dataset into training and validation sets
+2. Fit the model to the training data
+3. Predict the outcomes of the model on the validation data
+4. Convert the predictions into categorical variables that match the field being predicted (like Yes/No, etc)
+5. The error rate is the proportion of correct predictions in the validation data.
+
+The code below runs 10 iterations and accumulates the results into a results data frame.
+
+```{r eval=FALSE}
+set.seed(1235)
+train_pct <- .75
+
+p5_3a.results <- data.frame(iteration = integer(),
+                           error_rate = double())
+
+for (i in 1:10) {
+  p5_3a.training <- rep(FALSE, nrow(Default))
+  p5_3a.training[sample(nrow(Default), nrow(Default) * train_pct)] <- TRUE
+  p5_3a.validation <- !p5_3a.training
+
+  p5_3a.fit <-
+    glm(
+      formula = default ~ income + balance,
+      family = binomial,
+      data = Default,
+      subset = p5_3a.training
+    )
+
+  p5_3a.probs <- predict(p5_3a.fit, Default[p5_3a.validation, ], type = "response")
+  p5_3a.pred <- ifelse(p5_3a.probs > 0.5, "Yes", "No")
+  p5_3a.validation_error_rate <-
+    mean(p5_3a.pred == Default[p5_3a.validation, ]$default)
+  p5_3a.validation_error_rate_print <-
+    round(p5_3a.validation_error_rate, 2)
+
+  p5_3a.results[i,] <- data.frame(
+    iteration = i,
+    error_rate = 1 - p5_3a.validation_error_rate
+  )
+}
+
+```
+
+### Cross-validation: leave-one-out (LOOCV)
+
+This method is similar to the validation set approach, except that the validation set is only one element, and the training set is the remaining $n-1$ observations.
+It is run automatically $n$ times and returns an average of the test errors.
+
+The process is simpler than the validation set approach:
+
+1. Fit the model to the entire data set with `glm()` (omitting the `subset` argument)
+2. Run `cv.glm` on the fit
+3. The `delta` factor of the return data is the cross-validation error rate for the test data.
+
+Code from the book:
+
+```{r eval=FALSE}
+library(boot)
+glm.fit <- glm(mpg ~ horsepower, data = Auto)
+cv.error <- cv.glm(Auto, glm.fit)
+cv.error$delta
+```
+
+### Cross-validation: k-fold
+
+K-fold cross-validation is like leave-one-out cross-validation but with a larger test set; it splits the data up into $k$ segments and runs the cross-validation on each.
+It may give better results if run repeatedly and averaged, as with the validation set approach.
+It is faster than LOOCV, obviously, because it uses less data.
+
+The process is the same as LOOCV, except a parameter such as `K = 10` is added to the `cv.glm()` call. 5 to 10 is a common range for values of $k$.
+
 <!--

### 2022-03-20 _academics/ps753-hw5/ps753-hw5.Rmd
New post

---
title: "Homework 5: Brokerage and Betweenness in the Florentine Wedding dataset"
description: |
  Examining additional centrality measures in the Florentine network
author:
  - name: Steve Linberg
    url: {}
date: 2022-02-16
output:
  distill::distill_article:
    self_contained: false
    code_folding: true
    toc: true
categories:
  - r
  - social networks
---

<p style="background: #eee">
[<i class="fas fa-home"></i>](/) &gt; Academics &gt; [PS753](../ps753/index.html)
&gt; `r rmarkdown::metadata$title`
</p>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	fig.height = 8,
	message = FALSE,
	warning = FALSE
)
library(network)
library(igraph)
library(tidyverse)
```

### 2022-03-20 _academics/ms797-glossary/ms797-glossary.Rmd
build

--- 2022-02-19 _academics/ms797-glossary/ms797-glossary.Rmd
@@
 : _"the amount by which $\hat{f}$ would change if we estimated it using a different training data set"_
+More practically: the *average* of squared differences from the mean, often expressed as $\sigma^2$, where $\sigma$ (or the square root of the variance) is the standard deviation
+per [StatQuest](https://www.youtube.com/watch?v=EuBBz3bI-aA&t=246s): "the difference in fits between data sets" (like training and test)
 
@@
 More flexible models increase variance and decrease bias.
+per [StatQuest](https://www.youtube.com/watch?v=EuBBz3bI-aA&t=136s): "The inability for a machine learning method (like linear regression) to capture the true relationship is called **bias**" - a straight line trying to model a curved separation in classes will never get it right and always be biased
+
+bias (65)
+: in an estimator, something that systematically misses the true parameter; for an unbiased estimator, $\hat{\mu} = \mu$ when averaged over (huge) numbers of observations
 
@@
 : the difference between $y_i$ and $\hat{y}_i$, also $e_i$; the difference between the $i$th response variable and the $i$th response variable predicted by the model
+[StatQuest](https://youtu.be/g9c66TUylZ4?t=542): the difference between the *observed* and *predicted* value
 
@@
 : the regression line made of the least-squares estimates for $\beta_0$ and $\beta_1$
-
-bias (65)
-: in an estimator, something that systematically misses the true parameter; for an unbiased estimator, $\hat{\mu} = \mu$ when averaged over (huge) numbers of observations
 
@@
 
+Note: sensitivity and specificity can and should be separately considered when evaluating ML methods.
+Each column (class) of a confusion matrix has a sensitivity and a specificity.
+In a 2x2, this is simple; in larger than 2x2, it involves summation.
+
+$$\text{Sensitivity} = \frac{\text{True Positives}}{\text{True Positives + False Negatives}}$$
+True positives is the cell in the class column containing the number of correct predictions in a category; the false negatives are the rest of the column.
+
+$$\text{Specificity} = \frac{\text{True Negatives}}{\text{True Negatives + False Positives}}$$
+True negatives is the sum of cells, in the other columns, that correctly do *not* predict the class (even if they wrongly predict some other class as well); the false positives are the rest of the *row* for the class that incorrectly predicted that class.
+
+ROC charts show the true positive rate (sensitivity) on the Y axis, and the false positive rate (1 - sensitivity) on the X axis.
+
 ROC curve (150)
-: Receiver Operating Characteristics; name of a curve showing the overall performance of a classifier resembling the top left corner of a rounded rectangle; area under the curve (AUC) indicates the percentage of correct classifications; the closer it is to square, the bigger the AUC, the better the classifier
+: Receiver Operating Characteristics; name of a curve showing the overall performance of a classifier resembling the top left corner of a rounded rectangle; area under the curve (AUC) indicates the percentage of correct classifications; the closer it is to square, the bigger the AUC, the better the classifier.
+The Y axis is true positive rate (sensitivity).
+the X axis is false positive rate (1 - sensitivity).
+
+[StatQuest](https://www.youtube.com/watch?v=4jRBRDbJemM&list=PLblh5JKOoLUICTaGLRoHQDuF_7q2GfuJF&index=6): AUC (area under the curve) of the ROC shows the overall effectiveness / quality of a model; good models will fill the space more.
+And the points of the individual curve help indicate the best thresholds to pick for the classifier; points with X = 0 have no false positives, and points with Y = 1 get all of the true positives.
+
+Also, *precision* is another option for the X-axis, rather than false positives, if it's more important to know the proportion of the true positives were correctly identified.
+
+$$\text{Precision} = \frac{\text{True Positives}}{\text{True Positives + False Positives}}$$
 

### 2022-03-20 _academics/ps753-glossary/ps753-glossary.Rmd
build

@@
 two-mode data / bipartite data
-: ties between different sets (or categories) of agents, such as ties between nations and international organizations; a network split into two parts, each of whose nodes only connect to nodes in the opposite part, not to nodes within its own part
+: ties between different sets (or categories) of agents, such as ties between nations and international organizations; a network split into two parts, each of whose nodes only connect to nodes in the opposite part, not to nodes within its own part; "affiliation network" in @wassermanSocialNetworkAnalysis1994
 

### 2022-03-20 _academics/ps753-hw4/ps753-hw4.Rmd
build

@@
 More influential families according to Bonacich would include the Ridolfi and Tornabuoni families, who suddenly become dominant from this perspective.
-The Bonacich calculation is complex and includes notions of centrality and power, and is not purely a measure of adjacency; see how small the Albizzi's Bonacich's score is, even though one of its three connections is to the pendant Ginori family, which is nonetheless in a slightly stronger position according to Bonacich.
+The Bonacich calculation is complex and includes notions of centrality and power, and is not purely a measure of adjacency; see how small the Albizzi's Bonacich score is, even though one of its three connections is to the pendant Ginori family, which is nonetheless in a slightly stronger position according to Bonacich.
 A fuller investigation of this dynamic would require some careful research.

### 2022-03-20 _academics/ps753-useful-r-tidbits/ps753-useful-r-tidbits.Rmd
build

@@
 This document contains R snippets pertaining to work in the [Social Networks class](../ps753/index.html). There is also one for the [Machine Learning class](../ms797-useful-r-tidbits/index.html).
+
+## Prelude: finding data to play with
+
+Since in the class tutorials, the network objects are pre-loaded in the web-based runtime environments, we don't actually see how to load some of the data sets that the exercises are based on, for running outside of this environment.
+
+There are gazillions of ways to get data into R, but for the purposes of playing with networks, there are some packages with pre-existing network data that can be easily loaded.
+Two of these are `igraphdata`, which contains igraph-formatted network datasets, and `ergm`, which contains statnet-compatible network datasets.
+
+The datasets in each can be listed with the `data` command:
+
+```r
+> data(package = "igraphdata")
+
+Data sets in package ‘igraphdata’:
+
+Koenigsberg              Bridges of Koenigsberg from Euler's times
+UKfaculty                Friendship network of a UK university faculty
+USairports               US airport network, 2010 December
+enron                    Enron Email Network
+foodwebs                 A collection of food webs
+immuno                   Immunoglobulin interaction network
+karate                   Zachary's karate club network
+kite                     Krackhardt's kite
+macaque                  Visuotactile brain areas and connections
+rfid                     Hospital encounter network data
+yeast                    Yeast protein interaction network
+
+> data(package = "ergm")
+
+Data sets in package ‘ergm’:
+
+cohab_MixMat (cohab)     Target statistics and model fit to a hypothetical
+                         50,000-node network population with 50,000 nodes
+                         based on egocent
+cohab_PopWts (cohab)     Target statistics and model fit to a hypothetical
+                         50,000-node network population with 50,000 nodes
+                         based on egocent
+cohab_TargetStats (cohab)
+                         Target statistics and model fit to a hypothetical
+                         50,000-node network population with 50,000 nodes
+                         based on egocent
+ecoli1 (ecoli)           Two versions of an E. Coli network dataset
+ecoli2 (ecoli)           Two versions of an E. Coli network dataset
+faux.desert.high         Faux desert High School as a network object
+faux.dixon.high          Faux dixon High School as a network object
+faux.magnolia.high       Goodreau's Faux Magnolia High School as a network
+                         object
+faux.mesa.high           Goodreau's Faux Mesa High School as a network
+                         object
+flobusiness (florentine)
+                         Florentine Family Marriage and Business Ties Data
+                         as a "network" object
+flomarriage (florentine)
+                         Florentine Family Marriage and Business Ties Data
+                         as a "network" object
+g4                       Goodreau's four node network as a "network"
+                         object
+kapferer                 Kapferer's tailor shop data
+kapferer2 (kapferer)     Kapferer's tailor shop data
+molecule                 Synthetic network with 20 nodes and 28 edges
+samplike (sampson)       Cumulative network of positive affection within a
+                         monastery as a "network" object
+samplk1 (samplk)         Longitudinal networks of positive affection
+                         within a monastery as a "network" object
+samplk2 (samplk)         Longitudinal networks of positive affection
+                         within a monastery as a "network" object
+samplk3 (samplk)         Longitudinal networks of positive affection
+                         within a monastery as a "network" object
+```
+
+To load the statnet "florentine" data from the `ergm` package:
+
+```r
+> data("florentine", package = "ergm")
+```
+
+This creates the objects `flobusiness` and `flomarriage`, both from this dataset, ready to go in statnet. Note that another version of this dataset is found in the `network` package `flo`:
+
+```r
+data("flo", package = "network")
+```
+
+This loads the very simple data into a matrix, but it is not yet a network object. To convert it to an igraph object:
+
+```r
+library(network)
+library(igraph)
+
+data(flo)
+ntwk.ig <- graph_from_adjacency_matrix(flo)
+```
 
@@
 karate.stat %e% "weight"
+```
+
+### Getting the basic network descriptors
+
+First look at any network is to examine the network *size*, *type* (un/directed, un/weighted, bipartite) and available *attributes of vertices and edges*.
+
+#### igraph
+
+Basic descriptors first:
+
+```r
+vcount(climpref.ig)
+[1] 34
+
+ecount(climpref.ig)
+[1] 531
+
+is_bipartite(climpref.ig)
+[1] FALSE
+
+is_directed(climpref.ig)
+[1] FALSE
+
+is_weighted(climpref.ig)
+[1] TRUE
+```
+
+Listing vertex and edge attributes:
+
+```r
+> vertex_attr_names(climpref.ig)
+[1] "name"                 "Climate.council"      "Klimaallianz"        
+[4] "Stiftung.Klimarappen" "type3"                "type5"               
+[7] "dm"
+
+> edge_attr_names(climpref.ig)
+[1] "weight"
+
+```
+
+Accessing vertex and edge attributes:
+
+```r
+> V(climpref.ig)$name %>% head()
+[1] "AA" "AB" "AC" "AD" "AE" "AF"
+
+E(climpref.ig)$weight %>% head()
+[1] 0.69 0.08 0.04 0.18 0.42 0.21
+```
+
+#### statnet
+
+The basic descriptors in statnet are all *shown* by print():
+
+```r
+> print(flobusiness)
+ Network attributes:
+  vertices = 16 
+  directed = FALSE 
+  hyper = FALSE 
+  loops = FALSE 
+  multiple = FALSE 
+  bipartite = FALSE 
+  total edges= 15 
+    missing edges= 0 
+    non-missing edges= 15
+
+ Vertex attribute names: 
+    priorates totalties vertex.names wealth
+
+No edge attributes
+```
+
+It appears, though it's thinly documented, that these attributes are programmatically accessible through the `$gal` attribute, as in `is_directed <- flobusiness$gal$directed`:
+
+```r
+> flobusiness$gal
+$n
+[1] 16
+
+$mnext
+[1] 16
+
+$directed
+[1] FALSE
+
+$hyper
+[1] FALSE
+
+$loops
+[1] FALSE
+
+$multiple
+[1] FALSE
+
+$bipartite
+[1] FALSE
 ```
@@
 
-## Transitivity
+### Transitivity
 
@@
 isolates(gotbook.stat)
+
 # There is no direct command in igraph, but we can do this:
@@
 deg_counts <- igraph::degree(gotbook.ig, loops = FALSE)
+
 # filter and count the nodes with 0 degrees (or any other quantity of interest)
@@
 # subset vertex.names attribute to get names of isolates
-as.vector(gotbook.stat %v% 'vertex.names')[c(isolates(gotbook.stat))]
+as.vector(gotbook.stat %v% 'vertex.names')[c(isolates(gotbook.stat))] %>%
+  head()
+##   [1] "Aegon Frey (Jinglebell)"         "Alebelly"                       
+##   [3] "Alfyn"                           "Allar Deem"                     
+##   [5] "Antario Jast"                    "Balman Byrch"    
 ```
@@
 
-Again, there appear to be some inconsistency between `igraph` and `statnet` in the calculations, with statnet apparently not incorporating weights and failing on singluar matrices.
+Again, there appear to be some inconsistency between `igraph` and `statnet` in the calculations, with statnet apparently not incorporating weights and failing on singular matrices.
 
@@
 ```
+
+## Week 5: Big Block of Basic Code
+
+Big Blocks of Basic Code to get a bunch of measures of a network:
+
+```r
+# Get the basic stuff we can do all at once with igraph
+climinfl.nodes <- data.frame(
+    name      = V(climinfl.ig)$name,
+    totdegree = igraph::degree(climinfl.ig, loops=FALSE),
+    indegree  = igraph::degree(climinfl.ig, mode="in", loops=FALSE),
+    outdegree = igraph::degree(climinfl.ig, mode="out", loops=FALSE),
+    eigen     = centr_eigen(climinfl.ig, directed = T)$vector,
+    bonanich  = power_centrality(climinfl.ig),
+    centr_clo = igraph::closeness(climinfl.ig),
+    centr_btw = igraph::betweenness(climinfl.ig, directed = FALSE),
+    # igraph only
+    burt      = constraint(climinfl.ig)
+)
+# Network-level measures:
+#   closeness centralization
+climinfl.centr_clo = centr_clo(climpref.ig)$centralization
+#   betweenness centralization
+climinfl.centr_btw = centr_betw(climpref.ig, directed = FALSE)$centralization
+
+# statnet version
+climinfl.nodes <- data.frame(
+    name      = climinfl.stat %v% "vertex.names",
+    totdegree = sna::degree(climinfl.stat),
+    indegree  = sna::degree(climinfl.stat, cmode = "indegree"),
+    outdegree = sna::degree(climinfl.stat, cmode = "outdegree"),
+    eigen     = evcent(climinfl.stat, ignore.eval = TRUE),
+    bonanich  = bonpow(climinfl.stat),
+    centr_clo = sna::closeness(climinfl.stat, gmode = "graph"),
+    centr_btw = sna::betweenness(climinfl.stat, gmode = "graph")
+)
+# Network-level measures:
+#   closeness centralization
+climinfl.centr_clo = centralization(climinfl.stat, FUN = "closeness", mode = "graph")
+#   betweenness centralization
+climinfl.centr_btw = centralization(climinfl.stat, FUN = "betweenness", mode = "graph")
+
+# Statnet-only: Gould-Fernandez Brokerage
+temp <- data.frame(brokerage(climinfl.stat, cl = climinfl.nodes$orgtype5)$z.nli)
+climinfl.nodes <- climinfl.nodes %>%
+  mutate(broker.tot = temp$t,
+         broker.coord = temp$w_I,
+         broker.itin = temp$w_O,
+         broker.rep = temp$b_IO,
+         broker.gate = temp$b_OI,
+         broker.lia = temp$b_O)
+
+# Calculated measures not specific to igraph or statnet:
+# Build the derived and reflected centrality (dc/rc) measures
+# "To calculate the proportion of centrality that is received, we first
+# square the adjacency matrix. The diagonal of the adjacency matrix is
+# equal to the the square of node degree. We then divide this diagonal
+# (squared degree) by total squared indegree (calculated by rowSums) to get
+# the proportion of received centrality."
+
+mat_climinfl <- as.matrix(as_adjacency_matrix(climinfl.ig))  # not " attr='weight'"
+mat_climinfl_sq <- t(mat_climinfl) %*% mat_climinfl
+climinfl.nodes$rc <- diag(mat_climinfl_sq) / rowSums(mat_climinfl_sq)
+# replace missing values with 0
+climinfl.nodes$rc <- ifelse(is.nan(climinfl.nodes$rc), 0, climinfl.nodes$rc)
+climinfl.nodes$dc <- 1 - climinfl.nodes$rc
+
+# Build the derived and reflected eigenvector measures
+climinfl.nodes$eigen.rc <- climinfl.nodes$eigen * climinfl.nodes$rc
+climinfl.nodes$eigen.dc <- climinfl.nodes$eigen * climinfl.nodes$dc
+```
+
+### Closeness centrailty
+
+From text: "The closeness centrality of a node is defined as the sum of the geodesic distances between that node and all other nodes in a network."
+
+Both are called in the above blocks, with `sna::closeness()` or `igraph::closeness()` on the respective network object.
+
+From igraph::closeness help: "Closeness centrality measures how many steps is required to access every other vertex from a given vertex."
+
+igraph and statnet have very different implementations, with options that have to be carefully set.
+
+#### igraph
+
+igraph uses inverse closeness.
+
+For directed networks, use `mode=("in", "out", "all", "total")`, describing the path type; `in` is paths *to* a vertex, `out` is paths *from* a vertex.
+Undirected networks ignore this parameter.
+It will use the "weight" edge attribute automatically if it's there, or can be overriden with something else.
+
+#### statnet
+
+Must specify `gmode` (type of graph) as `graph` (undirected) or `digraph` (directed, default).
+`cmode` (type of closeness centrality being measured) is one of: `directed`, `undirected` (both standard closeness), `suminvdir` (directed case) and `suminvundir` (undirected case), and `gil-schmidt` for that.
+The `suminv` options correspond to igraph's default inversion, though they're still calculated slightly differently, so they're generally preferred.
+
+statnet ignores the edge weights by default; `ignore.eval = FALSE` to use them, according to the documentation, but the results appear not to use them.
+
+`sna::closeness(climpref.stat, gmode="graph", cmode="suminvundir", ignore.eval=FALSE))`
+
+### Closeness centralization
+
+Closeness centralization is the network-level measure of the *closeness centrality* node-level measure.
+
+```r
+climinfl.centr_clo = centr_clo(climpref.ig)$centralization
+climinfl.centr_clo = centralization(climinfl.stat, FUN = "closeness", mode = "graph")
+```
+
+### Betweenness centrality
+
+Betweenness centrality is the node-level measure of the number of geodesics (shortest path between two nodes) on which a node sits.
+A high betweenness centrality measure means a node is on many shortest-paths, suggesting a measure of influence or power.
+
+```r
+igraph::betweenness(climpref.ig, directed = FALSE)
+sna::betweenness(climpref.ig, gmode = "graph")
+```
+
+The igraph version `directed` argument, for whether or not direction should be considered, defaults to `TRUE`; it might be wondered why it doesn't read the directedness of the graph as a default, but oh well.
+
+The statnet version would use `gmode` of `digraph` for a directed network, and `cmode` for a variant undirected form; see `?sna::betweenness` for more.
+Statnet appears to use weights by default; `weights = NA` disables weights in igraph.
+
+### Betweenness centralization
+
+The network-level measure of betweenness centralization represents [Freeman centralization](https://en.wikipedia.org/wiki/Centrality#Freeman_centralization), at least according to the statnet documentation.
+
+This is a "a measure of how central its most central node is in relation to how central all the other nodes are".
+
+```r
+climinfl.centr_btw = centr_betw(climpref.ig, directed = FALSE)$centralization
+climinfl.centr_btw = centralization(climpref.stat, FUN = "betweenness", mode = "graph")
+```
+
+### (Burt's) network constraint
+
+igraph-only function measuring a node's connection redundancy from 0 (none) to 1. Uses weights.
+
+```r
+constraint(climinfl.ig)
+```
+
+### Gould-Fernandez Brokerage
+
+(Statnet-only)
+
+From `?brokerage()`: "Gould and Fernandez (following Marsden and others) describe brokerage as the role played by a social actor who mediates contact between two alters."
+
+Requires a directed network with a vertex attribute used for grouping. Returns a structure with a lot of information is returned; tutorial refers mainly to `znli` containing the following roles:
+
+|prefix|Role|Action|Path|
+|---|---|---|---|
+|w_I|Coordinator| mediates contact between two individuals from his or her own group. |A -> A -> A|
+|w_O|Itinerant broker| mediates contact between two individuals from a single group to which he or she does not belong. |A -> B -> A|
+|b_{OI}|Gatekeeper| mediates an incoming contact from an out-group member to an in-group member. |A -> B -> B|
+|b_{IO}|Representative| mediates an outgoing contact from an in-group member to an out-group member. |A -> A -> B|
+|b_O|Liaison| mediates contact between two individuals from different groups, neither of which is the group to which he or she belongs. |A -> B -> C|
+|t|Total|Total (cumulative) brokerage role occupancy|(Any two)|
+
+```r
+brokerage(climinfl.stat, cl = climinfl.nodes$orgtype5)
+```

### 2022-03-20 _academics/ps753/ps753.Rmd
update index
build
Add hw4 link
Add to index:

@@
 - Some [useful R tidbits](../ps753-useful-r-tidbits/index.html) from the course material and related sources.
+- [Homework 3](../ps753-hw3/index.html) for the course: a network analysis of the Florentine Wedding data.
+- [Homework 4](../ps753-hw4/index.html) for the course: a centrality analysis of the Florentine Wedding data.
+- [Homework 5](../ps753-hw5/index.html) for the course: brokerage and betweenness analysis of the Florentine Wedding data.

### 2022-03-20 _academics/ps753-useful-r-tidbits/ps753-useful-r-tidbits.Rmd
Specify package for sna::evcent and bonpow

@@
     outdegree = sna::degree(climinfl.stat, cmode = "outdegree"),
-    eigen     = evcent(climinfl.stat, ignore.eval = TRUE),
-    bonanich  = bonpow(climinfl.stat),
+    eigen     = sna::evcent(climinfl.stat, ignore.eval = TRUE),
+    bonanich  = sna::bonpow(climinfl.stat),
     centr_clo = sna::closeness(climinfl.stat, gmode = "graph"),

### 2022-03-20 _academics/ps753-hw5/ps753-hw5.Rmd
knit

@@
 ```
+
+```{r}
+data(flo)
+ntwk.ig <- graph_from_adjacency_matrix(flo)
+
+ntwk.vcount <- vcount(ntwk.ig)
+ntwk.ecount <- ecount(ntwk.ig)
+ntwk.is_bipartite <- is_bipartite(ntwk.ig)
+ntwk.is_directed <- is_directed(ntwk.ig)
+ntwk.is_weighted <- is_weighted(ntwk.ig)
+ntwk.vertex_attr_names <- vertex_attr_names(ntwk.ig)
+ntwk.edge_attr_names <- edge_attr_names(ntwk.ig)
+ntwk.names <- V(ntwk.ig)$name
+ntwk.dyad_census <- igraph::dyad.census(ntwk.ig)
+ntwk.triad_census <- igraph::triad_census(ntwk.ig)
+ntwk.global_clustering_coef <- transitivity(ntwk.ig, type = "global")
+ntwk.local_avg_clustering_coef <- transitivity(ntwk.ig, type = "average")
+ntwk.avg_path_length <- average.path.length(ntwk.ig, directed = T)
+ntwk.component_names <- names(igraph::components(ntwk.ig))
+ntwk.components_no <- igraph::components(ntwk.ig)$no
+ntwk.components_csize <- igraph::components(ntwk.ig)$csize
+
+# distances(ntwk.ig,"Acciaiuoli","Strozzi", weights=NA)
+
+ntwk.graph_density <- graph.density(ntwk.ig)
+ntwk.graph_density_noloops <- graph.density(ntwk.ig, loops = TRUE)
+
+ntwk.avg_degree <- igraph::degree(ntwk.ig)
+
+# Get the degree info and make the data frame
+
+ntwk.nodes <- data.frame(
+  name = V(ntwk.ig)$name, 
+  degree = igraph::degree(ntwk.ig),
+  indegrees = igraph::degree(ntwk.ig, mode = "in", loops = FALSE),
+  outdegrees = igraph::degree(ntwk.ig, mode = "out", loops = FALSE)
+  )
+# ntwk.nodes[ntwk.nodes$name %in% c("Medici", "Strozzi"),]
+
+# Get degree centralization
+
+ntwk.indegree_centralization <- 
+  centr_degree(ntwk.ig, loops = FALSE, mode = "in")$centralization
+ntwk.outdegree_centralization <- 
+  centr_degree(ntwk.ig, loops = FALSE, mode = "out")$centralization
+ntwk.total_centralization <- 
+  centr_degree(ntwk.ig, loops = FALSE, mode = "total")$centralization
+```
+
+Building on [last week's look](../ps753-hw3/index.html) at basic network centrality attributes in the Florentine Wedding dataset, we're going to look at two new measurements of centrality: brokerage and betweenness.
+
+One of the under-the-hood consequences of this is that we must switch from iGraph to Statnet for this analysis, as brokerage is only found in Stanet.
+
+```{r}
+# statnet version
+climinfl.nodes <- data.frame(
+    name      = climinfl.stat %v% "vertex.names",
+    totdegree = sna::degree(climinfl.stat),
+    indegree  = sna::degree(climinfl.stat, cmode = "indegree"),
+    outdegree = sna::degree(climinfl.stat, cmode = "outdegree"),
+    eigen     = sna::evcent(climinfl.stat, ignore.eval = TRUE),
+    bonanich  = sna::bonpow(climinfl.stat),
+    centr_clo = sna::closeness(climinfl.stat, gmode = "graph"),
+    centr_btw = sna::betweenness(climinfl.stat, gmode = "graph")
+)
+# Network-level measures:
+#   closeness centralization
+climinfl.centr_clo = centralization(climinfl.stat, FUN = "closeness", mode = "graph")
+#   betweenness centralization
+climinfl.centr_btw = centralization(climinfl.stat, FUN = "betweenness", mode = "graph")
+
+# Statnet-only: Gould-Fernandez Brokerage
+temp <- data.frame(brokerage(climinfl.stat, cl = climinfl.nodes$orgtype5)$z.nli)
+climinfl.nodes <- climinfl.nodes %>%
+  mutate(broker.tot = temp$t,
+         broker.coord = temp$w_I,
+         broker.itin = temp$w_O,
+         broker.rep = temp$b_IO,
+         broker.gate = temp$b_OI,
+         broker.lia = temp$b_O)
+```

### 2022-03-21 _academics/ps753-hw5/ps753-hw5.Rmd
HW5 and knit

--- 2022-03-20 _academics/ps753-hw5/ps753-hw5.Rmd
@@
 ---
-title: "Homework 5: Brokerage and Betweenness in the Florentine Wedding dataset"
+title: "Homework 5: Constraint and Brokerage in the Florentine Wedding dataset"
 description: |
@@
 library(igraph)
+library(sna)
 library(tidyverse)
 ```
+
+```{r}
+data("florentine", package = "ergm")
+ntwk_stat <- flomarriage
+
+data(flo)
+ntwk_ig <- graph_from_adjacency_matrix(flo)
+
+```
+
+Building on [last week's look](../ps753-hw3/index.html) at basic network centrality attributes in the Florentine Wedding dataset, we're going to look at two new measurements of centrality: brokerage and constraints.
+
+(Note: these two concepts begin to break the "use either one" conclusion when choosing statnet or igraph; brokerage is a function of statnet, and constraints are a function of igraph.)
+
+```{r}
+
+ntwk_ig.nodes <- data.frame(
+    name      = V(ntwk_ig)$name,
+    totdegree = igraph::degree(ntwk_ig, loops = FALSE),
+    indegree  = igraph::degree(ntwk_ig, mode = "in", loops = FALSE),
+    outdegree = igraph::degree(ntwk_ig, mode = "out", loops = FALSE),
+    eigen     = centr_eigen(ntwk_ig, directed = T)$vector,
+    bonanich  = power_centrality(ntwk_ig),
+    centr_clo = igraph::closeness(ntwk_ig),
+    centr_btw = igraph::betweenness(ntwk_ig, directed = FALSE),
+    # igraph only
+    burt      = constraint(ntwk_ig)
+)
+# Network-level measures:
+#   closeness centralization
+ntwk_ig.centr_clo = centr_clo(ntwk_ig)$centralization
+#   betweenness centralization
+ntwk_ig.centr_btw = centr_betw(ntwk_ig, directed = FALSE)$centralization
+
+ntwk_stat.nodes <- data.frame(
+    name      = ntwk_stat %v% "vertex.names",
+    totdegree = sna::degree(ntwk_stat),
+    indegree  = sna::degree(ntwk_stat, cmode = "indegree"),
+    outdegree = sna::degree(ntwk_stat, cmode = "outdegree"),
+    eigen     = sna::evcent(ntwk_stat, ignore.eval = TRUE),
+    bonanich  = sna::bonpow(ntwk_stat),
+    centr_clo = sna::closeness(ntwk_stat, gmode = "graph",
+                               cmode = "suminvundir"),
+    centr_btw = sna::betweenness(ntwk_stat, gmode = "graph")
+)
+# Network-level measures:
+#   closeness centralization
+ntwk_stat.centr_clo = sna::centralization(ntwk_stat.nodes, sna::closeness, 
+                                     mode = "graph")
+#   betweenness centralization
+ntwk_stat.centr_btw = sna::centralization(ntwk_stat.nodes, sna::betweenness,
+                                     mode = "graph")
+
+# Statnet-only: Gould-Fernandez Brokerage
+# temp <- data.frame(brokerage(ntwk_stat, cl = ntwk.nodes$orgtype5)$z.nli)
+# ntwk.nodes <- ntwk.nodes %>%
+#   mutate(broker.tot = temp$t,
+#          broker.coord = temp$w_I,
+#          broker.itin = temp$w_O,
+#          broker.rep = temp$b_IO,
+#          broker.gate = temp$b_OI,
+#          broker.lia = temp$b_O)
+```
+
+## Burt's constraint
+
+*Burt's constraint* can be thought of as a measure of a node's dependency on other nodes for connections or information.
+If a node has a small number of connections, or the connections are largely redundant (i.e.
+a small group of friends without many other connections), that node has a higher constraint than one that has connections to nodes that are less redundant to each other.
+It provides more pathways, or opportunities, for information or power to flow.
+
+It should come as no surprise, given all of the various ways that the Medici are the dominant node in this dataset, that the Medici family has the lowest overall constraint.
+
+As a reminder, here is the original network graph of the Florence wedding data with no scaling applied:
+
+```{r}
+set.seed(1235)
+plot(
+  ntwk_ig,
+  layout = layout_components(ntwk_ig),
+  edge.arrow.mode = 0,
+  vertex.label.color = "black",
+  vertex.shape = "circle",
+  vertex.label.font = 2,
+  main = "Florence interfamily marriages, 1282-1500"
+)
+```
+
+And here we scale the nodes by their Burt's constraint measure:
+
+```{r}
+
+ntwk_ig_b <- ntwk_ig
+ntwk_ig_b.nodes <- ntwk_ig.nodes
+ntwk_ig_b.nodes$burt <- ifelse(is.na(ntwk_ig_b.nodes$burt), 1, ntwk_ig_b.nodes$burt)
+
+ntwk_ig_b.nodes$burt_inv <- 5 * (1 - ntwk_ig_b.nodes$burt)
+V(ntwk_ig_b)$size <- ntwk_ig_b.nodes[,"burt_inv"] * 5
+V(ntwk_ig_b)$label.cex <- ntwk_ig_b.nodes[,"burt_inv"] / 4
+V(ntwk_ig_b)$label.cex <- pmax(V(ntwk_ig_b)$label.cex, 1)
+
+# Ensure minimum node size
+V(ntwk_ig_b)$size <- pmax(V(ntwk_ig_b)$size, 1)
+
+burts_labels <- sprintf("%s\n%s", V(ntwk_ig_b)$name,
+                            # Omit leading "0" can't be done in sprintf
+                            substr(as.character(round(
+                              ntwk_ig_b.nodes$burt, 3
+                            )), 2, 10))
+set.seed(1235)
+plot(
+  ntwk_ig_b,
+  layout = layout_components(ntwk_ig_b),
+  vertex.label = burts_labels,
+  edge.arrow.mode = 0,
+  vertex.label.color = "black",
+  vertex.shape = "circle",
+  vertex.label.font = 2,
+  main = "Florence families by Burt's constraint"
+)
+
+```
+
+Once again, we see the clear dominance of the Medici family in this measure.
+Note that all of the pendant nodes (Lamberteschi, Ginori, Pazzi and Acciaiuoli) have maximal constraints of 1; they are completely constrained by their single connections, by definition.
+The Pucci family's constraint is undefined, as it is a disconnected node.
+
+The Medici have the lowest constraints of any family, being well-connected to other well-connected nodes with some spread among them.
+It is interesting to note that the Guadagni family is not far behind, though, at 0.25 to 0.21 for the Medici.
+They are connected to 3 well-connected nodes besides the Lamberteschi, and the Bischeri node in particular gives them indirect access to the Strozzi and Peruzzi nodes, besides the Albizzi and Tornabouni nodes that they share with the Medici.
+It can be seen from this that the Bischeri node is a strength for the Guadagni family in much the same way that the Barbadori node is a strength of the Medici.
+
+## Brokerage
+
+With Gould-Fernandez Brokerage, we begin to look at the ways in which nodes mediate connections between other groups of nodes.
+This requires a directed network, since the direction of flow of access is a crucial aspect of the core concept.
+There are five roles recognized by Gould-Fernandez Brokerage:
+
+Coordinator
+: mediates contact between two individuals from his or her own group
+
+Itinerant broker
+: mediates contact between two individuals from a single group to which he or she does not belong
+
+Gatekeeper
+: mediates an incoming contact from an out-group member to an in-group member
+
+Representative
+: mediates an outgoing contact from an in-group member to an out-group member
+
+Liaison
+: mediates contact between two individuals from different groups, neither of which is the group to which he or she belongs
+
+Unfortunately for the present moment, however, we cannot apply this concept to the Florence wedding data set, because at present there are no subgroups or categories between which to identify brokers!
+We will need to add data to the set to find a basis for this sort of analysis, and this is where we begin to transition to a research project.
+
+I am interested in the patronage system in Renaissance Florence, particularly as it applied to artists and scientists, as a measure of influence of the families.
+We may be able to return to brokerage concepts once we map in data related to this concept.

### 2022-03-21 _academics/ps753-useful-r-tidbits/ps753-useful-r-tidbits.Rmd
HW5 and knit

--- 2022-03-20 _academics/ps753-useful-r-tidbits/ps753-useful-r-tidbits.Rmd
@@
     outdegree = sna::degree(climinfl.stat, cmode = "outdegree"),
-    eigen     = evcent(climinfl.stat, ignore.eval = TRUE),
-    bonanich  = bonpow(climinfl.stat),
-    centr_clo = sna::closeness(climinfl.stat, gmode = "graph"),
+    eigen     = sna::evcent(climinfl.stat, ignore.eval = TRUE),
+    bonanich  = sna::bonpow(climinfl.stat),
+    centr_clo = sna::closeness(climinfl.stat, gmode = "graph",
+                               cmode = "suminvundir"),
     centr_btw = sna::betweenness(climinfl.stat, gmode = "graph")
@@
 #   closeness centralization
-climinfl.centr_clo = centralization(climinfl.stat, FUN = "closeness", mode = "graph")
+climinfl.centr_clo = centralization(climinfl.stat, sna::closeness, mode = "graph")
 #   betweenness centralization
-climinfl.centr_btw = centralization(climinfl.stat, FUN = "betweenness", mode = "graph")
+climinfl.centr_btw = centralization(climinfl.stat, sna::betweenness, mode = "graph")
 

### 2022-03-24 _academics/ps753-hw5/ps753-hw5.Rmd
fix link

@@
 
-Building on [last week's look](../ps753-hw3/index.html) at basic network centrality attributes in the Florentine Wedding dataset, we're going to look at two new measurements of centrality: brokerage and constraints.
+Building on [last week's look](../ps753-hw4/index.html) at basic network centrality attributes in the Florentine Wedding dataset, we're going to look at two new measurements of centrality: brokerage and constraints.
 

### 2022-03-26 _academics/ps753-hw6/ps753-hw6.Rmd
initial commit

---
title: "Homework 6: Clustering and Blockmodels in the Florentine Wedding dataset"
description: |
  Yet more ways to look at the Florence data.
date: 2022-03-26
author:
  - name: Steve Linberg
    url: {}
output:
  distill::distill_article:
    self_contained: false
    code_folding: true
    toc: true
categories:
  - r
  - social networks
---

<p style="background: #eee">
[<i class="fas fa-home"></i>](/) &gt; Academics &gt; [PS753](../ps753/index.html)
&gt; `r rmarkdown::metadata$title`
</p>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	fig.height = 8,
	message = FALSE,
	warning = FALSE
)
library(network)
library(igraph)
library(sna)
library(tidyverse)
```

This week we will look at structural equivalence and blockmodels in the Florentine wedding data set. Structural equivalence is a way to look at subsections of a network that identifies nodes with similar connections to other nodes, such as parents of the same child, or children of the same parent.

We begin, as usual, by loading the Florence data in both the igraph and statnet packages.

```{r}
# Load the statnet version
data("florentine", package = "ergm")
flo_stat <- flomarriage

# Load the igraph version
data(flo)
# If the matrix is symmetric, it is undirected.
# https://mathworld.wolfram.com/UndirectedGraph.html
# We know the Florence marriage network is undirected, but code is always better.
flo_ig <- graph_from_adjacency_matrix(flo, mode = ifelse(isSymmetric(flo), "undirected", "directed"))
```

The structural equivalence object is created with the default parameters, except for forcing the undirected mode. Printing the resulting object shows its creation parameters.

```{r}
flo_stat.se <- sna::equiv.clust(flo_stat, mode = "graph")
print(flo_stat.se)
```

The first thing to do is get a visual look at the default clustering dendrogram.

```{r}
plot(flo_stat.se, labels = flo_stat.se$glabels)
```

What does this show us? Well, as with nearly every other observation we've looked at in this data set, one of the most obvious observations is how the Medici are once again separate from the other nodes (families) in the set. The fact that they are alone in the top left shows that they are dissimilar to all other nodes in the network.

If we look at the lower left corner of the graph, cam see that the Peruzzi and Strozzi families are structurally similar. Let's look again at the basic network diagram, highlighting these families:

```{r}
set.seed(1235)

oranges <- colorRampPalette(c("gold", "skyblue"))
col <- oranges(5)

vcol <- rep("gray80", vcount(flo_ig))
vcol[which(V(flo_ig)$name == "Peruzzi")] <- col[1]
vcol[which(V(flo_ig)$name == "Strozzi")] <- col[1]

plot(
  flo_ig,
  layout = layout_components(flo_ig),
  edge.arrow.mode = 0,
  vertex.label.color = "black",
  vertex.shape = "circle",
  vertex.color = vcol,
  vertex.label.font = 2,
  main = "Florence interfamily marriages, 1282-1500"
)
```
They are not identical - there are no structurally identical nodes in this network - but they are similar. They both connect to the Bischeri and Castellani families.^[(Note that their connections to each other are not part of this calculation; it is only assessing nodes they have in common with each other.] The Strozzi also connect to the Ridolfi. This difference is represented by the height of the connector lines in the cluster dendrogram above.

Three other pairs of nodes show the same level of similarity:

```{r}
set.seed(1235)

vcol <- rep("gray80", vcount(flo_ig))
vcol[which(V(flo_ig)$name == "Peruzzi")] <- col[1]
vcol[which(V(flo_ig)$name == "Strozzi")] <- col[1]
vcol[which(V(flo_ig)$name == "Acciaiuoli")] <- col[2]
vcol[which(V(flo_ig)$name == "Barbadori")] <- col[2]
vcol[which(V(flo_ig)$name == "Ginori")] <- col[3]
vcol[which(V(flo_ig)$name == "Pucci")] <- col[3]
vcol[which(V(flo_ig)$name == "Pazzi")] <- col[4]
vcol[which(V(flo_ig)$name == "Salviati")] <- col[4]

plot(
  flo_ig,
  layout = layout_components(flo_ig),
  edge.arrow.mode = 0,
  vertex.label.color = "black",
  vertex.shape = "circle",
  vertex.color = vcol,
  vertex.label.font = 2,
  main = "Florence interfamily marriages, 1282-1500"
)
```

This seems believable, overall; the least intuitive impression on this graph is that the Pucci family, which is the network's lone isolate, is similar to the Ginori family on the other side of the graph, which connects only to the Albizzi family; nodes with similar structural equivalence are usually direct neighbors, since they *are* structurally equivalent due to their shared connections to other nodes.

Where this begins to be a bit less convincing is in the triad of the Tornabuoni, Albizzi and Lamberteschi families:

```{r}
set.seed(1235)

vcol <- rep("gray80", vcount(flo_ig))
vcol[which(V(flo_ig)$name == "Tornabuoni")] <- col[5]
vcol[which(V(flo_ig)$name == "Albizzi")] <- col[5]
vcol[which(V(flo_ig)$name == "Lamberteschi")] <- col[5]

plot(
  flo_ig,
  layout = layout_components(flo_ig),
  edge.arrow.mode = 0,
  vertex.label.color = "black",
  vertex.shape = "circle",
  vertex.color = vcol,
  vertex.label.font = 2,
  main = "Florence interfamily marriages, 1282-1500"
)
```
The Tornabuoni and Albizzi seem reasonably similar, sharing 2 other connections with each other, but the Lamberteschi share only one with either of them (the Guadagni). It seems like a bit of a stretch to suggest that they are of the same level of similarity to each other, but it could be argued that the Lamberteschi are more similar to the Tornabuoni and Albizzi than to any other nodes in the network.

Finally, it's instructive to look at the three nodes with no structurally similar peers: the Medici, Guadagni and Ridolfi:

```{r}
set.seed(1235)

vcol <- rep("gray80", vcount(flo_ig))
vcol[which(V(flo_ig)$name == "Medici")] <- col[1]
vcol[which(V(flo_ig)$name == "Guadagni")] <- col[3]
vcol[which(V(flo_ig)$name == "Ridolfi")] <- col[5]

plot(
  flo_ig,
  layout = layout_components(flo_ig),
  edge.arrow.mode = 0,
  vertex.label.color = "black",
  vertex.shape = "circle",
  vertex.color = vcol,
  vertex.label.font = 2,
  main = "Florence interfamily marriages, 1282-1500"
)
```
These nodes are plainly quite different from any other nodes in the network, in terms of their connections.

If we wanted another way to describe the diagram - already quite complex despite the data set's relatively small size - one way we might seek to do that is by breaking the diagram into a manageable number of sub-clusters that are similar to each other. And one way to achieve that is to use rectangular clustering based on height, describing the similarity threshold we feel is appropriate. For example, if we set the height to 8, we get:

```{r}
plot(flo_stat.se, labels = flo_stat.se$glabels)
rect.hclust(flo_stat.se$cluster, h = 8)
```

This results in four clusters containing data (and two empty clusters). If we plot the results, we can see the clusters whose components the algorithm considers more similar to each other than to other nodes:

```{r}
vcol <- rep("gray80", vcount(flo_ig))
vcol[which(V(flo_ig)$name == "Peruzzi")] <- col[1]
vcol[which(V(flo_ig)$name == "Strozzi")] <- col[1]
vcol[which(V(flo_ig)$name == "Bischeri")] <- col[1]
vcol[which(V(flo_ig)$name == "Castellani")] <- col[1]
vcol[which(V(flo_ig)$name == "Ridolfi")] <- col[2]
vcol[which(V(flo_ig)$name == "Acciaiuoli")] <- col[3]
vcol[which(V(flo_ig)$name == "Barbadori")] <- col[3]
vcol[which(V(flo_ig)$name == "Albizzi")] <- col[3]
vcol[which(V(flo_ig)$name == "Tornabuoni")] <- col[3]
vcol[which(V(flo_ig)$name == "Lamberteschi")] <- col[3]
vcol[which(V(flo_ig)$name == "Ginori")] <- col[4]
vcol[which(V(flo_ig)$name == "Pucci")] <- col[4]
vcol[which(V(flo_ig)$name == "Pazzi")] <- col[4]
vcol[which(V(flo_ig)$name == "Salviati")] <- col[4]

plot(
  flo_ig,
  layout = layout_components(flo_ig),
  edge.arrow.mode = 0,
  vertex.label.color = "black",
  vertex.shape = "circle",
  vertex.color = vcol,
  vertex.label.font = 2,
  main = "Florence interfamily marriages, 1282-1500"
)

```

On the whole, this appears to be a pretty reasonable partitioning of the network into clusters of structurally similar nodes. As we observed earlier, the Medici and Guadagni families are distinct enough not to have any structurally equivalent peers in the network, and the Ridolfi, while itself also lacking structurally similar peers, is still more similar to the other clusters than it is to the Medici or Guadagni.

### Blockmodel partitions

Finally, a brief look at blockmodel partitions, which are another way to visualize the clusters described above. Keeping the same height parameter of 8 from the above diagram, we can generate a blockmodel partition from the data:

```{r}
flo_stat.bm <- blockmodel(flo_stat, flo_stat.se, h = 8, mode = "graph")
flo_stat.bm
```

```{r}

# Code from the tutorial
plot.block<-function(x=blk_mod, main=NULL, cex.lab=1){
  plot.sociomatrix(x$blocked.data, labels=list(x$plabels,x$plabels),
                   main=main, drawlines = FALSE, cex.lab=cex.lab)
  for (j in 2:length(x$plabels)) if (x$block.membership[j] !=
                                     x$block.membership[j-1]) 
    abline(v = j - 0.5, h = j - 0.5, lty = 3, xpd=FALSE)
}

plot.block(flo_stat.bm, cex.lab=.5)
```

THe four clusters are indicated by the dotted horizontal lines in the diagram; each black box is a connection between the nodes along the left and top axes. The more filled-in a partition is, the more similar the clustered elements are to each other; this is also indicated numerically in the chart above with measures closer to 1.

The most obvious observations to make here are:

1. The cluster with the highest inner similarity rating is 2, with a numeric rate of 0.83 (on a scale of 0 to 1). This can be interpreted to mean that the Peruzzi, Strozzi, Bischeri and Castallani familes mostly intermarried among each other. (Cousins!)

2. The grouping returned by `rect.hclust` with the same `h` parameter of 8 does not match the grouping returned by `blockmodel` with the same value. It gets the first three groups correct, but all of the remaining nodes are packed together into the fourth group. I suspect this may be a bug in the blockmodel code, related to the fact that there were 4 clusters in the `rect.hclust` code that had data, and 2 without, but blockmodel considered the empty ones as valid, and then stopped at 4. This will need to be investigated.

### 2022-03-26 _academics/ps753-useful-r-tidbits/ps753-useful-r-tidbits.Rmd
General updates

@@
 ```r
-igraph::degree(trade2007.ig, v = V(trade2007.ig)[c("China", "Canada", "United Kingdom", "Denmark")])
+> igraph::degree(trade2007.ig, v = V(trade2007.ig)[c("China", "Canada", "United Kingdom", "Denmark")])
          China         Canada United Kingdom        Denmark 
            364            364            364            362 
+```
+
+This can also be done by index, as in:
+
+```r
+> igraph::degree(flo_ig, v = 1:3)
+Acciaiuoli    Albizzi  Barbadori 
+         2          6          4 
+```
+
+Statnet degrees aren't named:
+
+```r
+> sna::degree(flo_stat)
+ [1]  2  6  4  6  6  2  8  2 12  2  6  0  6  4  8  6
+> which(sna::degree(flo_stat) == 0)
+[1] 12
 ```
@@
 Component structure and membership
+
+*Note*: a graph is fully connected if its number of components is 1.
+igraph returns this as the `no` parameter of `igraph::components()`; it appears that statnet has no parallel function, but if the number of isolates is 0, the graph is connected (`sna::isolates()`).
 
@@
 # Statnet-only: Gould-Fernandez Brokerage
-temp <- data.frame(brokerage(climinfl.stat, cl = climinfl.nodes$orgtype5)$z.nli)
+# replace ATTR with the vector of the desired attribute, such as
+# `climinfl.nodes$orgtype5`
+temp <- data.frame(brokerage(climinfl.stat, cl = ATTR)$z.nli)
 climinfl.nodes <- climinfl.nodes %>%
@@
 mat_climinfl_sq <- t(mat_climinfl) %*% mat_climinfl
+# alternately:
+mat_climinfl <- as.matrix.network(climinfl.stat, attr = "weight")
+diag(mat_climinfl) <- 0
+mat_climinfl_sq <- mat_climinfl %*% mat_climinfl
+
 climinfl.nodes$rc <- diag(mat_climinfl_sq) / rowSums(mat_climinfl_sq)
@@
 
-Requires a directed network with a vertex attribute used for grouping. Returns a structure with a lot of information is returned; tutorial refers mainly to `znli` containing the following roles:
+Requires a directed network with a vertex attribute used for grouping. Returns a structure with a lot of information; tutorial refers mainly to `znli` containing the following roles:
 
@@
 ```
+
+## Week 6
+
+(All statnet unless otherwise specified)
+
+Calculating structural equivalence ("SE clusters"): look for nodes that have the same pattern of ties with the same neighbors, like siblings of a parent.
+Unexplained distance functions include "hamming", "correlation", "gamma".
+The `sedist()` function ignores edge values (weights?).
+
+```r
+flomarr.se <- equiv.clust(flomarr.stat, equiv.fun = "sedist", method = "hamming", mode = "graph")
+```
+
+The output of `equiv.clust()` can be plotted as a "Cluster Dendogram":
+
+```r
+flomarr.se <- equiv.clust(flomarriage, equiv.fun = "sedist", method = "hamming", mode = "graph")
+plot(flomarr.se,labels = flomarr.se$glabels)
+```
+
+Annoyingly, the different `cluster.method`s "single, average, or ward.D" are not explained in the tutorial. The default is "complete".
+
+`rect.hclust()` (visually) "cuts" the diagram at a given height, making separate clusters:
+
+```r
+plot(flomarr.se, labels = flomarr.se$glabels)
+rect.hclust(flomarr.se$cluster, h = 9)
+```
+
+```r
+flomarr.se <- equiv.clust(flomarriage, equiv.fun = "sedist", cluster.method = "single",
+                          method = "hamming", mode = "graph")
+plot(flomarr.se,labels = flomarr.se$glabels)
+
+```

### 2022-03-26 _academics/ps753/ps753.Rmd
HW5 and knit
Add link

@@
 - [Homework 4](../ps753-hw4/index.html) for the course: a centrality analysis of the Florentine Wedding data.
-- [Homework 5](../ps753-hw5/index.html) for the course: brokerage and betweenness analysis of the Florentine Wedding data.
+- [Homework 5](../ps753-hw5/index.html) for the course: constraint and brokerage analysis of the Florentine Wedding data.
+- [Homework 6](../ps753-hw6/index.html) for the course: clustering and blockmodel analysis of the Florentine Wedding data.

### 2022-04-02 _posts/2022-04-02-quickie-for-importing-tasks-to-omnifocus-3/quickie-for-importing-tasks-to-omnifocus-3.Rmd
new post

---
title: "Quickie for importing tasks to OmniFocus 3"
description: |
  Copy and paste formatted text to mass-create tasks.
author:
  - name: Steve Linberg
    url: https://slinberg.net
date: 2022-04-02
output:
  distill::distill_article:
    self_contained: false
preview: https://www.omnigroup.com/assets/img/icons/omnifocus-web@2x.png
categories:
  - omnifocus
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

I've been an [OmniFocus](https://www.omnigroup.com/omnifocus/) user since version 1.0.

Well, sort of. I used it a ton in version 1.0 and then eventually stopped. Then I used it again a ton wehen version 2 came out, and then eventually stopped. It worked well when I really forced myself to use it, but eventually I'd run out of gas, and it always bugged me a bit that I couldn't get a handle on why. I read David Allen's "[Getting Things Done](https://gettingthingsdone.com/)", which it was inspired by, and it all made sense, but it still just had a lot of friction to keep using.

One of the problems, I now realize, was the notion of *contexts* - each to-do list could be assigned one, *and only one*, context, like "online" or "supermarket" or "home", describing the setting where a task could be done. The problem was that there were many tasks that could be done in multiple contexts, but no way to mark them as such. One user I talked with on one of the forums explained that the way to think about it was that I myself could be in multiple simultaneous contexts, like "home" and "online", and that tasks that matched either of those would appear. It sort of made sense, but it still never felt natural and I eventually dropped it again.

I checked it out again in version 3, and the problem has been solved: contexts are gone, replaced by the much more versatile and familiar concept of *tags*. Tags can be freely defined, and any item can have as many tags as you want. So now I'm using it again and finding it much easier to get along with.

(I also really appreciate that there is the option to just pay for it straight up, though they also offer the odious subscription-based pricing that everybody except developers hates. There is upgrade pricing from prior versions that makes it more affordable. They also offer [educational pricing](https://support.omnigroup.com/sales-educational-pricing/), so between my educational pricing and upgrade discount, OF3 went from \$99 to about \$38. Sold.)

It's full of really nice little interface touches that are almost invisible but that make life very pleasant. One of them is the ability to import tasks in [TaskPaper](https://guide.taskpaper.com/) format, which is very straightforward. I needed to add a list of 24 assignments (sigh) to my inbox, one per day for the next 24 days. I had them typed out, and with the help of a few [emacs macros](http://www.gnu.org/software/emacs/manual/html_node/emacs/Keyboard-Macros.html) added due dates and tags, using the following format:

```
- task description @tags(tag1, tag2) @due(YYYY-MM-DD HH:MM)
```

So my full list was:

```
- (ch 3) Simple linear regression @tags(Machine Learning, Homework) @projects(Machine Learning) @due(2022-04-02 23:59)
- (ch 3) Multiple linear regression @tags(Machine Learning, Homework) @due(2022-04-03 23:59)
- (ch 4) Logistic regression @tags(Machine Learning, Homework) @due(2022-04-04 23:59)
- (ch 4) Multiple logistic regression @tags(Machine Learning, Homework) @due(2022-04-05 23:59)
- (ch 4) Linear discriminant analysis @tags(Machine Learning, Homework) @due(2022-04-06 23:59)
- (ch 4) Quadratic discriminant analysis @tags(Machine Learning, Homework) @due(2022-04-07 23:59)
- (ch 4) Naive Bayes @tags(Machine Learning, Homework) @due(2022-04-08 23:59)
- (ch 4) K-Nearest Neighbors @tags(Machine Learning, Homework) @due(2022-04-09 23:59)
- (ch 4) Poisson regression @tags(Machine Learning, Homework) @due(2022-04-10 23:59)
- (ch 5) Cross-validation @tags(Machine Learning, Homework) @due(2022-04-11 23:59)
- (ch 5) Bootstrap @tags(Machine Learning, Homework) @due(2022-04-12 23:59)
- (ch 6) Best subset selection @tags(Machine Learning, Homework) @due(2022-04-13 23:59)
- (ch 6) Stepwise selection @tags(Machine Learning, Homework) @due(2022-04-14 23:59)
- (ch 6) Ridge regression @tags(Machine Learning, Homework) @due(2022-04-15 23:59)
- (ch 6) Lasso @tags(Machine Learning, Homework) @due(2022-04-16 23:59)
- (ch 6) Principal component regression @tags(Machine Learning, Homework) @due(2022-04-17 23:59)
- (ch 8) Bagging @tags(Machine Learning, Homework) @due(2022-04-18 23:59)
- (ch 8) Random forests @tags(Machine Learning, Homework) @due(2022-04-19 23:59)
- (ch 8) Boosting @tags(Machine Learning, Homework) @due(2022-04-20 23:59)
- (ch 8) Bayesian Additive Regression Trees @tags(Machine Learning, Homework) @due(2022-04-21 23:59)
- (ch 9) Support Vector Machines @tags(Machine Learning, Homework) @due(2022-04-22 23:59)
- (ch 12) Principal component analysis @tags(Machine Learning, Homework) @due(2022-04-23 23:59)
- (ch 12) K-means clustering @tags(Machine Learning, Homework) @due(2022-04-24 23:59)
- (ch 12) Hierarchical clustering @tags(Machine Learning, Homework) @due(2022-04-25 23:59)
```

Copy the entire list, and then just paste into the inbox, and voilà:

![Inbox](of_inbox.png)

Now the problem is I have to actually *do* all of those things. (But at least I can check off the top item for the new blog post.)

### 2022-04-02 _posts/2022-04-02-quickie-for-importing-tasks-to-omnifocus-3/quickie-for-importing-tasks-to-omnifocus-3.Rmd
typos

@@
 I used it a ton in version 1.0 and then eventually stopped.
-Then I used it again a ton wehen version 2 came out, and then eventually stopped.
+Then I used it again a ton when version 2 came out, and then eventually stopped.
 It worked well when I really forced myself to use it, but eventually I'd run out of gas, and it always bugged me a bit that I couldn't get a handle on why.
@@
 
-One of the problems, I now realize, was the notion of *contexts* - each to-do list could be assigned one, *and only one*, context, like "online" or "supermarket" or "home", describing the setting where a task could be done.
+One of the problems, I now realize, was the notion of *contexts* - each to-do item could be assigned to one, *and only one*, context, like "online" or "supermarket" or "home", describing the setting where a task could be done.
 The problem was that there were many tasks that could be done in multiple contexts, but no way to mark them as such.
@@
 ```
-- (ch 3) Simple linear regression @tags(Machine Learning, Homework) @projects(Machine Learning) @due(2022-04-02 23:59)
+- (ch 3) Simple linear regression @tags(Machine Learning, Homework) @due(2022-04-02 23:59)
 - (ch 3) Multiple linear regression @tags(Machine Learning, Homework) @due(2022-04-03 23:59)

### 2022-04-02 _posts/2022-04-02-quickie-for-importing-tasks-to-omnifocus-3/quickie-for-importing-tasks-to-omnifocus-3.Rmd
typos

@@
 
-One of the problems, I now realize, was the notion of *contexts* - each to-do item could be assigned to one, *and only one*, context, like "online" or "supermarket" or "home", describing the setting where a task could be done.
+One of the problems, I now realize, was the notion of *contexts* - each to-do list could be assigned one, *and only one*, context, like "online" or "supermarket" or "home", describing the setting where a task could be done.
 The problem was that there were many tasks that could be done in multiple contexts, but no way to mark them as such.

### 2022-08-18 steve.Rmd
text updates

@@
 
-I am a Masters student in the [DACSS program](https://www.umass.edu/sbs/data-analytics-and-computational-social-science-program) at the University of Massachusetts, Amherst; anticipated graduation Summer 2022.
-The day job is Senior Systems and Analytics Engineer at the [Gandara Center](https://www.gandaracenter.org/) in West Springfield, MA.
+In summer 2022, I completed the [Data Analytics and Computational Social Science (DACSS) Masters program](https://www.umass.edu/sbs/data-analytics-and-computational-social-science-program) at the [University of Massachusetts, Amherst](https://umass.edu).
+My present position is Senior Systems and Analytics Engineer at the [Gandara Center](https://www.gandaracenter.org/) in West Springfield, MA.
 
@@
 
-**Hampshire College** | Amherst, MA
+*MS, Data Analytics and Computational Social Science* | 2022
 
-BA, Social Science, 1990-3
+University of Massachusetts, Amherst | Amherst, MA
 
-**University of Massachusetts, Amherst** | Amherst, MA
+*BA, Social Science* | 1992
 
-M.S, Data Analytics and Computational Social Science | Anticipated graduation summer 2022
+Hampshire College | Amherst, MA
 
@@
 
-**Gandara Center** | Senior Systems and Analytics Engineer / CIO / IT Director | 2005 - Present
+**Gandara Center**, West Springfield, MA | Senior Systems and Analytics Engineer / CIO / IT Director | 2005 - Present
 
-**Silicon Goblin Technologies** | Owner / Programmer | 1999 - 2015
+**Silicon Goblin Technologies**, Amherst, MA | Owner / Programmer | 1999 - 2015
 
-**National Center on Adult Literacy**, University of Pennsylvania Graduate School of Education | Systems Programmer | 1997-1999
+**Rojo Networks**, San Francisco, CA | Systems Administrator and Customer Service Lead | 2006-2008
 
-**Holyoke Public Schools** | Literacy and GED Instructor | 1994-1997
+**National Center on Adult Literacy**, University of Pennsylvania Graduate School of Education, Philadelphia, PA | Systems Programmer and Lead Web Developer | 1997-1999
 
-**Unicom, Inc.** | Systems Programmer and Technical Writer | 1983-1997
+**Holyoke Public Schools**, Holyoke, MA | Literacy and GED Instructor | 1994-1997
+
+**Unicom, Inc.**, Providence, RI | Systems Programmer and Technical Writer | 1983-1997
 
@@
 
-- Data analytics and statistics
-- Programming languages
-- The state of the world
+- Data analytics, statistics, machine learning
+- Software engineering
+- Politics, culture and the state of the world
 - Drawing, painting, guitar, motorcycles, baking, gaming
